### 2014-05-07 rules/rules14.tex
Next draft of the 2014 rules.

\documentclass[12pt]{article}
\usepackage{times}
\usepackage{fullpage}
\usepackage{url}
\usepackage{epsf}

\newcommand{\akey}[1]{\textbf{#1}}

\begin{document}

\date{\small This version revised \the\year-\the\month-\the\day}

\title{DRAFT\\Satisfiability Modulo Theories Competition (SMT-COMP) 2014: Rules and 
Procedures}

% [morgan] do our own layout of authors; the four-author layout spacing
% was screwed up...
\def\doauthor#1{{%
  \hsize.5\hsize \advance\hsize by-1cm %
  \def\\{\hss\egroup\hbox to\hsize\bgroup\strut\hss}%
  \vbox{\hbox to\hsize\bgroup\strut\hss#1\hss\egroup}}}%

\def\header#1{\medskip\noindent\textbf{#1}}

\author{
David R. Cok \\
GrammaTech, Inc. \\
Ithaca, NY (USA) \\
cok@frontiernet.net \\
\and
David Deharbe \\
Federal University of Rio Grande do Norte \\
Brazil \\
David.Deharbe@loria.fr \\
\and
Tjark Weber \\
Uppsala University \\
Sweden \\
tjark.weber@it.uu.se \\ 
}

\maketitle

\def\eg{\textit{e.g.}}
\def\ie{\textit{i.e.}}

\noindent Comments on this document should be emailed to the smtcomp mailing
list or, if necessary, directly to the organizers (cf. email addresses in the following section and in the authors' affiliation information).

\section{Important dates and information}
\begin{itemize}
\item May 15 - deadline for new benchmarks
\item May 25 - deadline for solver entrants
\item June 1 - deadline for final versions of solvers (but no new entrants after May 25).
\item July 17 - SMT workshop; completion of competition.
\item SMT-COMP mailing list: smt-comp@cs.nyu.edu
\item signup site for mailing list: \url{cs.nyu.edu/mailman/listinfo/smt-comp}
\item competition web site: \url{www.smtcomp.org}  or \url{smtcomp.sourceforge.net/2014}
\end{itemize}

\section{Introduction}
\label{sec:intro}

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{PSS02,leberre+03}.  More information
on the history and motivation for SMT-COMP can be found at the
SMT-COMP web site, \url{www.smtcomp.org}, and in reports on previous
competitions~(\cite{SMTCOMP-2008,BDOS08,SMTCOMP-FMSD,SMTCOMP-JAR,SMTCOMP2012,SMTEVAL2013}).
SMT-COMP 2014 is affiliated with the SMT workshop (\url {http://smt2014.it.uu.se/}) at the co-located meetings of CAV, IJCAR, and SAT at FLoC 2014 (\url{http://vsl2014.at/}); the competition is part of the
FLoC Olympic Games (\url{http://vsl2014.at/olympics/}).

Accordingly, researchers are highly encouraged to submit both new benchmarks
and new or improved solvers to raise the level of competition and advance
the state-of-the-art in automated SMT problem solving.

SMT-COMP 2014 will have two tracks: the conventional main track and a repeat of 2012's application track.
Within a track there are multiple divisions, where each division
uses benchmarks from a specific SMT-LIB logic (or group of logics).

The rest of this document, revised from the previous 
version,\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, Aaron Stump, Morgan Deters, David Cok, Alberto Griggio, and Roberto Bruttomesso.}
describes the rules and competition procedures for SMT-COMP~2014.
The principal changes from the previous competition rules are the following:
\begin{itemize}
\item The competition will be run with StarExec (\url{www.starexec.org}). Accordingly we are simplifying the competition somewhat, since some of the infrastructure needs to be recreated.
\item We will hold just a main and application track. [ The demonstration unsat core, and proof generation tracks will be deferred until benchmarks and result analysis tools are reorganized.]
\item In the main track, we will measure comparative performance within all divisions to which at least two solvers are submitted. We will use the data to determine both sequential and parallel winners.

\item We are retaining and accenting last year's ``application track" and
  encourage submission of benchmarks for this track and competition against these
  benchmarks.

\item We will recognize winners as measured by number of benchmarks solved; we will also recognize best open source contributions and best new entrants, as described below. 
Also, importantly, the FLoC Olympic Games is sponsoring award ceremonies (with actual medals). A selection of the SMT-COMP winners (as described below) will be awarded FLoC Olympic Games medals.

\end{itemize}

All results of the competition will be made public. However, 
winners will be declared only for divisions in which at least two solvers are submitted.
Although the organizers may add other solvers for comparison purposes, only solvers that
are explicitly submitted by their authors are eligible to be designated as winners.

It is important for competitors to track discussions on the SMT-COMP mailing
list, as clarifications and any updates to these rules will be posted there.

\section{Entrants}
\label{sec:entrants}

\header{Solver format.} %
An entrant to SMT-COMP is an SMT solver
submitted by its authors using the Star-Exec (\url{http://www.starexec.org}) service.  
The execution service enables members of the
SMT research community to run solvers on jobs consisting of benchmarks
from the SMT-LIB benchmark library.  Jobs are run on a shared computer
cluster.  The execution service is provided free of
charge, but it does require registration to create a login account.  Registered users
may then upload their own solvers to run, or may run public solvers
already uploaded to the service.

For participation in SMT-COMP, a solver must be
uploaded to Star-Exec, made publicly available, and the organizers informed of its presence {\em and the logic divisions in which it is participating}.
A submission must also include a 32-bit unsigned
integer.  These numbers, collected from all submissions, are used
to seed the pseudo-random benchmark selection algorithm, as well
as the benchmark scrambler.

\header{System description.} %
As part of a submission,
SMT-COMP entrants are encouraged to provide a short (1--2 pages) description of
the system.  This should include a list of all authors of the system,
 their present institutional affiliations, and any appropriate acknowledgements.  The programming
language(s) and basic SMT solving approach employed should be
described (\eg, lazy integration of a Nelson-Oppen combination with
SAT, translation to SAT, etc.).  System descriptions are encouraged to
include a URL for a web site for the submitted tool.

\header{Other systems.} %
The organizers' intent is to promote as wide a comparison among solver options as possible.
However, if the number of solver submissions is too large for the
computational resources available to the competition, the organizers reserve the right not to
accept multiple versions of solvers from the same solver team.  The organizers also reserve the right to
include other systems of interest (such as entrants in previous competitions) in the
competition.

\header{Wrapper tools.} %
A \emph{wrapper tool} is defined as any tool
which calls one or more SMT solvers not written by the author of the wrapper
tool.  The other solvers are called the \emph{wrapped} tools.  
A wrapper tool must explicitly acknowledge any tools that it wraps.
Its system description should make clear the technical innovations  by which the
wrapper tool expects to improve on the wrapped tools.

\header{Attendance.} %
As with previous SMT-COMPs, submitters of an SMT-COMP entrant need not 
be physically present at the competition to
participate or win.

\subsection*{Deadlines} %

%%%%%% Make main and parallel competitions together.

\paragraph{Main competition track.} %
SMT-COMP entries must be submitted via StarExec and email to the organizers by the end of May~25, 2014 (anywhere on earth).  At that time new entrants will be
closed to the public to prepare for the competition, with the
exception that resubmissions of existing entries will be accepted
until the end of June~1, 2014 (anywhere on earth).  We strongly encourage
participants to use this grace period \emph{only} for the
purpose of fixing any bugs that may be discovered and not for adding
new features as there will be no opportunity to do extensive testing
using StarExec after the original deadline on May~25.

The versions that are present on the execution service at the conclusion of the
grace period will be the ones used for the competition, and versions
submitted after this time will not be used.  The organizers reserve
the right to start the competition itself at any time after the open
of the New York Stock Exchange on the morning of June~2.  See
Section~\ref{sec:timeline} below for a full timeline.

\paragraph{Application track.} %
The same deadlines and procedure for submitting to the main track will be used
also for the application track. Submissions to both the application track and
the main competition are independent: participants must submit explicitly to
both events to participate in both, and they may submit different (or
differently configured) solvers to each.  Benchmarks will be scrambled also
for this division, using the same scrambler and seed as the main track.
Entrants should still include a system description, as for the regular
competition.

The infrastructure for the application track is in the process of being recreated for StarExec.
If it is not ready in time, the organizers may choose to drop this track or only run it as a demonstration track.

\section{Execution of Solvers}
\label{sec:exec}

Solvers will be publicly evaluated in the main and application tracks, each in several logic divisions. 
All divisions with at least two solver entrants are competition divisions, i.e., a winner will be announced. The other divisions are exhibition divisions, i.e., solvers are evaluated and results posted, but no awards are announced.

\subsection{Logistics}

\header{Dates of competition.}
%
We anticipate that the bulk of the main track of the competition will take place during
the weeks leading up to SMT 2014, from about June~2 to July~17.  Results will be
announced at SMT 2014 and in an awards ceremony scheduled by the FLoC Olympic Games coordinator.
  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

The organizers will prioritize the running of the competition tracks, and may shift the
time period or order of the competition or exhibition tracks in order to complete SMT-COMP
prior to the SMT workshop and awards ceremony.

\header{Input and Output.}
%
Participating solvers must read a single benchmark script presented on its standard input
channel. The script is in the concrete syntax of the SMT-LIB format,
version 2.0, though with a restricted set of commands.   A
\textbf{benchmark script} is just a text file containing a sequence of SMTLIBv2 commands in which:

\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
logic is the first command after any $\akey{set-option}$ commands described below.
% rb 25/3: added
% \item A (single) \akey{set-option :print-success false} command, sent just
%       after $\akey{set-logic}$. This is to
%       avoid the output ``\texttt{success}'' to interfere with the
%       answer of the solver to the benchmark. 
\item The $\akey{exit}$ command is the last command.
\item For tracks other than the application track, there is exactly one $\akey{check-sat}$ command,
following possibly several $\akey{assert}$ commands.
\item For the application track, there are one or more \akey{check-sat} commands, 
  each preceded by one or more \akey{assert} commands 
  and zero or more \akey{push 1} commands, 
  and followed by zero or more \akey{pop 1} commands.
\item Scripts for the application track will have an initial $\akey{set-option :print-success true}$ command.
\item There is at most one $\akey{set-info}$ command for \texttt{status}.
%%%%%%%%%%%%%%%%%%% TODO: Need to define the set-info command.
\item The formulas in the script belong to the benchmark's logic, with
any free symbols declared in the script.
\item Extra symbols are declared exactly once before any
  use, using $\akey{declare-sort}$
  %, $\akey{define-sort}$,
  %$\akey{declare-fun}$, or $\akey{define-fun}$.  
  or $\akey{declare-fun}$.
  They must be part of the allowed signature expansion for the logic.
  Moreover, all sorts declared with a $\akey{declare-sort}$ command must have zero arity.

\item No other commands besides the ones just mentioned may be used.

\end{enumerate}

\noindent The SMT-LIB format specification is publicly
available from the ``Documents'' section of the SMT-LIB
website~\cite{SMT-LIB}.  Solvers will be given formulas just from the
Problem Divisions indicated during their submission to SMT-COMP.

\subsection{Main track}
\label{sec:exec:main}

The main track competition will consist of selected benchmarks in each of the
logic divisions.
Each benchmark script will be presented to the standard input of the solver.
Each SMT-COMP entrant is then expected to attempt to report on its
standard output channel whether the formula is satisfiable
(``\texttt{sat}'', in lowercase, without the quotation marks) or unsatisfiable
(``\texttt{unsat}'').  An entrant may also report ``\texttt{unknown}''
to indicate that it cannot determine satisfiability of the formula.

The main track competition uses a StarExec postprocessor (named ``SMT2 results conserv") to accumulate the results.

\header{Timeouts.} %
Each SMT-COMP solver will be executed on an
unloaded competition machine for each given benchmark, up to a fixed
time limit.  The time limit is yet to be determined, but it is
anticipated to be 1500 seconds (25 minutes) of CPU time and the same limit on wall clock time.\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.} 
% TBD- TODO - how to constrain solvers to single processes?
Solvers that take more than this
time limit will be killed.  Solvers are allowed to spawn other
processes.  These will be killed at approximately the same time as the
first started process.  A
timeout scores the same as if the output is ``\texttt{unknown}''.

\header{Aborts and unparsable output.} %
Solvers which exit before the time
limit without reporting a result (\ie\ due to exhausting memory, crashing,
or producing output other than \texttt{sat}, \texttt{unsat}, or
\texttt{unknown})
will be considered to have aborted. 
An abort scores the same as if the output is ``\texttt{unknown}''. 
Also, as a further measure to prevent misjudgments of solvers,
any  ``\texttt{success}'' outputs will be 
ignored.\footnote{
Note that SMT-LIBv2 requires that a solver produce a ``\texttt{success}'' answer
after each \akey{set-logic}, \akey{declare-sort}, \akey{declare-fun} and
\akey{assert} command (among others), unless the option
\akey{:print-success} is set to false; ignoring the
\texttt{success} outputs therefore allows for submitting fully-compliant
solvers without the need of a wrapper script, while still allowing entrants
of previous competitions to run without changes.}

\header{Persistent state.} %
Solvers are allowed to create and write to
files and directories during the course of an execution, but they are
not allowed to read such files back during later executions.  Any
files written should be put in the directory in which the tool is
started, or in a subdirectory.

\subsection{Application track}
\label{sec:exec:application}

The application track evaluates SMT solvers when interacting
with an external verification framework, \eg, a model
checker. This interaction, ideally, happens by means of an online
communication between the model checker and the solver: the model
checker repeatedly sends queries to the SMT solver, which in turn
answers either \texttt{sat} or \texttt{unsat}.  In this interaction an SMT-solver is
required to accept queries incrementally via its standard input channel.

In order to facilitate the evaluation of the solvers in this track, we
will set up a ``simulation'' of the aforementioned interaction, as was done in 2012. In
particular each benchmark in the application track represents a realistic
communication trace, containing multiple \akey{check-sat} commands (possibly
with corresponding \akey{push 1}/\akey{pop 1} commands), which
is parsed by a {\em trace executor}. The trace executor serves the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to the SMT solver
      (through stdin);
\item it prevents ``look-ahead'' behaviours of SMT solvers;
\item it records time and answers for each call, possibly aborting the execution
      in case of a wrong answer;
\item it guarantees a fair execution for all solvers by abstracting from any possible
      crash, misbehaviour, etc. that may happen on the model checker side.
\end{itemize}

\header{Input and Output.}
Participating solvers will be connected to a trace executor 
which will incrementally send commands to the standard input channel of the solver
and read responses from the standard output channel of the solver.
The commands will be taken from an \textbf{incremental benchmark script},
which is an SMT-LIB 2.0 script which satisfies the rules for an application script given above.
Note also that the trace executor will send a single 
\akey{set-option :print-success true} command to the solver before 
sending commands from the incremental benchmark script.

\medskip
Solvers must respond to each command sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{set-logic}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
 with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands. There should be no \akey{get-unsat-core} or \akey{get-proof} commands in these benchmarks.

\header{Timeouts.}
A time limit is set for the whole execution of each application
benchmark (consisting of multiple queries).  We anticipate the
timeout to be around 25~minutes (as it has been in the past).\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.}

\section{Benchmarks and Problem Divisions}
\label{sec:theories}

In 2012 the competition ran only some logics as competitive divisions.
In 2014 we will revert to previous practice and run each logic as a competitive division, if there are two or more solvers submitted.

\subsection{Main track}

\header{Benchmark sources.} %
Benchmark formulas for each division of the main track
will be drawn from the SMT-LIB library. The deadline for new benchmarks is
May 15. The organizers will be checking and curating these until June 1;
the organizers reserve the option to exclude new benchmarks if any prove problematic for some reason.
SMT-COMP
attempts to give preference to benchmarks that are ``real-world,'' in
the sense of coming from or having some intended application outside
SMT.

\header{Benchmark availability.} % 
The deadline for new benchmarks is May 15, 2014.
The benchmarks will be made available as soon as possible after that, as they are checked and curated.
The goal for completing the set of benchmarks is June 1.
The set of selected benchmarks will be published when the
competition begins.

\header{Benchmark demographics.} %
In SMT-LIB, benchmarks are organized according to \emph{families}.  A benchmark
family contains problems that are similar in some significant way.  Typically
they come from the same source or application, or are all output by the same
tool.  \emph{Each top-level subdirectory within a division represents a distinct
family.}  
%

Each benchmark in SMT-LIB also has a \emph{category}.  There are four possible
categories:
%
\begin{itemize}
\item \emph{check.} These benchmarks are hand-crafted to test whether
  solvers support specific features of each division.  In particular,
  there are checks for integer completeness (\ie\ benchmarks that are
  satisfiable under the reals but not under the integers) and big
  number support (\ie\ benchmarks that are likely to fail if integers
  cannot be represented beyond some maximum value, such as
  $(2^{31}-1)$).

%  \noindent%\textbf{New for 2010:} 
%  Using the same random seed as
%  for benchmark selection and scrambling, 5 ``check'' benchmarks will
%  be randomly generated using Robert Brummayer's SMT fuzzing
%  tool~\cite{brummayer+09}.  The (exact version of this) tool will be
%  publicly available from the SMT-COMP~2012 web site before the
%  competition.  Since these benchmarks are generated after the random
%  seed is fixed, they cannot be known in advance to any competitors,
%  including the organizers.  The rationale for including these
%  benchmarks is to try to take a step towards stronger certification
%  that solvers are correct.  In future years, we envision SMT-COMP
%  requiring that solvers pass some kind of pre-qualifying round based
%  on SMT fuzzing.  The inclusion of these fuzzing benchmarks that are
%  not known in advance is intended to encourage (without requiring)
%  solver implementors to test their solvers using fuzzing or similar
%  bug-discovery techniques.

\item \emph{industrial.} These benchmarks come from some real application
      and are produced by tools such as bounded model checkers, static analyzers, extended
      static checkers, etc.
\item \emph{random.} These benchmarks are randomly generated.
\item \emph{crafted.} This category is for all other benchmarks.  Usually,
  benchmarks in this category are designed to be particularly difficult or to
  test a specific feature of the logic.
\end{itemize}

\header{Benchmark selection.} %

The benchmarks are assigned a difficulty. For 2014, the difficulty measure is the
CPU time taken by the best performing solver in the 2013 SMT-EVAL. These times will be posted 
as soon after May 15 as possible.

The selection of benchmarks for the competition will be biased towards more difficult benchmarks
as follows. First, the benchmark pool is culled as follows:
\begin{enumerate}

\item \textbf{Retire very easy benchmarks.} %
  Eliminate benchmarks for which all of the current solvers in SMT-EVAL-2013
  solved the benchmark in under 5 seconds,
  \emph{unless} doing so reduces the pool of
    benchmarks for the division to less than~300.

\item \textbf{Retire inappropriate benchmarks.} %
  The competition organizers will remove from the eligibility pool
  certain SMT-LIB benchmarks that are inappropriate or uninteresting
  for competition, or cut the size of certain benchmark families to
  avoid their over-representation.

\end{enumerate}

Then, for each division, the benchmarks are divided into quintiles according to the difficulty measure.
Then a random selection of N competition benchmarks is made as follows:
\begin{itemize}
\item N*40\% are taken from the top quintile, if sufficient benchmarks are available
\item additional benchmarks are taken from the second quintile, to produce N*60\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are taken from the third quintile, to produce N*75\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are taken from the fourth quintile, to produce N*90\% benchmarks, if sufficient benchmarks are available
\item any remaining needed benchmarks are taken from the bottom quintile.
\end{itemize}
Within each quintile selection,
\begin{itemize}
\item if the industrial benchmarks comprise less than 85\% of the benchmarks in that quintile, 85\% will be chosen from the industrial benchmarks (if a sufficient number are available), and the others from the population of random and crafted benchmarks;
\item if the industrial benchmarks comprise at least 85\% of the benchmarks in that quintile, the benchmarks are chosen at random from the population of industrial, random and crafted benchmarks in that quintile.
\end{itemize}

Since the organizers at this point are unsure how long a set of benchmarks may take (which will depend also on the number of solvers submitted), the competition will be run in {\em heats}. For each division, some number $N$ of benchmarks will be selected as described and randomly divided into a number of heats. The heats will be run in order. If the competition ends before all heats have completed, the remaining heats will be ignored. All benchmarks will be chosen at the beginning (that is, without reference to the results of any given heat).

The main purpose of the algorithm above is to have a balanced and complete set
of benchmarks.  The  built-in bias is towards industrial rather than crafted
or random benchmarks and towards more difficult benchmarks.
  This reflects a desire by the organizers and agreed upon
by the SMT community to emphasize problems that come from real applications.

Pseudo-random numbers will be generated using the standard C library
function \texttt{random()}, seeded (using \texttt{srandom()}) with the
sum, modulo $2^{30}$, of the numbers provided in the system
descriptions (see Section~\ref{sec:entrants} above) by all SMT-COMP
entrants other than the organizers.  Additionally, the integer part of
the opening value of the New York Stock Exchange Composite Index on
the first day the exchange is open on or after June 2
will be added to the other seeding values.  This helps provide transparency,
by guaranteeing that the organizers cannot manipulate the seed in
favor of or against any particular submitted solver.  Benchmarks will also be slightly
scrambled before the competition, using a simple benchmark scrambler
seeded with the same seed as the benchmark selector.  Both the
scrambler and benchmark selector will be publicly available before the
competition.  Naturally, solvers must not rely on previously
determined identifying syntactic characteristics of competition
benchmarks in testing satisfiability (violation of this is considered
cheating).

\subsection{Application track}

\header{Benchmark sources.} %
Benchmarks for the application track will be collected by the SMT-COMP organizers.
Any benchmark available to the organizers by May 15 will be considered eligible.

\header{Benchmark availability.} 
A first release of the application track
benchmarks will be made available May 15, 2014.
No additional
benchmarks will be added after this date, but benchmarks can be
modified or removed to fix possible bugs or other issues. 
The final selection that will be used for the competition will be posted when the competition begins.

\header{Benchmark demographics and selection.} 
A random selection of all the available benchmarks will be used for the competition. 
As was the case in 2012, no difficulty will be assigned to the benchmarks for the application track.
Benchmarks will be slightly scrambled before the competition, using the same scrambler and 
random seed as the main track. 
If the organizers determine that there is adequate time to complete the competition, all of the benchmarks will be used. Otherwise an unbiased random subset will be used to ensure timely completion of the competition.

\section{Judging and Scoring}
\label{sec:judging}

\subsection{Scoring}

The score for each benchmark is a triple $\langle e,n,m\rangle$, with
$e$ a non-negative number of erroneous results (usually 0, possibly 1 for main track benchmarks, and possibly more than one for application benchmarks),
$n\in[0,N]$ an integral number of points scored for the benchmark,
where $N$ is the number of \akey{check-sat} commands
in the benchmark, and $m\in[0,T]$ is the (real-valued) CPU time in seconds, where $T$ is
the timeout.  Recall that main track benchmarks will have just one \akey{check-sat} command;
application track benchmarks may have multiple \akey{check-sat} commands.

\header{Main track - sequential mode.}
For the main track, the score on a given benchmark is the single $\langle e,n,m\rangle$ triple, with $m$ being the CPU time it takes to process the benchmark, that is, until the solver process terminates, and
\begin{itemize}
\item $e=0, n=1$ if the result of the check-sat command is correct,
\item $e=0, n=0$ if the result is {\tt unknown}, and
\item $e=1, n=0$ if the result is incorrect.
\end{itemize}

\header{Main track - parallel mode.}
The Star-Exec infrastructure reports both wall-clock time and CPU time for a given benchmark-solver job-pair. Each Star-Exec node has multiple (currently 4) cores; a multi-threaded solver can in principle make use of these cores to reduce its wall-clock time. Since the competition has not to date emphasized parallel solvers, we will use total CPU time (across all cores and subprocesses) as the measure of performance in the sequential mode of the main track; the imposed time out will also be on the CPU time.

However, we can also measure the performance using wall-clock time, the key performance criterion for a parallel solver.
We will use the same competition data to determine a parallel-solver winner as follows. The competition will be run with a timeout T for both CPU time and wall-clock time. Since the computation nodes have C=4 cores, we can simulate a competition that had a time out of T CPU seconds and T/C wall-clock seconds by simply declaring that any solver taking more than T/C wall-clock time has timed out for the parallel competition. We then recompute the scoring as above but using wall-clock time and a timeout value of T/C.

\header{Application track.}
An application benchmark may have multiple check-sat commands and may partially solve the benchmark before timing out.
The benchmark is run by the benchmark executor, measuring the total time (summed over all individual commands) taken by
the solver to respond to commands. It is expected that nearly all of this time will be in response to a check-sat command, but assert, push or pop commands might also entail a reasonable amount of processing.
\begin{itemize}
\item If the solver completes the benchmark without timing out, the solver receives a score of $\langle e,n,m\rangle$,
where $e$ is the number of erroneous responses to commands, $n$ is the number of correct responses to check-sat commands, and $m$ is the CPU time taken.
\item If the solver times out while processing the benchmark, the solver receives a score of $\langle e,n,m\rangle$,
where $e$ is the number of erroneous responses to commands, $n$ is the number of correct responses to check-sat commands before timing out, and $m$ is equal to the timeout value.
\end{itemize}

As queries are only presented in order, this scoring system may mean
that relatively ``easier'' queries are hidden behind more difficult
ones located at the middle of the query sequence. We will not perform a parallel performance analysis of the application track.

\header{Competition scoring.}
Benchmarks' scores are summed component-wise to form a solver's total
score for that division of the competition.
Total scores are compared lexicographically---a score $\langle e,n,m\rangle$ is better than 
$\langle e',n',m'\rangle$ iff $e < e'$ or ($e = e'$ and $n > n'$) or ($e = e'$ and $n = n'$ and $m < m'$).
That is, fewer errors takes precedence over more correct solutions, which takes precedence over less time taken.

\subsection{Other recognitions}
The organizers will also recognize the following contributions:
\begin{itemize}
\item {\em Open source}. In addition to recognizing the overall
winner in each division, the top solver that provides its source code as open source will also be
recognized in each division.  
\item {\em Best new entrant}. The best performing entrant from a new solver implementation team, as measured by the average of 
(solver time)/(winning solver time) over the competition benchmarks.
\item The organizers reserve the right to recognize other outstanding contributions that become apparent in the competition results.
\end{itemize}

\subsection{FLoC Olympic Games medals}

The FLoC Olympics games medals will be awarded based on competition-wide critera. The organizers are still determining what those criteria should be. Comments are welcome.

%% TBD: DD suggestion: We could have a system similar to that of the Diamond's league. For each problem division with at least three entrants, attribute 4, 2 and 1 points to the first three solvers. The medals will be awarded based on the total number of points.

\section{Timeline}
\label{sec:timeline}

\nobreak
\vbox{% no page break here, please!
\begin{description}
\item[May 15] Deadline for benchmark contributions.
  First version of the benchmark scrambler, benchmark selector and trace executor made available.
\item[May 25] First versions of solver competitors are due to the organizers.
\item[June 1] Final version of the benchmark scrambler, benchmark selector and trace executor made available.
\item[June 1] Benchmark libraries are frozen; final versions of solver competitors due.
\item[June 1] Final versions of solvers due via StarExec (for
  all tracks), including divisions being entered, system
  descriptions and magic numbers for benchmark scrambling.
\item[June 2] Opening value of NYSE Composite Index used to complete random seed.
\item[July 17] SMT 2014; end of competition.
\end{description}}

\section{Mailing List}
\label{sec:ps}

Interested parties should subscribe to the SMT-COMP mailing list (at \url{cs.nyu.edu/mailman/listinfo/smt-comp}).
Important
late-breaking news and any necessary clarifications and edits to these
rules will be announced there, and it is the primary way that such
announcements will be communicated.

\section{Disclaimer}
\begin{itemize}
\item David Cok, David Deharbe and Tjark Weber are the organizers of SMT-COMP 2014. 
David Cok is the lead and is 
responsible for policy and procedure decisions, such as these
rules, with input from the co-organizers. He is not associated 
with any group creating or submitting solvers. He has used solvers
in industrial settings and is keenly interested to know which are the best.

\item David Deharbe is associated with the solver group producing the veriT solver.
\end{itemize}
\bibliographystyle{plain}
\bibliography{biblio}

\end{document}