### 2015-02-17 rules/rules15.tex
First draft of 2015 rules added.

\documentclass[12pt]{article}

\usepackage{color}
\usepackage{times}
\usepackage{fullpage}
\usepackage{hyperref}

\usepackage{draftwatermark}

\newcommand{\akey}[1]{\textbf{#1}}

\newcommand{\remark}[1]{{\small\textcolor{red}{[#1]}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\date{\small This version revised \the\year-\the\month-\the\day}

\title{10th International Satisfiability Modulo Theories Competition
  (SMT-COMP 2015): Rules and Procedures}

% [morgan] do our own layout of authors; the four-author layout spacing
% was screwed up...
\def\doauthor#1{{%
  \hsize.5\hsize \advance\hsize by-1cm %
  \def\\{\hss\egroup\hbox to\hsize\bgroup\strut\hss}%
  \vbox{\hbox to\hsize\bgroup\strut\hss#1\hss\egroup}}}%

\def\header#1{\medskip\noindent\textbf{#1}}

\author{%
Sylvain Conchon \\
Paris-Sud University \\
France \\
{\small\href{mailto:Sylvain.Conchon@lri.fr}{\tt Sylvain.Conchon@lri.fr}} \\
\and
David D{\'e}harbe \\
Federal University of Rio Grande do Norte \\
Brazil \\
{\small\href{mailto:David.Deharbe@loria.fr}{\tt David.Deharbe@loria.fr}} \\
\and
Tjark Weber \\
Uppsala University \\
Sweden \\
{\small\href{mailto:tjark.weber@it.uu.se}{\tt tjark.weber@it.uu.se}} \\
\\
}

\maketitle

\noindent Comments on this document should be emailed to the SMT-COMP
mailing list (see below) or, if necessary, directly to the organizers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Communication}
\label{sec:communication}

Interested parties should subscribe to the SMT-COMP mailing list.
Important late-breaking news and any necessary clarifications and
edits to these rules will be announced there, and it is the primary
way that such announcements will be communicated.

\begin{itemize}
\item SMT-COMP mailing list: \href{mailto:smt-comp@cs.nyu.edu}{\tt
  smt-comp@cs.nyu.edu}
\item Sign-up site for the mailing list:
  \url{cs.nyu.edu/mailman/listinfo/smt-comp}
\end{itemize}

\noindent Additional material will be made available at the
competition web site, \url{www.smtcomp.org} or
\url{smtcomp.sourceforge.net/2015}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\section{Important Dates}
\label{sec:important}

\begin{description}
\item[May~1] Deadline for new benchmark contributions.
\item[June~1] Deadline for first versions of solvers (for all tracks),
  including information about which tracks and divisions are being
  entered, and magic numbers for benchmark scrambling.
\item[June~7] Final versions of competition tools (e.g., benchmark
  scrambler) are made available.  Benchmark libraries are frozen.
\item[June~14] Deadline for final versions of solvers, including
  system descriptions.
\item[June~15] Opening value of NYSE Composite Index used to complete
  random seed for benchmark scrambling.
\item[July~18/19] SMT Workshop; end of competition, presentation of
  results.
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{leberre+03,PSS02}.  More information
on the history and motivation for SMT-COMP can be found at the
competition web site, \url{www.smtcomp.org}, and in reports on
previous
competitions~(\cite{SMTCOMP2012,SMTEVAL2013,SMTCOMP-JAR,SMTCOMP-FMSD,SMTCOMP-2008,BDOS08}).

SMT-COMP~2015 is affiliated with the SMT Workshop~2015
(\url{http://smt2015.csl.sri.com/}), which is associated with CAV~2015
(\url{http://i-cav.org/2015/}).  The SMT Workshop and the main CAV
conference will include a block of time to present the competitors and
results of the competition.

Accordingly, researchers are highly encouraged to submit both new
benchmarks and new or improved solvers to raise the level of
competition and advance the state-of-the-art in automated SMT problem
solving.

SMT-COMP 2015 will have two tracks: the conventional main track and an
application (i.e., incremental) track.  Within each track there are
multiple divisions, where each division uses benchmarks from a
specific SMT-LIB logic (or group of logics). \remark{Additional
  tracks, e.g., for unsat-core extraction or proof generation, are
  currently being considered.}  We will recognize winners as measured
by number of benchmarks solved; we will also recognize solvers based
on additional criteria.

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  David D{\'e}harbe, Morgan Deters, Alberto Griggio, Albert Oliveras,
  Aaron Stump, and Tjark Weber.}  describes the rules and competition
procedures for SMT-COMP~2015.  The principal changes from the previous
competition rules are the following:
\begin{itemize}
\item All solvers will be run with $n=4$ cores available, and the CPU
  timeout will be $n$~times the wall-clock timeout.  We will recognize
  both best sequential and best parallel performance in all divisions.
\item In addition to recognizing the best solver in each division, we
  will recognize the three solvers that perform best according to
  competition-wide criteria, emphasizing breadth of supported logics.
  These criteria will be a refinement of the criteria used for the
  FLoC Olympic Games medals in~2014.
\item \remark{Pending availability of suitable benchmarks and solvers,
  we may include a new division for floating-point arithmetic.  This
  should be considered experimental in 2015, and will not be included
  in the competition-wide ranking.}
\item \remark{Benchmark selection and relative weight of benchmark
  families are under review.}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Entrants}
\label{sec:entrants}

\header{Solver submission.}
%
An entrant to SMT-COMP is an SMT solver submitted by its authors using
the StarExec (\url{http://www.starexec.org}) service.  The execution
service enables members of the SMT research community to run solvers
on jobs consisting of benchmarks from the SMT-LIB benchmark library.
Jobs are run on a shared computer cluster.  The execution service is
provided free of charge, but it does require registration to create a
login account.  Registered users may then upload their own solvers to
run, or may run public solvers already uploaded to the service.

For participation in SMT-COMP, a solver must be uploaded to StarExec,
made publicly available, and the organizers informed of its presence
{\em and the tracks and divisions in which it is
  participating}. StarExec supports solver configurations; for
clarity, each submitted solver must have one configuration only.  A
submission must also include a 32-bit unsigned integer.  These
numbers, collected from all submissions, are used to seed the
pseudo-random benchmark selection algorithm, as well as the benchmark
scrambler.

Information about how to configure and upload a solver is contained in
the StarExec user guide,
\url{https://wiki.uiowa.edu/display/stardev/User+Guide}.

\header{System description.}
%
As part of a submission, SMT-COMP entrants are encouraged to provide a
short (1--2 pages) description of the system.  This should include a
list of all authors of the system, their present institutional
affiliations, and any appropriate acknowledgements.  The programming
language(s) and basic SMT solving approach employed should be
described (e.g., lazy integration of a Nelson-Oppen combination with
SAT, translation to SAT, etc.).  System descriptions are encouraged to
include a URL for a web site for the submitted tool.  System
descriptions may be submitted after the solver deadline, but to be
useful should be sent to the organizers before the competition ends.

\header{Other systems.}
%
The organizers' intent is to promote as wide a comparison among
solvers and solver options as possible.  However, if the number of
solver submissions is too large for the computational resources
available to the competition, the organizers reserve the right not to
accept multiple versions of solvers from the same solver team.  The
organizers also reserve the right to include other systems of interest
(such as entrants in previous competitions) in the competition.

\header{Wrapper tools.}
%
A \emph{wrapper tool} is defined as any tool which calls one or more
SMT solvers not written by the author of the wrapper tool.  The other
solvers are called the \emph{wrapped} tools.  A wrapper tool must
explicitly acknowledge any tools that it wraps.  Its system
description should make clear the technical innovations by which the
wrapper tool expects to improve on the wrapped tools.

\header{Attendance.}
%
Submitters of an SMT-COMP entrant need not be physically present at
the competition to participate or win.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Deadlines}

SMT-COMP entrants must be submitted via StarExec (solvers) \emph{and}
email to the organizers (accompanying information) until the end of
{\bf June~1, 2015} anywhere on earth.  After this date no new entrants
will be accepted.  However, updates to existing entrants on StarExec
will be accepted until the end of {\bf June~14, 2015} anywhere on
earth.

We strongly encourage participants to use this grace period
\emph{only} for the purpose of fixing any bugs that may be discovered,
and not for adding new features, as there may be no opportunity to do
extensive testing using StarExec after the initial deadline.

The solver versions that are present on StarExec at the conclusion of
the grace period will be the ones used for the competition, and
versions submitted after this time will not be used.  The organizers
reserve the right to start the competition itself at any time after
the open of the New York Stock Exchange on the day after the final
solver deadline.

These deadlines and procedures apply equally to all tracks of the
competition.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Execution of Solvers}
\label{sec:exec}

Solvers will be publicly evaluated in all tracks and divisions into
which they have been entered.  All results of the competition will be
made public.

\subsection{Logistics}

\header{Dates of competition.}
%
The bulk of the computation will take place during the weeks leading
up to SMT~2015, from about June~15 to July~17.  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

The organizers reserve the right to prioritize certain competition
tracks or divisions to ensure their timely completion, and in
exceptional circumstances to complete divisions after the SMT
Workshop.

\header{Input and output.}
%
In the main track, a participating solver must read a single benchmark
script, whose name is presented as the solver's first argument.  In
the application track, a trace executor will send commands from a
benchmark script to the solver's standard input channel.

The benchmark script is in the concrete syntax of the SMT-LIB format,
version~2.0, though with a restricted set of commands.  A benchmark
script is a text file containing a sequence of SMT-LIB~v2 commands in
which:
%
\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
  logic is the first command after any $\akey{set-option}$ commands
  described below.
\item In the main track, there may be a (single) \akey{set-option
  :print-success ...} command. Note that {\tt success} outputs are
  ignored by the post-processor used by the competition; in the
  application track, the :print-success option may not be disabled.
\item The application track executor will send an initial
  $\akey{set-option :print-success true}$ command to the solver.
\item The script next optionally contains (any number of)
  $\akey{set-info}$ commands, which may be ignored by the
  solver. Benchmarks are expected to have the following
  \akey{set-info} commands: \akey{(set-info :smt-lib-version 2.0)},
  \akey{(set-info :category ...)}, \akey{(set-info :source ...)}
  (indicating the authorship or tool that produced the benchmark), and
  \akey{(set-info :status ...)}.
\item The $\akey{exit}$ command is the last command.
\item For tracks other than the application track, there is exactly
  one $\akey{check-sat}$ command, following possibly several
  $\akey{assert}$ commands.
\item For the application track, there are one or more
  \akey{check-sat} commands, each preceded by one or more
  \akey{assert} commands; there may also be zero or more \akey{push 1}
  commands, and zero or more \akey{pop 1} commands, consistent with
  the uses of those commands in the SMT-LIB standard. Each
  \akey{check-sat} command may be preceded by a \akey{set-info
    :status} command that indicates the expected result of the
  \akey{check-sat} command.
\item The formulas in the script belong to the benchmark's logic, with
  any free symbols declared in the script.
\item All sorts declared with a $\akey{declare-sort}$ command must
  have zero arity.
\item Application track scripts may have $\akey{define-sort}$
  commands.
\item No other commands besides the ones just mentioned may be used.
\item Named formulas are not used in benchmark scripts.
\end{enumerate}
%
The SMT-LIB format specification is publicly available from the
``Documents'' section of the SMT-LIB website~\cite{SMT-LIB}.  Solvers
will be given formulas only from the divisions into which they have
been entered.

\header{Timeouts.}
%
Each SMT-COMP solver will be executed on a dedicated processor with
4~cores of a competition machine for each given benchmark, up to a
fixed time limit.  Detailed machine specifications are available on
the competition web site.  The time limit is yet to be determined, but
it is anticipated to be 40~minutes of wall-clock time (and 160~minutes
of CPU time).\footnote{The time limit may be adjusted once we know the
  number of competition entrants and eligible benchmarks.}  Solvers
that take more than this time limit will be killed.  Solvers are
allowed to spawn other processes; these will be killed at
approximately the same time as the first started process.

The StarExec service also limits the memory consumption of the solver
processes.  We expect the limit per solver-benchmark job-pair to be on
the order of 60\,GB.  The values of both the timeout limit and the
memory limit are available to a solver process through environment
variables defined and set by the StarExec service (see the StarExec
user guide for more information).

\subsection{Main track}
\label{sec:exec:main}

The main track competition will consist of selected benchmarks in each
of the logic divisions.  Each benchmark script will be presented to
the solver as its first command-line argument.  Each SMT-COMP entrant
is then expected to attempt to report on its standard output channel
whether the formula is satisfiable (\texttt{sat}, in lowercase) or
unsatisfiable (\texttt{unsat}).  An entrant may also report
\texttt{unknown} to indicate that it cannot determine satisfiability
of the formula.

The main track competition uses a StarExec postprocessor (named ``SMT2
results conserv") to accumulate the results.

\header{Aborts and unparsable output.}
%
Solvers which exit before the time limit without reporting a result
(i.e., due to exhausting memory or crashing) and do not produce output
that includes \texttt{sat}, \texttt{unsat} or \texttt{unknown} will be
considered to have aborted.  Also, as a further measure to prevent
misjudgments of solvers, any \texttt{success} outputs will be
ignored.\footnote{ Note that SMT-LIB~v2 requires that a solver produce
  a \texttt{success} answer after each \akey{set-logic},
  \akey{declare-sort}, \akey{declare-fun} and \akey{assert} command
  (among others), unless the option \akey{:print-success} is set to
  false.  Ignoring the \texttt{success} outputs therefore allows for
  submitting fully-compliant solvers without the need of a wrapper
  script, while still allowing entrants of previous competitions to
  run without changes.}

\header{Persistent state.}
%
Solvers are allowed to create and write to files and directories
during the course of an execution, but they are not allowed to read
such files back during later executions.  Any files written should be
put in the directory in which the tool is started, or in a
subdirectory.

The solver is executed with a temporary directory as the current
working directory.  Any generated files should be produced there (and
not, say, in the system's {\tt /tmp} directory).  The StarExec system
sets a limit on the amount of disk storage permitted---typically
20\,GB (see the User Guide).  The temporary directory and its contents
are deleted after the job is complete.

\subsection{Application track}
\label{sec:exec:application}

The application track evaluates SMT solvers when interacting with an
external verification framework, e.g., a model checker. This
interaction, ideally, happens by means of an online communication
between the model checker and the solver: the model checker repeatedly
sends queries to the SMT solver, which in turn answers either
\texttt{sat} or \texttt{unsat}.  In this interaction an SMT-solver is
required to accept queries incrementally via its {\em standard input
  channel}.

In order to facilitate the evaluation of the solvers in this track, we
will set up a ``simulation'' of the aforementioned interaction. In
particular each benchmark in the application track represents a
realistic communication trace, containing multiple \akey{check-sat}
commands (possibly with corresponding \akey{push 1}/\akey{pop 1}
commands), which is parsed by a {\em trace executor}. The trace
executor serves the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to
  the SMT solver (through stdin);
\item it prevents ``look-ahead'' behaviors of SMT solvers;
\item it records time and answers for each call, possibly aborting the
  execution in case of a wrong answer;
\item it guarantees a fair execution for all solvers by abstracting
  from any possible crash, misbehavior, etc. that may happen on the
  model checker side.
\end{itemize}
Note that the executor terminates processing the script upon receiving
an incorrect response.

The disk space and memory limits on the application track are the same
as for the main track.

\header{Input and Output.}
%
Participating solvers will be connected to a trace executor which will
incrementally send commands to the standard input channel of the
solver and read responses from the standard output channel of the
solver.  The commands will be taken from an \textbf{incremental
  benchmark script}, which is an SMT-LIBv2 script which satisfies the
rules for an application script given above.  Note also that the trace
executor will send a single \akey{set-option :print-success true}
command to the solver before sending commands from the incremental
benchmark script.

Solvers must respond to each command sent by the trace executor, with
the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for \akey{set-option},
\akey{set-logic}, \akey{declare-sort}, \akey{declare-fun},
\akey{define-sort}, \akey{define-fun}, \akey{assert}, \akey{push 1},
and \akey{pop 1} commands, and with a \texttt{sat}, \texttt{unsat}, or
\texttt{unknown} for \akey{check-sat} commands. There will be no
\akey{get-unsat-core}, \akey{get-proof}, \akey{get-model},
\akey{get-assignments}, or \akey{get-values} commands in these
benchmarks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Benchmarks and Problem Divisions}
\label{sec:theories}

\subsection{Main track}

\header{Benchmark sources.}
%
Benchmark formulas for each division of the main track will be drawn
from the SMT-LIB library. The deadline for new benchmarks is {\bf May
  1, 2015}. The organizers will be checking and curating these until
{\bf June 7, 2015}; the organizers reserve the option to exclude new
benchmarks if any prove problematic for some reason.  SMT-COMP
attempts to give preference to benchmarks that are ``real-world,'' in
the sense of coming from or having some intended application outside
SMT.

\header{Benchmark availability.}
%
New benchmarks will be made available as soon as possible after the
benchmark submission deadline, as they are checked and curated.  The
set of benchmarks selected for the competition will be published when
the competition begins.

\header{Benchmark demographics.} %
In SMT-LIB, benchmarks are organized according to \emph{families}.  A benchmark
family contains problems that are similar in some significant way.  Typically
they come from the same source or application, or are all output by the same
tool.  \emph{Each top-level subdirectory within a division represents a distinct
family.}
%

Each benchmark in SMT-LIB also has a \emph{category}.  There are four possible
categories:
%
\begin{itemize}
\item \emph{check.} These benchmarks are hand-crafted to test whether
  solvers support specific features of each division.  In particular,
  there are checks for integer completeness (i.e., benchmarks that are
  satisfiable under the reals but not under the integers) and big
  number support (i.e., benchmarks that are likely to fail if integers
  cannot be represented beyond some maximum value, such as
  $(2^{31}-1)$).

\item \emph{industrial.} These benchmarks come from some real
  application and are produced by tools such as bounded model
  checkers, static analyzers, extended static checkers, etc.

\item \emph{random.} These benchmarks are randomly generated.

\item \emph{crafted.} This category is for all other benchmarks.
  Usually, benchmarks in this category are designed to be particularly
  difficult or to test a specific feature of the logic.
\end{itemize}

\header{Benchmark selection.} \remark{This section is under review.}
%
Each benchmark is assigned a difficulty. For 2015, the difficulty
measure is the CPU time taken by the best performing solver in the
2014 SMT-COMP.  For new benchmarks, we will determine an approximate
difficulty using trial runs.  The difficulty values will be posted as
soon after May~1 as possible.  The difficulty values are best
estimates by the organizers and may not always reflect actual
difficulty.

The selection of benchmarks for the competition will be biased towards more difficult benchmarks
as follows. First, the benchmark pool is culled as follows:
\begin{enumerate}

\item \textbf{Retire very easy benchmarks.}
%
Eliminate benchmarks for which all \remark{or perhaps some} of the
current solvers in SMT-COMP 2014 solved the benchmark in under
5~seconds, \emph{unless} doing so reduces the pool of benchmarks for
the division to less than~300.

\item \textbf{Retire inappropriate benchmarks.}
%
The competition organizers will remove from the eligibility pool
certain SMT-LIB benchmarks that are inappropriate or uninteresting for
competition, or cut the size of certain benchmark families to avoid
their over-representation.

\item \textbf{Eliminate benchmarks whose status is unknown.}
%
The selection process ignores any benchmark whose expected output is
neither {\tt sat} nor {\tt unsat}.  There is a significant number of
benchmarks with {\tt unknown} status in the library; these are
certainly interesting and are a challenge to solve, but they are not
used in the competition.

\end{enumerate}

Then, for each division, the benchmarks are divided into quintiles according to the difficulty measure.
Next a random selection of N competition benchmarks is made as follows:
\begin{itemize}
\item N*40\% are randomly selected from the top quintile, if sufficient benchmarks are available
\item additional benchmarks are randomly selected from the second quintile, to produce N*60\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are randomly selected from the third quintile, to produce N*75\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are randomly selected from the fourth quintile, to produce N*90\% benchmarks, if sufficient benchmarks are available
\item any remaining needed benchmarks are randomly selected from the bottom quintile.
\end{itemize}
Within each quintile selection,
\begin{itemize}
\item if the industrial benchmarks comprise less than 85\% of the benchmarks in that quintile, 85\% will be chosen from the industrial benchmarks (if a sufficient number are available), and the others from the population of random and crafted benchmarks;
\item if the industrial benchmarks comprise at least 85\% of the benchmarks in that quintile, the benchmarks are chosen at random from the population of industrial, random and crafted benchmarks in that quintile.
\end{itemize}

Since the organizers at this point are unsure how long a set of
benchmarks may take (which will depend also on the number of solvers
submitted), the competition will be run in {\em heats}. For each
division, some number $N$ of benchmarks will be selected as described
and randomly divided into a number of (possibly unequal-sized)
heats. The heats will be run in order. If the competition ends before
all heats have completed, the remaining heats will be ignored. All
benchmarks will be chosen at the beginning (that is, without reference
to the results of any given heat).

The main purpose of the algorithm above is to have a balanced and
complete set of benchmarks.  The built-in bias is towards industrial
rather than crafted or random benchmarks and towards more difficult
benchmarks.  This reflects a desire by the organizers and agreed upon
by the SMT community to emphasize problems that come from real
applications.

Pseudo-random numbers will be generated using the standard C library
function \texttt{random()}, seeded (using \texttt{srandom()}) with the
sum, modulo $2^{30}$, of the numbers provided in the system
descriptions (see Section~\ref{sec:entrants} above) by all SMT-COMP
entrants other than the organizers.  Additionally, the integer part of
the opening value of the New York Stock Exchange Composite Index on
the first day the exchange is open on or after the date specified in
the timeline will be added to the other seeding values.  This helps
provide transparency, by guaranteeing that the organizers cannot
manipulate the seed in favor of or against any particular submitted
solver.  Benchmarks will also be slightly scrambled before the
competition, using a simple benchmark scrambler seeded with the same
seed as the benchmark selector.  Both the scrambler and benchmark
selector will be publicly available before the competition.
Naturally, solvers must not rely on previously determined identifying
syntactic characteristics of competition benchmarks in testing
satisfiability (violation of this is considered cheating).

\subsection{Application track}

\header{Benchmark sources.}
%
Benchmarks for the application track will be collected by the SMT-COMP
organizers.  Any benchmark available to the organizers by {\bf May~1,
  2015} will be considered eligible.

\header{Benchmark availability.}
A first release of the application track
benchmarks will be made available as soon after the benchmark deadline as the organizers can arrange.
No additional
benchmarks will be added after this date, but benchmarks can be
modified or removed to fix possible bugs or other issues.
The final selection that will be used for the competition will be posted when the competition begins.

\header{Benchmark demographics and selection.}
A random selection of all the available benchmarks will be used for the competition.
As was the case in 2012, no difficulty will be assigned to the benchmarks for the application track.
Benchmarks will be slightly scrambled before the competition, using the same scrambler and
random seed as the main track.
If the organizers determine that there is adequate time to complete the competition, all of the benchmarks will be used. Otherwise an unbiased random subset will be used to ensure timely completion of the competition.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Scoring}
\label{sec:scoring}

Scores will be computed for all solvers and divisions.  However,
winners will be declared only for \emph{competitive} divisions.  A
division is competitive if at least two substantially different
solvers (i.e., solvers from two different teams) were submitted.
Although the organizers may enter other solvers for comparison
purposes, only solvers that are explicitly submitted by their authors
determine whether a division is competitive, and are eligible to be
designated as winners.

\subsection{Benchmark scoring}

A solver's score for each benchmark is a quadruple $\langle e, n, w,
c\rangle$, with $e \in \{0, 1\}$ the number of erroneous results
(usually~$e = 0$), $n \in [0,N]$ an integral number of points scored
for the benchmark, where $N$ is the number of \akey{check-sat}
commands in the benchmark, $c \in [0,4T]$ is the (real-valued) CPU
time in seconds, and $w \in [0,T]$ is the (real-valued) wall-clock
time in seconds, where $T$ is the timeout.  Recall that main track
benchmarks will have just one \akey{check-sat} command; application
track benchmarks may have multiple \akey{check-sat} commands.

\header{Main track.}
%
\remark{Clarify how $\langle e, n, c, w\rangle$ depend on solver
  behavior.}

\header{Application track.}
%
An application benchmark may have multiple check-sat commands and may
partially solve the benchmark before timing out.  The benchmark is run
by the benchmark executor, measuring the total time (summed over all
individual commands) taken by the solver to respond to commands. It is
expected that nearly all of this time will be in response to
\akey{check-sat} commands, but \akey{assert}, \akey{push} or
\akey{pop} commands might also entail a reasonable amount of
processing.
%% \begin{itemize}
%% \item If the solver completes the benchmark without timing out, the
%%   solver receives a score of $\langle e,n,m\rangle$, where $e$ is the
%%   number of erroneous responses to commands, $n$ is the number of
%%   correct responses to check-sat commands, and $m$ is the CPU time
%%   taken.
%% \item If the solver times out while processing the benchmark, the
%%   solver receives a score of $\langle e,n,m\rangle$, where $e$ is the
%%   number of erroneous responses to commands, $n$ is the number of
%%   correct responses to check-sat commands before timing out, and $m$
%%   is equal to the timeout value.
%% \end{itemize}

\subsection{Division scoring}

To compute a solver's score for a division, the solver's individual
benchmark scores for all benchmarks in the division are summed
component-wise.  Total scores are compared lexicographically: a score
$\langle e, n, w, c\rangle$ is better than $\langle e', n', w',
c'\rangle$ iff $e < e'$ or ($e = e'$ and $n > n'$) or ($e = e'$ and $n
= n'$ and $w < w'$) or ($e = e'$ and $n = n'$ and $w = w'$ and $c <
c'$).  That is, fewer errors takes precedence over more correct
solutions, which takes precedence over less wall-clock time taken,
which takes precedence over less CPU time taken.

\subsection{Competition-wide scoring (main track)}

We define a competition-wide metric for the main track as follows.
Let $B_i$ be the total number of benchmarks in division~$i$ that were
used in the competition, and let $\langle e_i, n_i, w_i, c_i\rangle$
be a solver's score for this division.  The solver's competition-wide
score is $\sum_i (e_i == 0 ? (n_i/N_i)^2 : -e_i) \log N_i$, where the
sum is over all \emph{competitive} divisions into which the solver was
entered.  \remark{Rationale.}

%% {\em Rationale.} The metric M purposely emphasizes breadth of solver
%% participation---a solver participating in many logics need not be the
%% best in any one of them.  The use of the square in the metric limits
%% that somewhat---a solver still needs to do reasonably well compared to
%% winners to be able to catch up by breadth of participation.  The
%% non-linear metric also gives added weight to completing close to all
%% benchmarks in a division.  The metric is also quite simple and the
%% metric for a solver is independent of the performance of other
%% solvers.  Time is omitted from the metric both to simplify the metric
%% and because it is only of third importance in the regular competition
%% metric and is difficult to compare across divisions.

\subsection{Sequential performance}

SMT-COMP has traditionally emphasized sequential performance (i.e.,
CPU time) over parallel performance (i.e., wall-clock time).
Accordingly, we will recognize the best \emph{sequential} solver, both
in each division and according to the competition-wide metric (for the
main track).  A solver is eligible for this recognition if, for each
benchmark in the division, the solver's CPU time does not exceed its
wall-clock time.

\subsection{Other recognitions}

The organizers will also recognize the following contributions:
%
\begin{itemize}
\item {\em Open source}. In addition to recognizing the overall winner
  in each division, the top solver that provides its source code as
  open source will also be recognized in each division.
\item {\em Best new entrant}. The best performing entrant from a new
  solver implementation team, as measured by the Olympic Games metric
  defined below.
\item {\em Breadth of logics}. Tools that cover the most theories and
  logics.
\item {\em Benchmarks}. Contributors of new benchmarks.
\end{itemize}
%
The organizers reserve the right to recognize other outstanding
contributions that become apparent in the competition results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Judging}

The organizers reserve the right, with careful deliberation, to remove
a benchmark from the competition results if it is determined that the
benchmark is faulty (e.g., syntactically invalid in a way that affects
some solvers but not others); and to clarify ambiguities in these
rules that are discovered in the course of the competition.  Authors
of solver entrants may appeal to the organizers to request such
decisions.  Organizers that are affiliated with solver entrants will
be recused from these decisions.  The organizers' decisions are final.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Acknowledgments and Disclaimer}

SMT-COMP 2015 is organized under the direction of the SMT Steering
Committee. The organizing team is
%
\begin{itemize}
\setlength{\itemsep}{0pt}
\item \href{https://www.lri.fr/~conchon/}{Sylvain Conchon}~--
  Paris-Sud University, France (co-organizer)
\item \href{https://www.sites.google.com/site/deharbe/}{David
  D{\'e}harbe}~-- Federal University of Rio Grande do Norte, Brazil
  (co-organizer)
\item \href{http://user.it.uu.se/~tjawe125/}{Tjark Weber}~-- Uppsala
  University, Sweden (chair)
\end{itemize}
%
Tjark Weber is responsible for policy and procedure decisions, such as
these rules, with input from the co-organizers. He is not associated
with any group creating or submitting solvers.

Many others have contributed benchmarks, effort, and feedback.
\remark{Specific individuals will be named later.}
% In particular, Clark Barrett processed the many benchmark
% submissions.
The competition uses the \href{https://www.starexec.org/}{StarExec}
service, which is hosted at the
\href{http://www.cs.uiowa.edu/}{University of Iowa}.  Aaron Stump is
providing essential StarExec support.

David D{\'e}harbe is associated with the solver group producing the
veriT solver.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{plain}
\bibliography{biblio}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}