### 2019-03-15 2018/report/table-winners-main-new.tex
New layout for winners table (currently only main).

\begin{table}
\caption{Best main-track solvers (by division)}
\label{tab:winners-main}
\centering
\scalebox{.83}{%
\renewcommand{\arraystretch}{1.01}%
\begin{tabular}{r@{\hskip 3em}c@{\hskip 2em}c@{\hskip 2em}c@{\hskip 2em}c}
\toprule
Division     & 2015            & 2016            & 2017                   & 2018                  \\
\hline \hline
ABVFP        &                 &                 &                        & {CVC4}                \\
ALIA         & {{[}Z3{]}}      & {{[}Z3{]}}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
AUFBVDTLIA   &                 &                 & {CVC4}                 & {CVC4}                \\
AUFDTLIA     &                 &                 & {CVC4}                 & {CVC4}                \\
AUFLIA       & {CVC4}          & {CVC4}          & {CVC4}                 & {CVC4}                \\
AUFLIRA      & {{[}Z3{]}}      & {{[}Z3{]}}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
AUFNIRA      & {CVC4}          & {Vampire}       & {Vampire}              & {CVC4}                \\
BV           & {{[}Z3{]}}      & {Q3B}           & {{[}Z3{]}}             & {CVC4}                \\
BVFP         &                 &                 &                        & {CVC4}                \\
FP           &                 &                 &                        & {CVC4}                \\
LIA          & {CVC4}          & {CVC4}          & {{[}Z3{]}}             & {{[}Z3{]}}            \\
LRA          & {CVC4}          & {CVC4}          & {{[}Z3{]}}             & {{[}Z3{]}}            \\
NIA          & {{[}Z3{]}}      & {{[}Z3{]}}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
NRA          & {CVC4}          & {Vampire}       & {Redlog}               & {{[}Z3{]}/Vampire}    \\
QF\_ABV      & {Boolector}     & {Boolector}     & {Boolector}            & {Boolector}           \\
QF\_ABVFP    &                 &                 & {---}                  & {CVC4}                \\
QF\_ALIA     & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
QF\_ANIA     & {{[}Z3{]}}      & {CVC4}          & {CVC4}                 & {{[}Z3{]}}            \\
QF\_AUFBV    & {{[}MathSAT{]}} & {{[}MathSAT{]}} & {{[}MathSAT{]}}        & {CVC4}                \\
QF\_AUFLIA   & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
QF\_AUFNIA   & {CVC4}          & {CVC4}          & {{[}Z3{]}}             & {{[}Z3{]}}            \\
QF\_AX       & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
QF\_BV       & {Boolector}     & {Boolector}     & {Boolector/MinkeyRink} & {Boolector/MinkeyRink}\\
QF\_BVFP     & {Z3}            & {{[}Z3{]}}      & {{[}Z3{]}}             & {CVC4}                \\
QF\_DT       &                 &                 & {CVC4}                 & {CVC4}                \\
QF\_FP       & {Z3}            & {{[}MathSAT{]}} & {{[}Z3{]}}             & {COLIBRI}             \\
QF\_IDL      & {{[}Z3{]}}      & {{[}Z3{]}}      & {Yices2}               & {Yices2}              \\
QF\_LIA      & {{[}MathSAT{]}} & {{[}MathSAT{]}} & {{[}MathSAT{]}}        & {SPASS-SATT}          \\
QF\_LIRA     & {Yices2}        & {Yices2}        & {{[}Z3{]}}             & {{[}Z3{]}}            \\
QF\_LRA      &  {CVC4}         & {CVC4}          & {CVC4}                 & {CVC4}                \\
QF\_NIA      & {{[}Z3{]}}      & {{[}Z3{]}}      & {CVC4}                 & {CVC4}                \\
QF\_NIRA     & {CVC4}          & {CVC4}          & {SMT-RAT}              & {SMT-RAT}             \\
QF\_NRA      & {{[}Z3{]}}      & {{[}Z3{]}}      & {Yices2}               & {{[}Z3{]}}            \\
QF\_RDL      & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
QF\_SLIA     &                 &                 &                        & {CVC4}                \\
QF\_UF       & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
QF\_UFBV     & {Boolector}     & {Boolector}     & {Boolector}            & {Boolector}           \\
QF\_UFIDL    & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
QF\_UFLIA    & {{[}Z3{]}}      & {{[}Z3{]}}      & {Yices2}               & {Yices2}              \\
QF\_UFLRA    & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
QF\_UFNIA    & {CVC4}          & {Yices2/CVC4}   & {Yices2}               & {Yices2}              \\
QF\_UFNRA    & {{[}Z3{]}}      & {Yices2}        & {{[}Z3{]}}             & {Yices2}              \\
UF           & {CVC4}          & {CVC4}          & {Vampire}              & {CVC4/Vampire}        \\
UFBV         & {{[}Z3{]}}      & {{[}Z3{]}}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
UFDT         &                 &                 & {CVC4}                 & {CVC4}                \\
UFDTLIA      &                 &                 & {Vampire}              & {CVC4}                \\
UFIDL        & {{[}Z3{]}}      & {CVC4}          & {CVC4}                 & {{[}Z3{]}}            \\
UFLIA        & {CVC4}          & {CVC4}          & {CVC4}                 & {CVC4}                \\
UFLRA        & {CVC3}          & {{[}Z3{]}}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
UFNIA        & {CVC4}          & {Vampire}       & {Vampire}              & {{[}Z3{]}/Vampire}    \\
\bottomrule
\end{tabular}}
\end{table}