### 2019-03-16 2018/report/table-winners-main-new.tex
Beautify layout of main track winners table.

\begin{table}
\caption{Best main-track solvers (by division)}
\label{tab:winners-main}
\centering
\scalebox{.83}{%
\renewcommand{\arraystretch}{1.01}%
  \begin{tabular}{r@{\hskip 3em}>{\columncolor[gray]{.95}[.5em][1em]}c@{\hskip 2em}>{\columncolor[gray]{.95}[1em][1em]}c@{\hskip 2em}>{\columncolor[gray]{.95}[1em][1em]}c@{\hskip 2em}>{\columncolor[gray]{.95}[1em][0.5em]}c}
\toprule
\rc
Division         &  2015           &  2016           &  2017                  &  2018                 \\
\hline \hline
\cc ABVFP        & \cc             & \cc             & \cc                    & \cc {CVC4}            \\
\cc ALIA         & {{[}Z3{]}}      & {{[}Z3{]}}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
\drc
\cc AUFBVDTLIA   & \cc             & \cc             & {CVC4}                 & {CVC4}                \\
\cc AUFDTLIA     & \cc             & \cc             & {CVC4}                 & {CVC4}                \\
\drc
\cc AUFLIA       & {CVC4}          & {CVC4}          & {CVC4}                 & {CVC4}                \\
\cc AUFLIRA      & {{[}Z3{]}}      & {{[}Z3{]}}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
\drc
\cc AUFNIRA      & \cc {CVC4}      & {Vampire}       & {Vampire}              &\cc  {CVC4}            \\
\rc
\cc BV           &  {{[}Z3{]}}     & {Q3B}           & {{[}Z3{]}}             &  {CVC4}               \\
\cc BVFP         & \cc             & \cc             & \cc                    & \cc {CVC4}            \\
\cc FP           & \cc             & \cc             & \cc                    & \cc {CVC4}            \\
\cc LIA          & \dcc {CVC4}     & \dcc{CVC4}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
\cc LRA          & {CVC4}          & {CVC4}          & \dcc {{[}Z3{]}}        & \dcc {{[}Z3{]}}       \\
\cc NIA          & {{[}Z3{]}}      & {{[}Z3{]}}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
\rc
\cc NRA          & {CVC4}          & {Vampire}       & {Redlog}               & {{[}Z3{]}/Vampire}    \\
\cc QF\_ABV      & {Boolector}     & {Boolector}     & {Boolector}            & {Boolector}           \\
\cc QF\_ABVFP    & \cc             &  \cc            & \cc {---}              & \cc {CVC4}            \\
\cc QF\_ALIA     & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\drc
\cc QF\_ANIA     & \cc {{[}Z3{]}}  & {CVC4}          & {CVC4}                 & {{[}Z3{]}}            \\
\cc QF\_AUFBV    & {{[}MathSAT{]}} & {{[}MathSAT{]}} & {{[}MathSAT{]}}        & \cc {CVC4}            \\
\drc
\cc QF\_AUFLIA   & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\cc QF\_AUFNIA   & {CVC4}          & {CVC4}          & \dcc {{[}Z3{]}}        & \dcc {{[}Z3{]}}       \\
\cc QF\_AX       & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\drc
\cc QF\_BV       & {Boolector}     & {Boolector}     & {Boolector/MinkeyRink} & {Boolector/MinkeyRink}\\
\cc QF\_BVFP     & {Z3}            & {{[}Z3{]}}      & {{[}Z3{]}}             & \cc {CVC4}            \\
\drc
\cc QF\_DT       & \cc             & \cc             & {CVC4}                 & {CVC4}                \\
\rc
\cc QF\_FP       & {Z3}            & {{[}MathSAT{]}} & {{[}Z3{]}}             & {COLIBRI}             \\
\cc QF\_IDL      & \dcc {{[}Z3{]}} & \dcc {{[}Z3{]}} & {Yices2}               & {Yices2}              \\
\cc QF\_LIA      & {{[}MathSAT{]}} & {{[}MathSAT{]}} & {{[}MathSAT{]}}        & \cc {SPASS-SATT}      \\
\cc QF\_LIRA     & {Yices2}        & {Yices2}        & \dcc {{[}Z3{]}}        & \dcc {{[}Z3{]}}       \\
\cc QF\_LRA      &  {CVC4}         & {CVC4}          & {CVC4}                 & {CVC4}                \\
\cc QF\_NIA      & {{[}Z3{]}}      & {{[}Z3{]}}      & {CVC4}                 & {CVC4}                \\
\cc QF\_NIRA     & {CVC4}          & {CVC4}          & {SMT-RAT}              & {SMT-RAT}             \\
\cc QF\_NRA      & \dcc {{[}Z3{]}} & \dcc {{[}Z3{]}}      & \cc {Yices2}      & \cc {{[}Z3{]}}        \\
\cc QF\_RDL      & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\cc QF\_SLIA     & \cc             & \cc             & \cc                    & \cc {CVC4}            \\
\cc QF\_UF       & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\drc
\cc QF\_UFBV     & {Boolector}     & {Boolector}     & {Boolector}            & {Boolector}           \\
\cc QF\_UFIDL    & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\cc QF\_UFLIA    & \dcc {{[}Z3{]}} & \dcc {{[}Z3{]}} & {Yices2}               & {Yices2}              \\
\cc QF\_UFLRA    & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\drc
\cc QF\_UFNIA    & \cc {CVC4}      & {Yices2/CVC4}   & {Yices2}               & {Yices2}              \\
\rc
\cc QF\_UFNRA    & {{[}Z3{]}}      & {Yices2}        & {{[}Z3{]}}             & {Yices2}              \\
\cc UF           & {CVC4}          & {CVC4}          & \dcc {Vampire}         & \dcc {CVC4/Vampire}   \\
\cc UFBV         & {{[}Z3{]}}      & {{[}Z3{]}}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
\cc UFDT         & \cc             & \cc             & {CVC4}                 & {CVC4}                \\
\rc
\cc UFDTLIA      &                 &                 & {Vampire}              & {CVC4}                \\
\cc UFIDL        & \cc {{[}Z3{]}}  & {CVC4}          & {CVC4}                 & \cc {{[}Z3{]}}        \\
\drc
\cc UFLIA        & {CVC4}          & {CVC4}          & {CVC4}                 & {CVC4}                \\
\cc UFLRA        & \cc {CVC3}      & {{[}Z3{]}}      & {{[}Z3{]}}             & {{[}Z3{]}}            \\
\drc
\cc UFNIA        & \cc {CVC4}      & {Vampire}       & {Vampire}              & {{[}Z3{]}/Vampire}    \\
\bottomrule
\end{tabular}}
\end{table}