### 2019-03-16 2018/report/table-winners-main.tex
Removed old main track winners table.

\begin{table}
\caption{Best main-track solvers (by division)}
\label{tab:winners-main}
\centering
\scalebox{.83}{%
\renewcommand{\arraystretch}{1.01}%
  \begin{tabular}{r@{\hskip 1em}>{\columncolor[gray]{.95}[.25em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][0.25em]}c}
\toprule
\rc
Division         &  2015           &  2016           &  2017                  &  2018                 \\
\hline \hline
\cc ABVFP        & \cc             & \cc             & \cc                    & \cc {CVC4} *          \\
\cc ALIA         & CVC4 {{[}Z3{]}}      & CVC4 {{[}Z3{]}}      & CVC4 {{[}Z3{]}}             & CVC4 {{[}Z3{]}}            \\
\cc AUFBVDTLIA   & \cc             & \cc             & {CVC4} *               & {CVC4} *              \\
\cc AUFDTLIA     & \cc             & \cc             & {CVC4} *               & {CVC4}                \\
\cc AUFLIA       & {CVC4}          & {CVC4}          & {CVC4}                 & {CVC4}                \\
\drc
\cc AUFLIRA      & \cc CVC4 {{[}Z3{]}}      & Vampire {{[}Z3{]}}      & Vampire {{[}Z3{]}}             & \cc CVC4 {{[}Z3{]}}            \\
\drc
\cc AUFNIRA      & \cc {CVC4} *    & {Vampire}       & {Vampire}              &\cc  {CVC4}            \\
\cc BV           & \cc CVC4 {{[}Z3{]}} *    & {Q3B}           & Q3B {{[}Z3{]}}             &  \cc {CVC4}               \\
\cc BVFP         & \cc             & \cc             & \cc                    & \cc {CVC4} *          \\
\cc FP           & \cc             & \cc             & \cc                    & \cc {CVC4} *          \\
\cc LIA          & {CVC4}     & {CVC4}      & CVC4 {{[}Z3{]}}             & CVC4 {{[}Z3{]}}            \\
\cc LRA          & {CVC4}     & {CVC4}     & CVC4 {{[}Z3{]}}             & CVC4 {{[}Z3{]}}            \\
\cc NIA          & \cc CVC4 {{[}Z3{]}} *     & \cc ProB {{[}Z3{]}}      & CVC4 {{[}Z3{]}}             & CVC4 {{[}Z3{]}}            \\
\rc
\cc NRA          & {CVC4} *        & {Vampire}       & {Redlog}               & Vampire {{[}Z3{]}/Vampire}    \\
\cc QF\_ABV      & {Boolector}     & {Boolector}     & {Boolector}            & {Boolector}           \\
\cc QF\_ABVFP    & \cc             &  \cc            & \cc {---}              & \cc {CVC4}            \\
\cc QF\_ALIA     & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\drc
\cc QF\_ANIA     & CVC4 {{[}Z3{]}} *& {CVC4} *        & {CVC4} *               & CVC4 {{[}Z3{]}} *           \\
\cc QF\_AUFBV    & CVC4 {{[}MathSAT{]}} & CVC4 {{[}MathSAT{]}} & Yices2 {{[}MathSAT{]}}        & \cc {CVC4}            \\
\drc
\cc QF\_AUFLIA   & {Yices2}        & {Yices2}        & {Yices2}                 & {Yices2}              \\
\cc QF\_AUFNIA   & {CVC4} *        & {CVC4} *        & CVC4 {{[}Z3{]}} *        & CVC4 {{[}Z3{]}} *     \\
\cc QF\_AX       & {Yices2}        & {Yices2}        & {Yices2}                 & {Yices2}              \\
\drc
\cc QF\_BV       & {Boolector}     & {Boolector}     & {Boolector/MinkeyRink}   & {Boolector/MinkeyRink}\\
\cc QF\_BVFP     & {Z3} *          & {{[}Z3{]}} *    & \cc COLIBRI {{[}Z3{]}} * & \cc {CVC4}            \\
\drc
\cc QF\_DT       & \cc             & \cc             & {CVC4} *               & {CVC4} *              \\
\cc QF\_FP       & \cc {Z3} *      & \cc {{[}MathSAT{]}} * & COLIBRI {{[}Z3{]}}             & {COLIBRI}             \\
\drc
\cc QF\_IDL      & Yices2 {{[}Z3{]}} & Yices2 {{[}Z3{]}} & {Yices2}               & {Yices2}              \\
\cc QF\_LIA      & CVC4 {{[}MathSAT{]}} & CVC4 {{[}MathSAT{]}} & CVC4 {{[}MathSAT{]}}        & \cc {SPASS-SATT}      \\
\drc
\cc QF\_LIRA     & {Yices2}        & {Yices2}        & Yices2 {{[}Z3{]}}        & Yices2 {{[}Z3{]}}       \\
\cc QF\_LRA      &  {CVC4}         & {CVC4}          & {CVC4}                 & {CVC4}                \\
\cc QF\_NIA      & \cc AProVE {{[}Z3{]}} & \cc Yices2 {{[}Z3{]}} & {CVC4}                 & {CVC4}                \\
\cc QF\_NIRA     & {CVC4}          & {CVC4}          & \dcc {SMT-RAT}         & \dcc {SMT-RAT}        \\
\cc QF\_NRA      & Yices2 {{[}Z3{]}} & Yices2 {{[}Z3{]}}      & {Yices2}      & Yices2 {{[}Z3{]}}        \\
\cc QF\_RDL      & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\cc QF\_SLIA     & \cc             & \cc             & \cc                    & \cc {CVC4} *          \\
\cc QF\_UF       & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\drc
\cc QF\_UFBV     & {Boolector}     & {Boolector}     & {Boolector}            & {Boolector}           \\
\cc QF\_UFIDL    & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\cc QF\_UFLIA    & Yices2 {{[}Z3{]}} & Yices2 {{[}Z3{]}} & {Yices2}               & {Yices2}              \\
\cc QF\_UFLRA    & {Yices2}        & {Yices2}        & {Yices2}               & {Yices2}              \\
\cc QF\_UFNIA    & \cc {CVC4} *    & {Yices2/CVC4}   & {Yices2}               & {Yices2}              \\
\cc QF\_UFNRA    & \cc CVC3 {{[}Z3{]}} *     & {Yices2}        & Yices2 {{[}Z3{]}}             & {Yices2}              \\
\cc UF           & {CVC4}          & {CVC4}          & \dcc {Vampire}         & \dcc {CVC4/Vampire}   \\
\drc
\cc UFBV         & CVC4 {{[}Z3{]}} *    & CVC4 {{[}Z3{]}}      & CVC4 {{[}Z3{]}} *            & CVC4 {{[}Z3{]}} *           \\
\cc UFDT         & \cc             & \cc             & {CVC4} *               & {CVC4}                \\
\rc
\cc UFDTLIA      &                 &                 & {Vampire} *            & {CVC4}                \\
\cc UFIDL        & CVC4 {{[}Z3{]}}  & {CVC4}          & {CVC4}                 & CVC4 {{[}Z3{]}}        \\
\cc UFLIA        & {CVC4}          & {CVC4}          & {CVC4}                 & {CVC4}                \\
\cc UFLRA        & \cc {CVC3}      & \cc Vampire {{[}Z3{]}}      & CVC4 {{[}Z3{]}}             & CVC4 {{[}Z3{]}}            \\
\cc UFNIA        & \cc {CVC4} *    & {Vampire}       & {Vampire}              & Vampire {{[}Z3{]}/Vampire}    \\
\bottomrule
\end{tabular}}
\end{table}