### 2019-03-19 2018/report/table-winners-main.tex
New layout for application winners table.

\begin{table}
\caption{Best main-track solvers (by division)}
\label{tab:winners-main}
\centering
\scalebox{.83}{%
\renewcommand{\arraystretch}{1.01}%
  \begin{tabular}{r@{\hskip 1em}>{\columncolor[gray]{.95}[.25em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][0.5em]}c}
    \toprule
    \rc
    Division         &  2015              &  2016               &  2017                 &  2018                   \\
    \hline \hline
    \cc ABVFP        & \cc                & \cc                 & \cc                   & \cc {CVC4} *            \\
    \cc ALIA         & CVC4 \nc{Z3}       & CVC4 \nc{Z3}        & CVC4 \nc{Z3}          & CVC4 \nc{Z3}            \\
    \cc AUFBVDTLIA   & \cc                & \cc                 & {CVC4} *              & {CVC4} *                \\
    \cc AUFDTLIA     & \cc                & \cc                 & {CVC4} *              & {CVC4}                  \\
    \cc AUFLIA       & {CVC4}             & {CVC4}              & {CVC4}                & {CVC4}                  \\
    \drc
    \cc AUFLIRA      & \cc CVC4 \nc{Z3}   & Vampire \nc{Z3}     & Vampire \nc{Z3}       & \cc CVC4 \nc{Z3}        \\
    \drc
    \cc AUFNIRA      & \cc {CVC4} *       & {Vampire}           & {Vampire}             &\cc  {CVC4}              \\
    \cc BV           & \cc CVC4 \nc{Z3} * & {Q3B}               & Q3B \nc{Z3}           &  \cc {CVC4}             \\
    \cc BVFP         & \cc                & \cc                 & \cc                   & \cc {CVC4} *            \\
    \cc FP           & \cc                & \cc                 & \cc                   & \cc {CVC4} *            \\
    \cc LIA          & {CVC4}             & {CVC4}              & CVC4 \nc{Z3}          & CVC4 \nc{Z3}            \\
    \cc LRA          & {CVC4}             & {CVC4}              & CVC4 \nc{Z3}          & CVC4 \nc{Z3}            \\
    \cc NIA          & \cc CVC4 \nc{Z3} * & \cc ProB \nc{Z3}    & CVC4 \nc{Z3}          & CVC4 \nc{Z3}            \\
    \rc
    \cc NRA          & {CVC4} *           & {Vampire}           & {Redlog}              & Vampire \nc{Z3/Vampire} \\
    \cc QF\_ABV      & {Boolector}        & {Boolector}         & {Boolector}           & {Boolector}             \\
    \cc QF\_ABVFP    & \cc                &  \cc                & \cc {---}             & \cc {CVC4}              \\
    \cc QF\_ALIA     & {Yices2}           & {Yices2}            & {Yices2}              & {Yices2}                \\
    \drc
    \cc QF\_ANIA     & CVC4 \nc{Z3} *     & {CVC4} *            & {CVC4} *              & CVC4 \nc{Z3} *          \\
    \cc QF\_AUFBV    & CVC4 \nc{MathSAT}  & CVC4 \nc{MathSAT}   & Yices2 \nc{MathSAT}   & \cc {CVC4}              \\
    \drc
    \cc QF\_AUFLIA   & {Yices2}           & {Yices2}            & {Yices2}              & {Yices2}                \\
    \cc QF\_AUFNIA   & {CVC4} *           & {CVC4} *            & CVC4 \nc{Z3} *        & CVC4 \nc{Z3} *          \\
    \cc QF\_AX       & {Yices2}           & {Yices2}            & {Yices2}              & {Yices2}                \\
    \drc
    \cc QF\_BV       & {Boolector}        & {Boolector}         & {Boolector/MinkeyRink}& {Boolector/MinkeyRink}  \\
    \cc QF\_BVFP     & {Z3} *             & \nc{Z3} *           & \cc COLIBRI \nc{Z3} * & \cc {CVC4}              \\
    \drc
    \cc QF\_DT       & \cc                & \cc                 & {CVC4} *              & {CVC4} *                \\
    \cc QF\_FP       & \cc {Z3} *         & \cc \nc{MathSAT} *  & COLIBRI \nc{Z3}       & {COLIBRI}               \\
    \drc
    \cc QF\_IDL      & Yices2 \nc{Z3}     & Yices2 \nc{Z3}      & {Yices2}              & {Yices2}                \\
    \cc QF\_LIA      & CVC4 \nc{MathSAT}  & CVC4 \nc{MathSAT}   & CVC4 \nc{MathSAT}     & \cc {SPASS-SATT}        \\
    \drc
    \cc QF\_LIRA     & {Yices2}           & {Yices2}            & Yices2 \nc{Z3}        & Yices2 \nc{Z3}          \\
    \cc QF\_LRA      &  {CVC4}            & {CVC4}              & {CVC4}                & {CVC4}                  \\
    \cc QF\_NIA      & \cc AProVE \nc{Z3} & \cc Yices2 \nc{Z3}  & {CVC4}                & {CVC4}                  \\
    \cc QF\_NIRA     & {CVC4}             & {CVC4}              & \dcc {SMT-RAT}        & \dcc {SMT-RAT}          \\
    \cc QF\_NRA      & Yices2 \nc{Z3}     & Yices2 \nc{Z3}      & {Yices2}              & Yices2 \nc{Z3}          \\
    \cc QF\_RDL      & {Yices2}           & {Yices2}            & {Yices2}              & {Yices2}                \\
    \cc QF\_SLIA     & \cc                & \cc                 & \cc                   & \cc {CVC4} *            \\
    \cc QF\_UF       & {Yices2}           & {Yices2}            & {Yices2}              & {Yices2}                \\
    \drc
    \cc QF\_UFBV     & {Boolector}        & {Boolector}         & {Boolector}           & {Boolector}             \\
    \cc QF\_UFIDL    & {Yices2}           & {Yices2}            & {Yices2}              & {Yices2}                \\
    \cc QF\_UFLIA    & Yices2 \nc{Z3}     & Yices2 \nc{Z3}      & {Yices2}              & {Yices2}                \\
    \cc QF\_UFLRA    & {Yices2}           & {Yices2}            & {Yices2}              & {Yices2}                \\
    \cc QF\_UFNIA    & \cc {CVC4} *       & {Yices2/CVC4}       & {Yices2}              & {Yices2}                \\
    \cc QF\_UFNRA    & \cc CVC3 \nc{Z3} * & {Yices2}            & Yices2 \nc{Z3}        & {Yices2}                \\
    \cc UF           & {CVC4}             & {CVC4}              & \dcc {Vampire}        & \dcc {CVC4/Vampire}     \\
    \drc
    \cc UFBV         & CVC4 \nc{Z3} *     & CVC4 \nc{Z3}        & CVC4 \nc{Z3} *        & CVC4 \nc{Z3} *          \\
    \cc UFDT         & \cc                & \cc                 & {CVC4} *              & {CVC4}                  \\
    \rc
    \cc UFDTLIA      &                    &                     & {Vampire} *           & {CVC4}                  \\
    \cc UFIDL        & CVC4 \nc{Z3}       & {CVC4}              & {CVC4}                & CVC4 \nc{Z3}            \\
    \cc UFLIA        & {CVC4}             & {CVC4}              & {CVC4}                & {CVC4}                  \\
    \cc UFLRA        & \cc {CVC3}         & \cc Vampire \nc{Z3} & CVC4 \nc{Z3}          & CVC4 \nc{Z3}            \\
    \cc UFNIA        & \cc {CVC4} *       & {Vampire}           & {Vampire}             & Vampire \nc{Z3/Vampire} \\
    \bottomrule
  \end{tabular}}
\end{table}