### 2019-03-19 2018/report/table-winners-unsat-core.tex
Double check application and unsat score application winners table.

\begin{table}
  \caption{Best unsat-core track solvers (by division)}
  \label{tab:winners-unsat-core}
  \centering
  \scalebox{.88}{%
    \begin{tabular}{r@{\hskip 1em}>{\columncolor[gray]{.95}[.25em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c}
    \toprule
    \rc
    Division   & 2016             & 2017            & 2018               \\
    \hline \hline
    \cc ALIA       & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *           \\
    \cc AUFBVDTLIA & \cc                            & \cc                                 & CVC4 *                   \\
    \cc AUFLIA     & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 *                   \\
    \cc AUFLIRA    & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 *                   \\
    \cc AUFNIRA    & \cc \nc{Z3} *                  & CVC4 *                              & CVC4 *                   \\
    \cc BV         & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *           \\
    \cc LIA        & \cc \nc{veriT} *               & CVC4 *                              & CVC4 \nc{Z3} *           \\
    \cc LRA        & \cc \nc{veriT} *               & CVC4 *                              & CVC4 *                   \\
    \cc NIA        & \cc \nc{Z3} *                  & CVC4 / CVC4 \nc{Z3}*                & CVC4 *                   \\
    \cc NRA        & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *           \\
    \cc QF\_ABV    & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & Yices2                   \\
    \cc QF\_ABVFP  & \cc                            & \cc ---                             & \cc CVC4 *               \\
    \rc
    \cc QF\_ALIA   & SMTInterpol \nc{Z3} *          & SMTInterpol \nc{Z3}                 & Yices2 \nc{Z3}           \\
    \cc QF\_ANIA   & \cc \nc{Z3} *                  & CVC4 *                              & CVC4 *                   \\
    \rc
    \cc QF\_AUFBV  & \nc{Z3} *                      & CVC4 \nc{MathSAT} *                 & Yices2 \nc{MathSAT}      \\
    \cc QF\_AUFLIA & \cc SMTInterpol \nc{Z3} *      & CVC4                                & CVC4                     \\
    \cc QF\_AUFNIA & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *           \\
    \drc
    \cc QF\_AX     & SMTInterpol \nc{Z3} *          & SMTInterpol \nc{Z3}                 & \cc Yices2               \\
    \rc
    \cc QF\_BV     & \nc{MathSAT} *                 & CVC4 \nc{MathSAT} *                 & Yices2                   \\
    \cc QF\_BVFP   & \nc{Z3} *                      & \nc{Z3} *                           & \cc CVC4 \nc{Z3} *       \\
    \rc
    \cc QF\_DT     &                                &                                     & CVC4 *                   \\
    \rc
    \cc QF\_FP     & \nc{MathSAT} *                 & \nc{Z3} *                           & CVC4 *                   \\
    \cc QF\_IDL    & \cc SMTInterpol \nc{Z3} *      & CVC4 \nc{Z3}                        & CVC4 \nc{Z3}             \\
    \drc
    \cc QF\_LIA    & SMTInterpol \nc{Z3} *          & SMTInterpol / SMTInterpol \nc{Z3}   & SMTInterpol              \\
    \cc QF\_LIRA   & \cc SMTInterpol \nc{Z3} *      & CVC4 \nc{Z3}                        & CVC4 \nc{Z3}             \\
    \drc
    \cc QF\_LRA    & SMTInterpol *                  & SMTInterpol                         & SMTInterpol              \\
    \cc QF\_NIA    & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *           \\
    \cc QF\_NIRA   & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *           \\
    \cc QF\_NRA    & \cc \nc{Z3} *                  & CVC4 *                              & CVC4 *                   \\
    \cc QF\_RDL    & \cc SMTInterpol \nc{veriT} *   & CVC4 \nc{Z3}                        & CVC4 \nc{Z3}             \\
    \cc QF\_UF     & \cc SMTInterpol \nc{MathSAT} * & CVC4                                & CVC4                     \\
    \rc
    \cc QF\_UFBV   & \nc{MathSAT} *                 & CVC4 \nc{MathSAT} *                 & Yices2 \nc{Z3}           \\
    \drc
    \cc QF\_UFIDL  & SMTInterpol \nc{Z3} *          & SMTInterpol \nc{Z3}                 & \cc Yices2               \\
    \drc
    \cc QF\_UFLIA  & SMTInterpol \nc{MathSAT} *     & SMTInterpol \nc{Z3}                 & SMTInterpol \nc{Z3}      \\
    \drc
    \cc QF\_UFLRA  & SMTInterpol \nc{Z3} *          & SMTInterpol \nc{MathSAT}            & \cc Yices2 \nc{MathSAT}  \\
    \cc QF\_UFNIA  & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *           \\
    \cc QF\_UFNRA  & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *           \\
    \cc UF         & \cc \nc{Z3} *                  & CVC4 *                              & CVC4 *                   \\
    \cc UFBV       & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 *                   \\
    \rc
    \cc UFDT       &                                &                                     & CVC4 *                   \\
    \cc UFIDL      & \cc \nc{Z3} *                  & CVC4 *                              & CVC4 *                   \\
    \cc UFLIA      & \cc \nc{Z3} *                  & CVC4 *                              & CVC4 *                   \\
    \cc UFLRA      & \cc \nc{Z3} *                  & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *           \\
    \cc UFNIA      & \cc \nc{Z3} *                  & CVC4 *                              & CVC4 *                   \\
  \end{tabular}
  }
\end{table}