### 2019-03-20 2018/report/table-winners-main.tex
Use colors for winners tables.

\begin{table}
\caption{Best main-track solvers (by division)}
\label{tab:winners-main}
\centering
\scalebox{.83}{%
\renewcommand{\arraystretch}{1.01}%
  \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][0.5em]}c}
    \toprule
    Division         &  2015                    &  2016                     &  2017                          &  2018                     \\
    \hline \hline
    \wc ABVFP        &                          &                           &                                & \cc{cvc4} {CVC4} *        \\
    \rc{cvc4}
    \wc ALIA         & CVC4 \nc{Z3}             & CVC4 \nc{Z3}              & CVC4 \nc{Z3}                   & CVC4 \nc{Z3}              \\
    \wc AUFBVDTLIA   &                          &                           & \cc{cvc4} {CVC4} *             & \cc{cvc4} {CVC4} *        \\
    \wc AUFDTLIA     &                          &                           & \cc{cvc4} {CVC4} *             & \cc{cvc4} {CVC4}          \\
    \rc{cvc4}
    \wc AUFLIA       & {CVC4}                   & {CVC4}                    & {CVC4}                         & {CVC4}                    \\
    \rc{vamp}
    \wc AUFLIRA      & \cc{cvc4} CVC4 \nc{Z3}   & Vampire \nc{Z3}           & Vampire \nc{Z3}                & \cc{cvc4} CVC4 \nc{Z3}    \\
    \rc{vamp}
    \wc AUFNIRA      & \cc{cvc4} {CVC4} *       & {Vampire}                 & {Vampire}                      & \cc{cvc4} {CVC4}          \\
    \rc{q3b}
    \wc BV           & \cc{cvc4} CVC4 \nc{Z3} * & {Q3B}                     & Q3B \nc{Z3}                    &  \cc{cvc4} {CVC4}         \\
    \wc BVFP         &                          &                           &                                & \cc{cvc4} {CVC4} *        \\
    \wc FP           &                          &                           &                                & \cc{cvc4} {CVC4} *        \\
    \rc{cvc4}
    \wc LIA          & {CVC4}                   & {CVC4}                    & CVC4 \nc{Z3}                   & CVC4 \nc{Z3}              \\
    \rc{cvc4}
    \wc LRA          & {CVC4}                   & {CVC4}                    & CVC4 \nc{Z3}                   & CVC4 \nc{Z3}              \\
    \rc{cvc4}
    \wc NIA          & CVC4 \nc{Z3} *           & \cc{prob} ProB \nc{Z3}    & CVC4 \nc{Z3}                   & CVC4 \nc{Z3}              \\
    \rc{vamp}
    \wc NRA          & \cc{cvc4} {CVC4} *       & {Vampire}                 & \cc{redl} {Redlog}             & Vampire \nc{Z3} / Vampire \\
    \rc{bool}
    \wc QF\_ABV      & {Boolector}              & {Boolector}               & {Boolector}                    & {Boolector}               \\
    \wc QF\_ABVFP    &                          &                           & {---}                          & \cc{cvc4} {CVC4}          \\
    \rc{yices}
    \wc QF\_ALIA     & {Yices2}                 & {Yices2}                  & {Yices2}                       & {Yices2}                  \\
    \rc{cvc4}
    \wc QF\_ANIA     & CVC4 \nc{Z3} *           & {CVC4} *                  & {CVC4} *                       & CVC4 \nc{Z3} *            \\
    \rc{cvc4}
    \wc QF\_AUFBV    & CVC4 \nc{MathSAT}        & CVC4 \nc{MathSAT}         & \cc{yices} Yices2 \nc{MathSAT} & {CVC4}                    \\
    \rc{yices}
    \wc QF\_AUFLIA   & {Yices2}                 & {Yices2}                  & {Yices2}                       & {Yices2}                  \\
    \rc{cvc4}
    \wc QF\_AUFNIA   & {CVC4} *                 & {CVC4} *                  & CVC4 \nc{Z3} *                 & CVC4 \nc{Z3} *            \\
    \rc{yices}
    \wc QF\_AX       & {Yices2}                 & {Yices2}                  & {Yices2}                       & {Yices2}                  \\
    \rc{bool}
    \wc QF\_BV       & {Boolector}              & {Boolector}               & {Boolector/MinkeyRink}         & {Boolector/MinkeyRink}    \\
    \wc QF\_BVFP     & \cc{z} {Z3} *            & \wc \nc{Z3} *             & \cc{coli} COLIBRI \nc{Z3} *    & \cc{cvc4} {CVC4}          \\
    \wc QF\_DT       &                          &                           & \cc{cvc4} {CVC4} *             & \cc{cvc4} {CVC4} *        \\
    \rc{coli}
    \wc QF\_FP       & \cc{z} {Z3} *            & \wc \nc{MathSAT} *        & COLIBRI \nc{Z3}                & {COLIBRI}                 \\
    \rc{yices}
    \wc QF\_IDL      & Yices2 \nc{Z3}           & Yices2 \nc{Z3}            & {Yices2}                       & {Yices2}                  \\
    \rc{cvc4}
    \wc QF\_LIA      & CVC4 \nc{MathSAT}        & CVC4 \nc{MathSAT}         & CVC4 \nc{MathSAT}              & \cc{spass} {SPASS-SATT}   \\
    \rc{yices}
    \wc QF\_LIRA     & {Yices2}                 & {Yices2}                  & Yices2 \nc{Z3}                 & Yices2 \nc{Z3}            \\
    \rc{cvc4}
    \wc QF\_LRA      &  {CVC4}                  & {CVC4}                    & {CVC4}                         & {CVC4}                    \\
    \rc{cvc4}
    \wc QF\_NIA      & \cc{apr} AProVE \nc{Z3}  & \cc{yices} Yices2 \nc{Z3} & {CVC4}                         & {CVC4}                    \\
    \rc{cvc4}
    \wc QF\_NIRA     & {CVC4}                   & {CVC4}                    & \cc{rat} {SMT-RAT}             & \cc{rat} {SMT-RAT}        \\
    \rc{yices}
    \wc QF\_NRA      & Yices2 \nc{Z3}           & Yices2 \nc{Z3}            & {Yices2}                       & Yices2 \nc{Z3}            \\
    \rc{yices}
    \wc QF\_RDL      & {Yices2}                 & {Yices2}                  & {Yices2}                       & {Yices2}                  \\
    \wc QF\_SLIA     &                          &                           &                                & \cc{cvc4} {CVC4} *        \\
    \rc{yices}
    \wc QF\_UF       & {Yices2}                 & {Yices2}                  & {Yices2}                       & {Yices2}                  \\
    \rc{bool}
    \wc QF\_UFBV     & {Boolector}              & {Boolector}               & {Boolector}                    & {Boolector}               \\
    \rc{yices}
    \wc QF\_UFIDL    & {Yices2}                 & {Yices2}                  & {Yices2}                       & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFLIA    & Yices2 \nc{Z3}           & Yices2 \nc{Z3}            & {Yices2}                       & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFLRA    & {Yices2}                 & {Yices2}                  & {Yices2}                       & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFNIA    & \cc{cvc4} {CVC4} *       & {Yices2/CVC4}             & {Yices2}                       & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFNRA    & \cc{cvc3} CVC3 \nc{Z3} * & {Yices2}                  & Yices2 \nc{Z3}                 & {Yices2}                  \\
    \rc{cvc4}
    \wc UF           & {CVC4}                   & {CVC4}                    & \cc{vamp} {Vampire}            &  {CVC4/Vampire}           \\
    \rc{cvc4}
    \wc UFBV         & CVC4 \nc{Z3} *           & CVC4 \nc{Z3}              & CVC4 \nc{Z3} *                 & CVC4 \nc{Z3} *            \\
    \wc UFDT         &                          &                           & \cc{cvc4} {CVC4} *             & \cc{cvc4} {CVC4}          \\
    \wc UFDTLIA      &                          &                           & \cc{vamp} {Vampire} *          & \cc{cvc4} {CVC4}          \\
    \rc{cvc4}
    \wc UFIDL        & CVC4 \nc{Z3}             & {CVC4}                    & {CVC4}                         & CVC4 \nc{Z3}              \\
    \rc{cvc4}
    \wc UFLIA        & {CVC4}                   & {CVC4}                    & {CVC4}                         & {CVC4}                    \\
    \rc{cvc4}
    \wc UFLRA        & \cc{cvc3} {CVC3}         & \cc{vamp} Vampire \nc{Z3} & CVC4 \nc{Z3}                   & CVC4 \nc{Z3}              \\
    \rc{vamp}
    \wc UFNIA        & \cc{cvc4} {CVC4} *       & {Vampire}                 & {Vampire}                      & Vampire \nc{Z3} / Vampire \\
    \bottomrule
  \end{tabular}}
\end{table}