### 2019-03-20 2018/report/table-winners-unsat-core.tex

\begin{table}
  \caption{Best unsat-core track solvers (by division)}
  \label{tab:winners-unsat-core}
  \centering
  \scalebox{.88}{%
    \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c}
      \toprule
      \rc{white}
      Division       & 2016                                 & 2017                                & 2018                           \\
      \hline \hline
      \rc{cvc4}
      \wc ALIA       & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *                 \\
      \wc AUFBVDTLIA &                                      &                                     & \cc{cvc4} CVC4 *               \\
      \rc{cvc4}
      \wc AUFLIA     & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 *                         \\
      \rc{cvc4}
      \wc AUFLIRA    & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 *                         \\
      \rc{cvc4}
      \wc AUFNIRA    & \wc \nc{Z3} *                        & CVC4 *                              & CVC4 *                         \\
      \rc{cvc4}
      \wc BV         & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *                 \\
      \rc{cvc4}
      \wc LIA        & \wc \nc{veriT} *                     & CVC4 *                              & CVC4 \nc{Z3} *                 \\
      \rc{cvc4}
      \wc LRA        & \wc \nc{veriT} *                     & CVC4 *                              & CVC4 *                         \\
      \rc{cvc4}
      \wc NIA        & \wc \nc{Z3} *                        & CVC4 / CVC4 \nc{Z3}*                & CVC4 *                         \\
      \rc{cvc4}
      \wc NRA        & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *                 \\
      \wc QF\_ABV    & \wc \nc{Z3} *                        & \cc{cvc4} CVC4 \nc{Z3} *            & \cc{yices} Yices2              \\
      \wc QF\_ABVFP  &                                      & ---                                 & \cc{cvc4} CVC4 *               \\
      \rc{smti}
      \wc QF\_ALIA   & SMTInterpol \nc{Z3} *                & SMTInterpol \nc{Z3}                 & \cc{yices} Yices2 \nc{Z3}      \\
      \rc{cvc4}
      \wc QF\_ANIA   & \wc \nc{Z3} *                        & CVC4 *                              & CVC4 *                         \\
      \wc QF\_AUFBV  & \nc{Z3} *                            & \cc{cvc4} CVC4 \nc{MathSAT} *       & \cc{yices} Yices2 \nc{MathSAT} \\
      \rc{cvc4}
      \wc QF\_AUFLIA & \cc{smti} SMTInterpol \nc{Z3} *      & CVC4                                & CVC4                           \\
      \rc{cvc4}
      \wc QF\_AUFNIA & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *                 \\
      \rc{smti}
      \wc QF\_AX     & SMTInterpol \nc{Z3} *                & SMTInterpol \nc{Z3}                 & \cc{yices} Yices2              \\
      \wc QF\_BV     & \nc{MathSAT} *                       & \cc{cvc4} CVC4 \nc{MathSAT} *       & \cc{yices} Yices2              \\
      \wc QF\_BVFP   & \nc{Z3} *                            & \nc{Z3} *                           & \cc{cvc4} CVC4 \nc{Z3} *       \\
      \wc QF\_DT     &                                      &                                     & \cc{cvc4} CVC4 *               \\
      \wc QF\_FP     & \nc{MathSAT} *                       & \nc{Z3} *                           & \cc{cvc4} CVC4 *               \\
      \rc{cvc4}
      \wc QF\_IDL    & \cc{smti} SMTInterpol \nc{Z3} *      & CVC4 \nc{Z3}                        & CVC4 \nc{Z3}                   \\
      \rc{smti}
      \wc QF\_LIA    & SMTInterpol \nc{Z3} *                & SMTInterpol / SMTInterpol \nc{Z3}   & SMTInterpol                    \\
      \rc{cvc4}
      \wc QF\_LIRA   & \cc{smti} SMTInterpol \nc{Z3} *      & CVC4 \nc{Z3}                        & CVC4 \nc{Z3}                   \\
      \rc{smti}
      \wc QF\_LRA    & SMTInterpol *                        & SMTInterpol                         & SMTInterpol                    \\
      \rc{cvc4}
      \wc QF\_NIA    & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *                 \\
      \rc{cvc4}
      \wc QF\_NIRA   & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *                 \\
      \rc{cvc4}
      \wc QF\_NRA    & \wc \nc{Z3} *                        & CVC4 *                              & CVC4 *                         \\
      \rc{cvc4}
      \wc QF\_RDL    & \cc{smti} SMTInterpol \nc{veriT} *   & CVC4 \nc{Z3}                        & CVC4 \nc{Z3}                   \\
      \rc{cvc4}
      \wc QF\_UF     & \cc{smti} SMTInterpol \nc{MathSAT} * & CVC4                                & CVC4                           \\
      \wc QF\_UFBV   & \nc{MathSAT} *                       & \cc{cvc4} CVC4 \nc{MathSAT} *       & \cc{yices} Yices2 \nc{Z3}      \\
      \rc{smti}
      \wc QF\_UFIDL  & SMTInterpol \nc{Z3} *                & SMTInterpol \nc{Z3}                 & \cc{yices} Yices2              \\
      \rc{smti}
      \wc QF\_UFLIA  & SMTInterpol \nc{MathSAT} *           & SMTInterpol \nc{Z3}                 & SMTInterpol \nc{Z3}            \\
      \rc{smti}
      \wc QF\_UFLRA  & SMTInterpol \nc{Z3} *                & SMTInterpol \nc{MathSAT}            & \cc{yices} Yices2 \nc{MathSAT} \\
      \rc{cvc4}
      \wc QF\_UFNIA  & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *                 \\
      \rc{cvc4}
      \wc QF\_UFNRA  & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *                 \\
      \rc{cvc4}
      \wc UF         & \wc \nc{Z3} *                        & CVC4 *                              & CVC4 *                         \\
      \rc{cvc4}
      \wc UFBV       & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 *                         \\
      \wc UFDT       &                                      &                                     & \cc{cvc4} CVC4 *               \\
      \rc{cvc4}
      \wc UFIDL      & \wc \nc{Z3} *                        & CVC4 *                              & CVC4 *                         \\
      \rc{cvc4}
      \wc UFLIA      & \wc \nc{Z3} *                        & CVC4 *                              & CVC4 *                         \\
      \rc{cvc4}
      \wc UFLRA      & \wc \nc{Z3} *                        & CVC4 \nc{Z3} *                      & CVC4 \nc{Z3} *                 \\
      \rc{cvc4}
      \wc UFNIA      & \wc \nc{Z3} *                        & CVC4 *                              & CVC4 *                         \\
    \end{tabular}
  }
\end{table}