### 2019-03-20 2018/report/table-winners-main.tex
Use gray text font instead of * for non-competitive divisions.

\begin{table}
\caption{Best main-track solvers (by division)}
\label{tab:winners-main}
\centering
\scalebox{.83}{%
\renewcommand{\arraystretch}{1.01}%
  \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][0.5em]}c}
    \toprule
    Division         &  2015                        &  2016                     &  2017                           &  2018                     \\
    \hline \hline
    \wc ABVFP        &                              &                           &                                 & \nonc \cc{cvc4} {CVC4}    \\
    \rc{cvc4}
    \wc ALIA         & CVC4 \nc{Z3}                 & CVC4 \nc{Z3}              & CVC4 \nc{Z3}                    & CVC4 \nc{Z3}              \\
    \wc AUFBVDTLIA   &                              &                           & \nonc \cc{cvc4} {CVC4}          & \nonc \cc{cvc4} {CVC4}    \\
    \wc AUFDTLIA     &                              &                           & \nonc \cc{cvc4} {CVC4}          & \cc{cvc4} {CVC4}          \\
    \rc{cvc4}
    \wc AUFLIA       & {CVC4}                       & {CVC4}                    & {CVC4}                          & {CVC4}                    \\
    \rc{vamp}
    \wc AUFLIRA      & \cc{cvc4} CVC4 \nc{Z3}       & Vampire \nc{Z3}           & Vampire \nc{Z3}                 & \cc{cvc4} CVC4 \nc{Z3}    \\
    \rc{vamp}
    \wc AUFNIRA      & \nonc \cc{cvc4} {CVC4}       & {Vampire}                 & {Vampire}                       & \cc{cvc4} {CVC4}          \\
    \rc{q3b}
    \wc BV           & \nonc \cc{cvc4} CVC4 \nc{Z3} & {Q3B}                     & Q3B \nc{Z3}                     & \cc{cvc4} {CVC4}          \\
    \wc BVFP         &                              &                           &                                 & \nonc \cc{cvc4} {CVC4}    \\
    \wc FP           &                              &                           &                                 & \nonc \cc{cvc4} {CVC4}    \\
    \rc{cvc4}
    \wc LIA          & {CVC4}                       & {CVC4}                    & CVC4 \nc{Z3}                    & CVC4 \nc{Z3}              \\
    \rc{cvc4}
    \wc LRA          & {CVC4}                       & {CVC4}                    & CVC4 \nc{Z3}                    & CVC4 \nc{Z3}              \\
    \rc{cvc4}
    \wc NIA          & \nonc CVC4 \nc{Z3}           & \cc{prob} ProB \nc{Z3}    & CVC4 \nc{Z3}                    & CVC4 \nc{Z3}              \\
    \rc{vamp}
    \wc NRA          & \nonc \cc{cvc4} {CVC4}       & {Vampire}                 & \cc{redl} {Redlog}              & Vampire \nc{Z3} / Vampire \\
    \rc{bool}
    \wc QF\_ABV      & {Boolector}                  & {Boolector}               & {Boolector}                     & {Boolector}               \\
    \wc QF\_ABVFP    &                              &                           & {---}                           & \cc{cvc4} {CVC4}          \\
    \rc{yices}
    \wc QF\_ALIA     & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{cvc4}
    \wc QF\_ANIA     & \nonc CVC4 \nc{Z3}           & \nonc {CVC4}                & \nonc {CVC4}                  & \nonc CVC4 \nc{Z3}        \\
    \rc{cvc4}
    \wc QF\_AUFBV    & CVC4 \nc{MathSAT}            & CVC4 \nc{MathSAT}         & \cc{yices} Yices2 \nc{MathSAT}  & {CVC4}                    \\
    \rc{yices}
    \wc QF\_AUFLIA   & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{cvc4}
    \wc QF\_AUFNIA   & \nonc {CVC4}                 & \nonc {CVC4}              & \nonc CVC4 \nc{Z3}              & \nonc CVC4 \nc{Z3}        \\
    \rc{yices}
    \wc QF\_AX       & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{bool}
    \wc QF\_BV       & {Boolector}                  & {Boolector}               & {Boolector/MinkeyRink}          & {Boolector/MinkeyRink}    \\
    \wc QF\_BVFP     & \nonc \cc{z} {Z3}            & \nonc \wc \nc{Z3}         & \nonc \cc{coli} COLIBRI \nc{Z3} & \cc{cvc4} {CVC4}          \\
    \wc QF\_DT       &                              &                           & \nonc \cc{cvc4} {CVC4}          & \nonc \cc{cvc4} {CVC4}    \\
    \rc{coli}
    \wc QF\_FP       & \nonc \cc{z} {Z3}            & \nonc \wc \nc{MathSAT}    & COLIBRI \nc{Z3}                 & {COLIBRI}                 \\
    \rc{yices}
    \wc QF\_IDL      & Yices2 \nc{Z3}               & Yices2 \nc{Z3}            & {Yices2}                        & {Yices2}                  \\
    \rc{cvc4}
    \wc QF\_LIA      & CVC4 \nc{MathSAT}            & CVC4 \nc{MathSAT}         & CVC4 \nc{MathSAT}               & \cc{spass} {SPASS-SATT}   \\
    \rc{yices}
    \wc QF\_LIRA     & {Yices2}                     & {Yices2}                  & Yices2 \nc{Z3}                  & Yices2 \nc{Z3}            \\
    \rc{cvc4}
    \wc QF\_LRA      &  {CVC4}                      & {CVC4}                    & {CVC4}                          & {CVC4}                    \\
    \rc{cvc4}
    \wc QF\_NIA      & \cc{apr} AProVE \nc{Z3}      & \cc{yices} Yices2 \nc{Z3} & {CVC4}                          & {CVC4}                    \\
    \rc{cvc4}
    \wc QF\_NIRA     & {CVC4}                       & {CVC4}                    & \cc{rat} {SMT-RAT}              & \cc{rat} {SMT-RAT}        \\
    \rc{yices}
    \wc QF\_NRA      & Yices2 \nc{Z3}               & Yices2 \nc{Z3}            & {Yices2}                        & Yices2 \nc{Z3}            \\
    \rc{yices}
    \wc QF\_RDL      & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \wc QF\_SLIA     &                              &                           &                                 & \nonc \cc{cvc4} {CVC4}    \\
    \rc{yices}
    \wc QF\_UF       & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{bool}
    \wc QF\_UFBV     & {Boolector}                  & {Boolector}               & {Boolector}                     & {Boolector}               \\
    \rc{yices}
    \wc QF\_UFIDL    & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFLIA    & Yices2 \nc{Z3}               & Yices2 \nc{Z3}            & {Yices2}                        & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFLRA    & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFNIA    & \nonc \cc{cvc4} {CVC4}       & {Yices2/CVC4}             & {Yices2}                        & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFNRA    & \nonc \cc{cvc3} CVC3 \nc{Z3} & {Yices2}                  & Yices2 \nc{Z3}                  & {Yices2}                  \\
    \rc{cvc4}
    \wc UF           & {CVC4}                       & {CVC4}                    & \cc{vamp} {Vampire}             &  {CVC4/Vampire}           \\
    \rc{cvc4}
    \wc UFBV         & \nonc CVC4 \nc{Z3}           & CVC4 \nc{Z3}              & \nonc CVC4 \nc{Z3}              & \nonc CVC4 \nc{Z3}        \\
    \wc UFDT         &                              &                           & \nonc \cc{cvc4} {CVC4}          & \cc{cvc4} {CVC4}          \\
    \wc UFDTLIA      &                              &                           & \nonc \cc{vamp} {Vampire}       & \cc{cvc4} {CVC4}          \\
    \rc{cvc4}
    \wc UFIDL        & CVC4 \nc{Z3}                 & {CVC4}                    & {CVC4}                          & CVC4 \nc{Z3}              \\
    \rc{cvc4}
    \wc UFLIA        & {CVC4}                       & {CVC4}                    & {CVC4}                          & {CVC4}                    \\
    \rc{cvc4}
    \wc UFLRA        & \cc{cvc3} {CVC3}             & \cc{vamp} Vampire \nc{Z3} & CVC4 \nc{Z3}                    & CVC4 \nc{Z3}              \\
    \rc{vamp}
    \wc UFNIA        & \nonc \cc{cvc4} {CVC4}       & {Vampire}                 & {Vampire}                       & Vampire \nc{Z3} / Vampire \\
    \bottomrule
  \end{tabular}}
\end{table}