### 2019-03-20 2018/report/table-winners-unsat-core.tex

\begin{table}
  \caption{Best unsat-core track solvers (by division)}
  \label{tab:winners-unsat-core}
  \centering
  \scalebox{.88}{%
    \begin{tabular}{r@{\hskip 1em}>{\nonc \columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c}
      \toprule
      \rc{white}
      Division       & 2016                                & 2017                              & 2018                           \\
      \hline \hline
      \rc{cvc4}
      \wc ALIA       & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4 \nc{Z3}             \\
      \wc AUFBVDTLIA &                                     &                                   & \nonc \cc{cvc4} CVC4           \\
      \rc{cvc4}
      \wc AUFLIA     & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4                     \\
      \rc{cvc4}
      \wc AUFLIRA    & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4                     \\
      \rc{cvc4}
      \wc AUFNIRA    & \wc \nc{Z3}                         & \nonc CVC4                        & \nonc CVC4                     \\
      \rc{cvc4}
      \wc BV         & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4 \nc{Z3}             \\
      \rc{cvc4}
      \wc LIA        & \wc \nc{veriT}                      & \nonc CVC4                        & \nonc CVC4 \nc{Z3}             \\
      \rc{cvc4}
      \wc LRA        & \wc \nc{veriT}                      & \nonc CVC4                        & \nonc CVC4                     \\
      \rc{cvc4}
      \wc NIA        & \wc \nc{Z3}                         & \nonc CVC4 / CVC4 \nc{Z3}         & \nonc CVC4                     \\
      \rc{cvc4}
      \wc NRA        & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4 \nc{Z3}             \\
      \wc QF\_ABV    & \wc \nc{Z3}                         & \nonc \cc{cvc4} CVC4 \nc{Z3}      & \cc{yices} Yices2              \\
      \wc QF\_ABVFP  &                                      & ---                              & \nonc \cc{cvc4} CVC4           \\
      \rc{smti}
      \wc QF\_ALIA   & SMTInterpol \nc{Z3}                 & SMTInterpol \nc{Z3}               & \cc{yices} Yices2 \nc{Z3}      \\
      \rc{cvc4}
      \wc QF\_ANIA   & \wc \nc{Z3}                         & \nonc CVC4                        & \nonc CVC4                     \\
      \wc QF\_AUFBV  & \nc{Z3}                             & \nonc \cc{cvc4} CVC4 \nc{MathSAT} & \cc{yices} Yices2 \nc{MathSAT} \\
      \rc{cvc4}
      \wc QF\_AUFLIA & \cc{smti} SMTInterpol \nc{Z3}       & CVC4                              & CVC4                           \\
      \rc{cvc4}
      \wc QF\_AUFNIA & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4 \nc{Z3}             \\
      \rc{smti}
      \wc QF\_AX     & SMTInterpol \nc{Z3}                 & SMTInterpol \nc{Z3}               & \cc{yices} Yices2              \\
      \wc QF\_BV     & \nc{MathSAT}                        & \nonc \cc{cvc4} CVC4 \nc{MathSAT} & \cc{yices} Yices2              \\
      \wc QF\_BVFP   & \nc{Z3}                             & \nonc \nc{Z3}                     & \nonc \cc{cvc4} CVC4 \nc{Z3}   \\
      \wc QF\_DT     &                                     &                                   & \nonc \cc{cvc4} CVC4           \\
      \wc QF\_FP     & \nc{MathSAT}                        & \nonc \nc{Z3}                     & \nonc \cc{cvc4} CVC4           \\
      \rc{cvc4}
      \wc QF\_IDL    & \cc{smti} SMTInterpol \nc{Z3}       & CVC4 \nc{Z3}                      & CVC4 \nc{Z3}                   \\
      \rc{smti}
      \wc QF\_LIA    & SMTInterpol \nc{Z3}                 & SMTInterpol / SMTInterpol \nc{Z3} & SMTInterpol                    \\
      \rc{cvc4}
      \wc QF\_LIRA   & \cc{smti} SMTInterpol \nc{Z3}       & CVC4 \nc{Z3}                      & CVC4 \nc{Z3}                   \\
      \rc{smti}
      \wc QF\_LRA    & SMTInterpol                         & SMTInterpol                       & SMTInterpol                    \\
      \rc{cvc4}
      \wc QF\_NIA    & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4 \nc{Z3}             \\
      \rc{cvc4}
      \wc QF\_NIRA   & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4 \nc{Z3}             \\
      \rc{cvc4}
      \wc QF\_NRA    & \wc \nc{Z3}                         & \nonc CVC4                        & \nonc CVC4                     \\
      \rc{cvc4}
      \wc QF\_RDL    & \cc{smti} SMTInterpol \nc{veriT}    & CVC4 \nc{Z3}                      & CVC4 \nc{Z3}                   \\
      \rc{cvc4}
      \wc QF\_UF     & \cc{smti} SMTInterpol \nc{MathSAT}  & CVC4                              & CVC4                           \\
      \wc QF\_UFBV   & \nc{MathSAT}                        & \nonc \cc{cvc4} CVC4 \nc{MathSAT} & \cc{yices} Yices2 \nc{Z3}      \\
      \rc{smti}
      \wc QF\_UFIDL  & SMTInterpol \nc{Z3}                 & SMTInterpol \nc{Z3}               & \cc{yices} Yices2              \\
      \rc{smti}
      \wc QF\_UFLIA  & SMTInterpol \nc{MathSAT}            & SMTInterpol \nc{Z3}               & SMTInterpol \nc{Z3}            \\
      \rc{smti}
      \wc QF\_UFLRA  & SMTInterpol \nc{Z3}                 & SMTInterpol \nc{MathSAT}          & \cc{yices} Yices2 \nc{MathSAT} \\
      \rc{cvc4}
      \wc QF\_UFNIA  & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4 \nc{Z3}             \\
      \rc{cvc4}
      \wc QF\_UFNRA  & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4 \nc{Z3}             \\
      \rc{cvc4}
      \wc UF         & \wc \nc{Z3}                         & \nonc CVC4                        & \nonc CVC4                     \\
      \rc{cvc4}
      \wc UFBV       & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4                     \\
      \wc UFDT       &                                     &                                   & \nonc \cc{cvc4} CVC4           \\
      \rc{cvc4}
      \wc UFIDL      & \wc \nc{Z3}                         & \nonc CVC4                        & \nonc CVC4                     \\
      \rc{cvc4}
      \wc UFLIA      & \wc \nc{Z3}                         & \nonc CVC4                        & \nonc CVC4                     \\
      \rc{cvc4}
      \wc UFLRA      & \wc \nc{Z3}                         & \nonc CVC4 \nc{Z3}                & \nonc CVC4 \nc{Z3}             \\
      \rc{cvc4}
      \wc UFNIA      & \wc \nc{Z3}                         & \nonc CVC4                        & \nonc CVC4                     \\
    \end{tabular}
  }
\end{table}