### 2019-04-14 2018/report/table-winners-main-sat-unsat.tex

\begin{table}
  \caption{Alternative best solvers in the Main track if only \emph{satisfiable} benchmarks had been considered. These results are in reference to those given in Table~\ref{tab:winners-main}.}
  \label{tab:results:sat}
\centering
  \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][0.5em]}c}
    \toprule
    Division         &  2015                &  2016                   &  2017                   &  2018                     \\
    \hline \hline
    ALIA             &                      &                         &                         & \cc{alt} Alt-Ergo         \\
    AUFLIA           & \cc{cvc3} CVC3       &                         &                         & \cc{alt} Alt-Ergo         \\
    AUFLIRA          &                      & \cc{cvc4} CVC4          & \cc{cvc4} CVC4          & \cc{alt} Alt-Ergo         \\
    AUFNIRA          & \nonc\nc{Z3}         & \cc{cvc4} CVC4 \nc{Z3}  & \cc{cvc4} CVC4 \nc{Z3}  & \nc{Z3}                   \\
    BV               &                      & \nc{Z3}                 & \cc{bool} Boolector     & \cc{bool} Boolector       \\
    LRA              &                      &                         & \cc{verit} veriT        &                           \\
    NIA              & \nonc \cc{cvc3} CVC3 &                         &                         &                           \\
    NRA              & \nonc \cc{cvc3} CVC3 & \cc{cvc4} CVC4          & \cc{cvc4} CVC4          & \cc{cvc4} CVC4 \nc{Z3}    \\
    QF\_ABV          &                      & \cc{yices} Yices2       & \cc{yices} Yices2       &                           \\
    QF\_AUFBV        & \cc{yices} Yices     & \cc{yices} Yices2       &                         & \cc{yices} Yices2         \\
    QF\_BV           &                      & \nc{Z3}/MinkeyRink      &                         &                           \\
    QF\_FP           &                      &                         &                         & \cc{cvc4} CVC4            \\
    QF\_LIA          & \nc{Z3}              &                         &                         &                           \\
    QF\_LRA          &                      &                         & \cc{yices} Yices2       & \cc{yices} Yices2         \\
    QF\_NIA          &                      & \cc{rat} SMT-RAT        &                         &                           \\
    QF\_NIRA         &                      &                         & \cc{cvc4} CVC4          &                           \\
    QF\_NRA          &                      &                         & \nc{Z3}                 &                           \\
    QF\_UFBV         &                      &                         &                         & \cc{yices} Yices2         \\
    QF\_UFIDL        & \nc{Z3}              & \nc{Z3}                 & \nc{Z3}                 & \cc{verit} veriT  \nc{Z3} \\
    QF\_UFNIA        & \nonc \cc{cvc3} CVC3 &                         & \cc{cvc4} CVC4          &                           \\
    UF               &                      & \cc{vamp} Vampire       &                         & \cc{vamp}Vampire/-        \\
    UFBV             &                      & \cc{bool} Boolector     &                         &                           \\
    UFDTLIA          &                      &                         & \nonc \cc{cvc4} CVC4    &                           \\
    UFLIA            & \nc{Z3}              &                         &                         &                           \\
    UFLRA            & \cc{cvc3} CVC3       & \cc{cvc4} CVC4          &                         &                           \\
    UFNIA            &                      & \cc{cvc4} CVC4          & \cc{cvc4} CVC4          & \cc{cvc4} CVC4            \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}
  \caption{Alternative best solvers in the Main track if only \emph{unsatisfiable} benchmarks had been considered. These results are in reference to those given in Table~\ref{tab:winners-main}.}
  \label{tab:results:unsat}
  \centering
  \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][0.5em]}c}
\toprule
Division        &  2015                  &  2016                     &  2017                 &  2018                   \\
\hline \hline
    BV          &                        & \cc{cvc4} CVC4            &                       &                         \\
    NIA         &                        & \cc{cvc4} CVC4            & \cc{vamp} Vampire     & \cc{vamp} Vampire       \\
    QF\_AUFBV   &                        &                           & \cc{cvc4} CVC4        &                         \\
    QF\_AUFNIA  & \nonc \nc{Z3}          & \nonc \nc{Z3}             &                       &                         \\
    QF\_BV      &                        &                           & -/Boolector           &                         \\
    QF\_IDL     & \cc{cvc4} CVC4         &                           & \nc{Z3}               & \cc{cvc4} CVC4 \nc{Z3}  \\
    QF\_LIA     & \cc{smti} SMTInterpol  & \cc{smti} SMTInterpol     & \cc{smti} SMTInterpol & \cc{smti} SMTInterpol   \\
    QF\_NIA     & \cc{cvc3} CVC3         &                           & \cc{yices} Yices2     &                         \\
    QF\_NRA     &                        &                           &                       & \cc{cvc4} CVC4          \\
    QF\_UFLIA   &                        &                           & \nc{Z3}               &                         \\
    QF\_UFNRA   &                        & \cc{cvc4} CVC4 \nc{Z3}    & \cc{cvc4} CVC4        & \cc{cvc4} CVC4 \nc{Z3}  \\
    UF          &                        &                           & \cc{cvc4} CVC4        & -/CVC4                  \\
    UFIDL       &                        & \cc{vamp} Vampire \nc{Z3} & \nc{Z3}               &                         \\
    UFLRA       &                        &                           & \cc{verit} veriT      & \cc{verit} veriT        \\
    \bottomrule
  \end{tabular}

\end{table}