### 2019-04-14 2018/report/table-winners-main-unknown.tex
Report: Layout and clean up main track winners tables for families and unknown.

\begin{table}
  \caption{Alternative best solvers if benchmarks with unknown status had not bee included in 2017 onwards.}
  \label{tab:results:unknowns}
  \centering
    \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][0.5em]}c}
    \toprule
    Division &  2017          &  2018                  \\
    \hline\hline
    AUFNIRA  & \cc{cvc4} CVC4 &                        \\
    QF\_NIRA & \nc{Z3}        & \cc{cvc4} CVC4 \nc{Z3} \\
    QF\_NRA  & \nc{Z3}        &                        \\
    UFDTLIA  & \cc{cvc4} CVC4 &                        \\
    BV       &                & \cc{bool} Boolector    \\
    QF\_FP   &                & \cc{cvc4} CVC4 \nc{Z3} \\
    UFNIA    &                & \cc{cvc4} CVC4         \\
    NRA      &                & -/\nc{Z3}              \\
    UF       &                & -/CVC4                 \\
  \bottomrule
  \end{tabular}
\end{table}