### 2019-08-16 2018/report/Reviews_67398_Revised_Response.txt
Add response to the reviews for the revised version

> *** Review 1 ***
>
>  The paper overall improved since the first submission, but there are still issues to be fixed, some of them already reported in the review of the first version.
>
>  * Organization
>
> - information about ranking should all stay in the related section, while
> there are related information, e.g. in the first part of Section 7 - Results.

TODO

> - outline of the content of Section 8 describes only part of its content (the first three sub-sections).

TODO

> * Quality
>
> - The paper still contains a number of issues, minor or not, detailed at the end of this review.
>
> -  In Section 2 the fact that "appointment should happen ... " is of relative interest and could be omitted. The same holds for the day in which the events took place.

TODO

> * Rules
>
> - basically incorrect results are "ignored" by omitting the benchmark if the status is not known - this comment still remains. The authors updated the text. I see their points and analysis, which I appreciated, through I think that this is an issue that should be dealt with more care. Correctness is very important for solvers when checking in particular real benchmarks (HW- and SW-verification, etc).

TODO

> * Significance of the results
>
>  - the paper and its analysis now contain many numbers and data, which is very appropriate for e competition report in my opinion. But, basically, no explanation is in place. What I expect to get out of a report, other than commenting the numbers, is their explanation, e.g.: what are the techniques implemented in solvers that overall performed better? Rr better, what techniques implemented in solvers led to an efficient solvers on a particular benchmark? why such techniques led to positive results on benchmarks with (a) given structure/characteristics?

TODO

>  - last paragraph of Section 7.1 that gives motivations for MathTSAT and Z3 to be industrial-strenght tool, and states Z3 training, are either not clear or somehow not convincing as motivations. The paragraph should be clarified, or can be omitted.

TODO

>  - Table 16 should not consider, in my opinion, divisions in which there was only one sound solvers, otherwise the concept of "solved uniquely" is trivially true.

TODO

>  * Relation to other competitions
>
>  - in the conclusions section authors mention the changes that will be implemented in 2019: it would be interesting to state which of the changes mentioned are already in place in other related solver's competitions (e.g., SAT, ASP, QBF, SV ...) .

TODO

>  * Minor points
>
>  Introduction
>
>  - "Hence there is" ->  - "Hence there is" ->

WONTFIX (comment is unclear)

>  - there is reference broken at pg. 2

DONE (http:// instead of https:// for www.satcompetition.org)

>  Section 2
>
>  - does the event was co-located with any conference in particular in 2018?

DONE (affiliated with IJCAR in 2018)

>  - "permitted" -> "allowed"

DONE

>  Section 3
>
>  - "increased by" -> "increased of" ?

WONTFIX

>  Section 4
>
>  - "discontinued partly" -> "partly discontinued" ?

DONE (added parentheses for clarity)

>  - "lacking tool support" -> "lacking of tool support" (This recurs later.)

DONE (lacking -> inadequate)

>  - "whose precise" ->  ", whose precise"

DONE

>  - "DT" looks not to be introduced

TODO

>  - the concept of "benchmarks families" is used at the beginning of pg. 7 but it has not been introduced

TODO

>  - "and in 2018, there were" ->   "while in 2018, there were"

DONE

> - e \in \{0,1 \} is defined, but than it is said to be the number of returned statuses...are they aligned? Please clarify

TODO

> Section 6
>
> - "executing" -> "performing"

DONE (executing -> exercising)

> - there is an huge space at pg. 14

TODO

> - Table 9: there are two team members for solver STP that miss affiliation

TODO

> - "Table 11 summarizes these figures". It would be probably useful to explicitly mention the figure numbers

DONE (figures -> numbers)

> Section 7
>
> - "are also taking" -> "were taken"

DONE

> - referring in the text to colors may not be a good choice for b/w printing

TODO

> - In Sec. 7.1 "assume was trained" -> "assume that it was trained"

DONE

> Section 8
>
> - "the fact that in 2015," ->  "the fact that, in 2015,"

DONE

>  - "degression of Z3" ?

DONE

>  - "rank highly" -> "rank in high position" ? (This recurs later.)

DONE

>  - "has relatively few" -> "has few"

DONE

>  - "quotient" -> "ratio" ?

TODO

>  - "a benchmarks" ->  "a benchmark"

DONE

>  - "problem" and "benchmark" are used for the same concept here. Maybe they can be unified

TODO

>  - in Table 17, does 1.0 means they are sequential? In case these numbers could be omitted

WONTFIX (we do not know for sure if those solvers are sequential; and
with these numbers present, the table also indicates whether a solver
was submitted in a given year)

>  - "As in the previous, this" -> "As in the previous analysis, this"

DONE

>  - "Unsurprisingly the" ->   "Unsurprisingly, the"

DONE

>  - "and not Vampire." -> - "and not by Vampire."

DONE

>  Section 9
>
> - "The last FLoC Olympiad was a period of ..." -> are you referring to "the time between the two FLoC Olympiads" ?

WONTFIX (yes; and this is correct usage of the word Olympiad)

>  References
>
>  They could be made more uniform, e.g.
>
>  - [7] vs. [12], where the same journal is referred with contracted form or full name

DONE

>  - Isn't [13] a CADE paper? In case, the way the conference name is mentioned is very different to [14], which is a CADE paper as well

DONE

>  - [38] contains the only acronym for journal name

DONE

> ============================================
>
> *** Review 2 ***
>
> ummary: This is a much improved revision of this paper. I'm very happy to see
> the additional sections and paragraphs, especially Sec 8! I think this paper is
> ready for publication, modulo a few minor (essentially typesetting) issues (see
> below).
>
> Details:
>
> 4.1 Scoring: A good addition. I think it would be great if the changes to the
> scoring system over the years would be analyzed, to see whether they did indeed
> have the intended effect. This has been started in 8.6, but a more
> comprehensive analysis would be great (possibly as a separate submission here
> or elsewhere).

TODO

> (28), Tbl 7: Thanks for the revision of the table, but I'm not sure listing
> more team members to "the best of our knowledge" is a good idea. If I were to
> do that, I would be bound to miss someone... Maybe just list the main
> institution? Or perhaps ask the submitter who they would want to be listed? I
> think there should be enough time to do that.

TODO

> General: Gray and grey are the worst colors to use. On my (Xerox color-laser)
> printer, all of the gray text in tables and figures are almost or completely
> unreadable. For instance, on white (bleached) paper there is not even a faint
> trace of the gray vertical line in Figures 3/4; I had to switch to the screen
> to see what this refers to.

TODO

> Fig 5/6: These could be significantly smaller :)

TODO

> Fig 3/4: Starting the X-axis at an arbitrary point is always a bit fishy. This
> way, any tiny difference can be overemphasized to achieve any level of
> impressiveness that the author choses. For instance, 144k vs 136k instances
> (casually read off the plot) is a 3.5% increase, but this is an irrelevant
> detail as explained in the text. All four series are also of exactly the same
> shape. I think a cactus plot is simply not the right tool for this job.

TODO

> ============================================
>
> *** Review 3 ***
>
>
> This is a revised version of a paper describing
> the years 2015-2018 of the SMT competition.
> The important changes requested last time were addressed,
> and new content was added in Section 8,
> which discusses interesting questions
> and adds substantial value to the paper.
>
> There are only a few minor improvements that should be made before publication:
>
> In the definition of the raw benchmark score (Sect. 4.1),
> it is not clear why CPU time is bounded at 4T
> (the fact that only 4 cores are available only comes later).

TODO

> In the remainder of Sect. 4.1, the term "benchmark family" is used,
> but this is never defined.
> This could be added to the definitions part in the beginning of Sect. 4.

TODO

> In the definition of "Competition-Wide Score (Main Track)",
> it says "since the weight already takes".
> This seems like the weighted division score is used as input for this computation,
> but another sentence of this definition explicitly uses
> the solver's raw score for this computation.
> So what weight is referred to here?

TODO

> In tables 3-5, please keep the regular numbers right-aligned.
> In LaTeX, the easiest way to do this is probably to use separate columns
> for the regular numbers and the number in brackets,
> and potentially shrink the space between each pair of columns.

TODO

> It is said that "StarExec ensures the reproducibility of the competition".
> If we use the terminology of ACM
> (https://www.acm.org/publications/policies/artifact-review-badging),
> the better term would be "replicability" instead of "reproducibility",
> because StarExec allows a different team to replicate the results
> with the same experimental setup.

DONE

> In the next sentence, commas should be placed before and after "e.g.".

DONE

> For tables 7-9, a smaller font size could be used
> (like in Table 12).

TODO

> The first paragraph of Sect. 8 mentions only three aspects,
> but more aspects are analyzed.

TODO

> Figure 5 seems unnecessary, because the difference is not visible anyway.
> Just listing the numbers would be sufficient.

TODO

> Especially in Sect. 8, there are many larger numbers.
> For readability, it would be better if all numbers larger than 9999
> would use a thin space as thousand separator,
> e.g., $1\,000\,000$ or using siunitx which can do this automatically
> with \num{1000000}.
> (Please do not use other characters as thousand separator
> because this can be confusing for international readers.)

TODO

> In Sect. 8.4, it would be better to refer to bounded model checking
> instead of just model checking.
> Other model checking algorithms use many SMT queries
> where the discussed property does not hold.

DONE

> On the middle of page 2, there is an unresolved reference ([?]).

DONE

> Footnote 1 is on the wrong page.

TODO