### 2019-08-29 2018/report/table-winners-unsat-core.tex
Report: Consistent table layout for App and UC.

\begin{table}
  \caption{Best \ucoretrack solvers (by division)}
  \label{tab:winners-unsat-core}
  \centering
  \scalebox{.88}{%
    \renewcommand{\arraystretch}{1.01}%
    \begin{tabular}{r>{\columncolor{white}[.25em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][0.5em]}l}
      \toprule
       Division       & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2016                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2017                              } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2018                          } \\
      \midrule
       ALIA       & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       AUFBVDTLIA & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{                                  } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{                                  } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       AUFLIA     & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       AUFLIRA    & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       AUFNIRA    & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4              } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       BV         & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       LIA        & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{veriT}                       } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4              } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       LRA        & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{veriT}                       } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4              } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       NIA        & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \nonc \cc{cvc4} CVC4 & \nonc \cc{cvc4} CVC4 \nc{Z3}                                      & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       NRA        & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       QF\_ABV    & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices2             } \\
       QF\_ABVFP  & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{                                  } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{---                               } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       QF\_ALIA   & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices2 \nc{Z3}     } \\
       QF\_ANIA   & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4              } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       QF\_AUFBV  & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nc{Z3}                           } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{MathSAT} } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices2 \nc{MathSAT}} \\
       QF\_AUFLIA & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4                    } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4                } \\
       QF\_AUFNIA & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       QF\_AX     & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices2             } \\
       QF\_BV     & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nc{MathSAT}                      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{MathSAT} } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices2             } \\
       QF\_BVFP   & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nc{Z3}                           } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \nc{Z3}                     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       QF\_DT     & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{                                  } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{                                  } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       QF\_FP     & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nc{MathSAT}                      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \nc{Z3}                     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       QF\_IDL    & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 \nc{Z3}            } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 \nc{Z3}        } \\
       QF\_LIA    & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \cc{smti} SMTInterpol & \cc{smti} SMTInterpol \nc{Z3}                                    & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol         } \\
       QF\_LIRA   & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 \nc{Z3}            } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 \nc{Z3}        } \\
       QF\_LRA    & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol             } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol             } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol         } \\
       QF\_NIA    & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       QF\_NIRA   & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       QF\_NRA    & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4              } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       QF\_RDL    & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{veriT}  } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 \nc{Z3}            } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 \nc{Z3}        } \\
       QF\_UF     & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{MathSAT}} & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4                    } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4                } \\
       QF\_UFBV   & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nc{MathSAT}                      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{MathSAT} } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices2 \nc{Z3}     } \\
       QF\_UFIDL  & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices2             } \\
       QF\_UFLIA  & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{MathSAT}} & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3} } \\
       QF\_UFLRA  & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{Z3}     } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol \nc{MathSAT}} & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices2 \nc{MathSAT}} \\
       QF\_UFNIA  & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       QF\_UFNRA  & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       UF         & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4              } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       UFBV       & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       UFDT       & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{                                  } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{                                  } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       UFIDL      & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4              } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       UFLIA      & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4              } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
       UFLRA      & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}      } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 \nc{Z3}  } \\
       UFNIA      & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{ \nc{Z3}                          } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4              } & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4          } \\
      \bottomrule
    \end{tabular}
  }
\end{table}