### 2019-08-29 2018/report/table-winners-main.tex
Report: Update layout of winners main table.

\begin{table}
\caption{Best \maintrack solvers (by division)}
\label{tab:winners-main}
\centering
\scalebox{.82}{%
\renewcommand{\arraystretch}{1.01}%
  \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][0.5em]}c}
    \toprule
    Division         &  2015                        &  2016                     &  2017                           &  2018                     \\
    \midrule
    \wc ABVFP        &                              &                           &                                 & \nonc \cc{cvc4} {CVC4}    \\
    \rc{cvc4}
    \wc ALIA         & CVC4 \nc{Z3}                 & CVC4 \nc{Z3}              & CVC4 \nc{Z3}                    & CVC4 \nc{Z3}              \\
    \wc AUFBVDTLIA   &                              &                           & \nonc \cc{cvc4} {CVC4}          & \nonc \cc{cvc4} {CVC4}    \\
    \wc AUFDTLIA     &                              &                           & \nonc \cc{cvc4} {CVC4}          & \cc{cvc4} {CVC4}          \\
    \rc{cvc4}
    \wc AUFLIA       & {CVC4}                       & {CVC4}                    & {CVC4}                          & {CVC4}                    \\
    \rc{vamp}
    \wc AUFLIRA      & \cc{cvc4} CVC4 \nc{Z3}       & Vampire \nc{Z3}           & Vampire \nc{Z3}                 & \cc{cvc4} CVC4 \nc{Z3}    \\
    \rc{vamp}
    \wc AUFNIRA      & \nonc \cc{cvc4} {CVC4}       & {Vampire}                 & {Vampire}                       & \cc{cvc4} {CVC4}          \\
    \rc{q3b}
    \wc BV           & \nonc \cc{cvc4} CVC4 \nc{Z3} & {Q3B}                     & Q3B \nc{Z3}                     & \cc{cvc4} {CVC4}          \\
    \wc BVFP         &                              &                           &                                 & \nonc \cc{cvc4} {CVC4}    \\
    \wc FP           &                              &                           &                                 & \nonc \cc{cvc4} {CVC4}    \\
    \rc{cvc4}
    \wc LIA          & {CVC4}                       & {CVC4}                    & CVC4 \nc{Z3}                    & CVC4 \nc{Z3}              \\
    \rc{cvc4}
    \wc LRA          & {CVC4}                       & {CVC4}                    & CVC4 \nc{Z3}                    & CVC4 \nc{Z3}              \\
    \rc{cvc4}
    \wc NIA          & \nonc CVC4 \nc{Z3}           & \cc{prob} ProB \nc{Z3}    & CVC4 \nc{Z3}                    & CVC4 \nc{Z3}              \\
    \rc{vamp}
    \wc NRA          & \nonc \cc{cvc4} {CVC4}       & {Vampire}                 & \cc{redl} {Redlog}              & Vampire \nc{Z3} / Vampire \\
    \rc{bool}
    \wc QF\_ABV      & {Boolector}                  & {Boolector}               & {Boolector}                     & {Boolector}               \\
    \wc QF\_ABVFP    &                              &                           & {---}                           & \cc{cvc4} {CVC4}          \\
    \rc{yices}
    \wc QF\_ALIA     & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{cvc4}
    \wc QF\_ANIA     & \nonc CVC4 \nc{Z3}           & \nonc {CVC4}                & \nonc {CVC4}                  & \nonc CVC4 \nc{Z3}        \\
    \rc{cvc4}
    \wc QF\_AUFBV    & CVC4 \nc{MathSAT}            & CVC4 \nc{MathSAT}         & \cc{yices} Yices2 \nc{MathSAT}  & {CVC4}                    \\
    \rc{yices}
    \wc QF\_AUFLIA   & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{cvc4}
    \wc QF\_AUFNIA   & \nonc {CVC4}                 & \nonc {CVC4}              & \nonc CVC4 \nc{Z3}              & \nonc CVC4 \nc{Z3}        \\
    \rc{yices}
    \wc QF\_AX       & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{bool}
    \wc QF\_BV       & {Boolector}                  & {Boolector}               & {Boolector/MinkeyRink}          & {Boolector/MinkeyRink}    \\
    \wc QF\_BVFP     & \nonc \cc{z} {Z3}            & \nonc \wc \nc{Z3}         & \nonc \cc{coli} COLIBRI \nc{Z3} & \cc{cvc4} {CVC4}          \\
    \wc QF\_DT       &                              &                           & \nonc \cc{cvc4} {CVC4}          & \nonc \cc{cvc4} {CVC4}    \\
    \rc{coli}
    \wc QF\_FP       & \nonc \cc{z} {Z3}            & \nonc \wc \nc{MathSAT}    & COLIBRI \nc{Z3}                 & {COLIBRI}                 \\
    \rc{yices}
    \wc QF\_IDL      & Yices2 \nc{Z3}               & Yices2 \nc{Z3}            & {Yices2}                        & {Yices2}                  \\
    \rc{cvc4}
    \wc QF\_LIA      & CVC4 \nc{MathSAT}            & CVC4 \nc{MathSAT}         & CVC4 \nc{MathSAT}               & \cc{spass} {SPASS-SATT}   \\
    \rc{yices}
    \wc QF\_LIRA     & {Yices2}                     & {Yices2}                  & Yices2 \nc{Z3}                  & Yices2 \nc{Z3}            \\
    \rc{cvc4}
    \wc QF\_LRA      &  {CVC4}                      & {CVC4}                    & {CVC4}                          & {CVC4}                    \\
    \rc{cvc4}
    \wc QF\_NIA      & \cc{apr} AProVE \nc{Z3}      & \cc{yices} Yices2 \nc{Z3} & {CVC4}                          & {CVC4}                    \\
    \rc{cvc4}
    \wc QF\_NIRA     & {CVC4}                       & {CVC4}                    & \cc{rat} {SMT-RAT}              & \cc{rat} {SMT-RAT}        \\
    \rc{yices}
    \wc QF\_NRA      & Yices2 \nc{Z3}               & Yices2 \nc{Z3}            & {Yices2}                        & Yices2 \nc{Z3}            \\
    \rc{yices}
    \wc QF\_RDL      & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \wc QF\_SLIA     &                              &                           &                                 & \nonc \cc{cvc4} {CVC4}    \\
    \rc{yices}
    \wc QF\_UF       & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{bool}
    \wc QF\_UFBV     & {Boolector}                  & {Boolector}               & {Boolector}                     & {Boolector}               \\
    \rc{yices}
    \wc QF\_UFIDL    & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFLIA    & Yices2 \nc{Z3}               & Yices2 \nc{Z3}            & {Yices2}                        & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFLRA    & {Yices2}                     & {Yices2}                  & {Yices2}                        & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFNIA    & \nonc \cc{cvc4} {CVC4}       & {Yices2/CVC4}             & {Yices2}                        & {Yices2}                  \\
    \rc{yices}
    \wc QF\_UFNRA    & \nonc \cc{cvc3} CVC3 \nc{Z3} & {Yices2}                  & Yices2 \nc{Z3}                  & {Yices2}                  \\
    \rc{cvc4}
    \wc UF           & {CVC4}                       & {CVC4}                    & \cc{vamp} {Vampire}             &  {CVC4/Vampire}           \\
    \rc{cvc4}
    \wc UFBV         & \nonc CVC4 \nc{Z3}           & CVC4 \nc{Z3}              & \nonc CVC4 \nc{Z3}              & \nonc CVC4 \nc{Z3}        \\
    \wc UFDT         &                              &                           & \nonc \cc{cvc4} {CVC4}          & \cc{cvc4} {CVC4}          \\
    \wc UFDTLIA      &                              &                           & \nonc \cc{vamp} {Vampire}       & \cc{cvc4} {CVC4}          \\
    \rc{cvc4}
    \wc UFIDL        & CVC4 \nc{Z3}                 & {CVC4}                    & {CVC4}                          & CVC4 \nc{Z3}              \\
    \rc{cvc4}
    \wc UFLIA        & {CVC4}                       & {CVC4}                    & {CVC4}                          & {CVC4}                    \\
    \rc{cvc4}
    \wc UFLRA        & \cc{cvc3} {CVC3}             & \cc{vamp} Vampire \nc{Z3} & CVC4 \nc{Z3}                    & CVC4 \nc{Z3}              \\
    \rc{vamp}
    \wc UFNIA        & \nonc \cc{cvc4} {CVC4}       & {Vampire}                 & {Vampire}                       & Vampire \nc{Z3} / Vampire \\
    \bottomrule
  \end{tabular}}
\end{table}

\begin{table}
\caption{Best \maintrack solvers (by division). NEW}
\label{tab:winners-main}
\centering
\scalebox{.82}{%
\renewcommand{\arraystretch}{1.01}%

\begin{tabular}{r>{\columncolor{white}[.25em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][0.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][0.5em]}l}
\toprule
Division &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2015} &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2016} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2017} &                   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2018} \\
\midrule
ABVFP      &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} \\
ALIA       &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
AUFBVDTLIA &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} \\
AUFDTLIA   &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
AUFLIA     &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
AUFLIRA    &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire [Z3]} &      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
AUFNIRA    &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &           \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
BV         &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{q3b} Q3B} &           \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{q3b} Q3B [Z3]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
BVFP       &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} \\
FP         &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} \\
LIA        &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
LRA        &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
NIA        &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{prob} ProB [Z3]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
NRA        &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{redl} Redlog} &                                   \cc{vamp} Vampire [Z3] & \cc{vamp} Vampire \\
QF\_ABV     &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} \\
QF\_ABVFP   &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{---} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_ALIA    &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_ANIA    &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} \\
QF\_AUFBV   &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [MathSAT]} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [MathSAT]} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [MathSAT]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_AUFLIA  &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_AUFNIA  &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} \\
QF\_AX      &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_BV      &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} &                                        \cc{bool} Boolector & \cc{mink} Minkeyrink &                                   \cc{bool} Boolector & \cc{mink} Minkeyrink \\
QF\_BVFP    &                 \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{z} Z3} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc [Z3]} &      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{coli} COLIBRI [Z3]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_DT      &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} \\
  QF\_FP      &                 \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{z} Z3} &                 \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc [MathSAT]} &      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{coli} COLIBRI [Z3]} &      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{coli} COLIBRI} \\
QF\_IDL     &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_LIA     &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [MathSAT]} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [MathSAT]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [MathSAT]} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{spass} SPASS-SATT} \\
QF\_LIRA    &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} \\
QF\_LRA     &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_NIA     &      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{apr} AProVE [Z3]} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_NIRA    &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{rat} SMT-RAT} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{rat} SMT-RAT} \\
QF\_NRA     &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} \\
QF\_RDL     &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_SLIA    &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_UF      &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_UFBV    &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} \\
QF\_UFIDL   &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_UFLIA   &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_UFLRA   &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_UFNIA   &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &                                               \cc{yices} Yices & \cc{cvc4} CVC4 &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_UFNRA   &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc3} CVC3 [Z3]} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
UF         &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &           \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &                                           \cc{cvc4} CVC4 & \cc{vamp} Vampire \\
UFBV       &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} \\
UFDT       &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
UFDTLIA    &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &           \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{vamp} Vampire} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
UFIDL      &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
UFLIA      &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
UFLRA      &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc3} CVC3} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire [Z3]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
UFNIA      &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &           \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &                                   \cc{vamp} Vampire [Z3] & \cc{vamp} Vampire \\
\bottomrule
\end{tabular}
}

\end{table}