### 2019-08-29 2018/report/table-winners-main-sat-unsat.tex
Report: Fix results and layout for sat/unsat score table.

\begin{table}
  \caption{Alternative best solvers in the \maintrack if only \emph{satisfiable} benchmarks had been considered. These results are in reference to those given in Table~\ref{tab:winners-main}.}
  \label{tab:results:sat}
\centering
  \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][0.5em]}c}
    \toprule
    Division         &  2015                &  2016                   &  2017                   &  2018                     \\
    \hline \hline
    ALIA             &                      &                         &                         & \cc{alt} Alt-Ergo         \\
    AUFLIA           & \cc{cvc3} CVC3       &                         &                         & \cc{alt} Alt-Ergo         \\
    AUFLIRA          &                      & \cc{cvc4} CVC4          & \cc{cvc4} CVC4          & \cc{alt} Alt-Ergo         \\
    AUFNIRA          & \nonc\nc{Z3}         & \cc{cvc4} CVC4 \nc{Z3}  & \cc{cvc4} CVC4 \nc{Z3}  & \nc{Z3}                   \\
    BV               &                      & \nc{Z3}                 & \cc{bool} Boolector     & \cc{bool} Boolector       \\
    LRA              &                      &                         & \cc{verit} veriT        &                           \\
    NIA              & \nonc \cc{cvc3} CVC3 &                         &                         &                           \\
    NRA              & \nonc \cc{cvc3} CVC3 & \cc{cvc4} CVC4          & \cc{cvc4} CVC4          & \cc{cvc4} CVC4 \nc{Z3}    \\
    QF\_ABV          &                      & \cc{yices} Yices2       & \cc{yices} Yices2       &                           \\
    QF\_AUFBV        & \cc{yices} Yices     & \cc{yices} Yices2       &                         & \cc{yices} Yices2         \\
    QF\_BV           &                      & \nc{Z3}/MinkeyRink      &                         &                           \\
    QF\_FP           &                      &                         &                         & \cc{cvc4} CVC4            \\
    QF\_LIA          & \nc{Z3}              &                         &                         &                           \\
    QF\_LRA          &                      &                         & \cc{yices} Yices2       & \cc{yices} Yices2         \\
    QF\_NIA          &                      & \cc{rat} SMT-RAT        &                         &                           \\
    QF\_NIRA         &                      &                         & \cc{cvc4} CVC4          &                           \\
    QF\_NRA          &                      &                         & \nc{Z3}                 &                           \\
    QF\_UFBV         &                      &                         &                         & \cc{yices} Yices2         \\
    QF\_UFIDL        & \nc{Z3}              & \nc{Z3}                 & \nc{Z3}                 & \cc{verit} veriT  \nc{Z3} \\
    QF\_UFNIA        & \nonc \cc{cvc3} CVC3 &                         & \cc{cvc4} CVC4          &                           \\
    UF               &                      & \cc{vamp} Vampire       &                         & \cc{vamp}Vampire/-        \\
    UFBV             &                      & \cc{bool} Boolector     &                         &                           \\
    UFDTLIA          &                      &                         & \nonc \cc{cvc4} CVC4    &                           \\
    UFLIA            & \nc{Z3}              &                         &                         &                           \\
    UFLRA            &                      & \cc{cvc4} CVC4          &                         &                           \\
    UFNIA            &                      & \cc{cvc4} CVC4          & \cc{cvc4} CVC4          & \cc{cvc4} CVC4            \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}
  \caption{Alternative best solvers in the \maintrack if only \emph{satisfiable} benchmarks had been considered. These results are in reference to those given in Table~\ref{tab:winners-main}. NEW}
  \label{tab:results:sat}
  \centering
  \scalebox{.82}{%
    \renewcommand{\arraystretch}{1.01}%
\begin{tabular}{r>{\columncolor{white}[.25em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][0.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][0.5em]}l}
\toprule
Division &                 \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2015} &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2016} &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2017} &                   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2018} \\
\midrule
ALIA      &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{[Z3]} \\
AUFLIRA   &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{[Z3]} &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{[Z3]} &                   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{[Z3]} \\
AUFNIRA   &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
BV        &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{q3b} Q3B [Z3]} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector [Z3]} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} \\
NIA       &           \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc [Z3]} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
NRA       &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{[Z3]} \\
QF\_ABV    &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_AUFBV  &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_AUFNIA &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} \\
QF\_BV     &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                 \cc{bool} Boolector [Z3] & \cc{mink} Minkeyrink &                                                         \cc{mink} Minkeyrink &  &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_FP     &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_IDL    &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_LIRA   &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_LRA    &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_NIA    &      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{apr} AProVE} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{rat} SMT-RAT [Z3]} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_NRA    &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_UFBV   &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_UFIDL  &                    \hspace*{3em}                   & \cc{yices} Yices [Z3] &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &                                                         & \cc{yices} Yices [Z3] &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{verit} veriT [Z3]} \\
QF\_UFLIA  &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
UF        &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                                         \cc{vamp} Vampire & \hspace*{3em} \\
UFBV      &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector [Z3]} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
UFLIA     &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
UFLRA     &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{[Z3]} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
UFNIA     &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
\bottomrule
\end{tabular}
  }
\end{table}

\begin{table}
  \caption{Alternative best solvers in the \maintrack if only \emph{unsatisfiable} benchmarks had been considered. These results are in reference to those given in Table~\ref{tab:winners-main}.}
  \label{tab:results:unsat}
  \centering
  \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][0.5em]}c}
\toprule
Division        &  2015                  &  2016                     &  2017                 &  2018                   \\
\hline \hline
    BV          &                        & \cc{cvc4} CVC4            &                       &                         \\
    NIA         &                        & \cc{cvc4} CVC4            & \cc{vamp} Vampire     & \cc{vamp} Vampire       \\
    QF\_AUFBV   &                        &                           & \cc{cvc4} CVC4        &                         \\
    QF\_AUFNIA  & \nonc \nc{Z3}          & \nonc \nc{Z3}             &                       &                         \\
    QF\_BV      &                        &                           & -/Boolector           &                         \\
    QF\_IDL     & \cc{cvc4} CVC4         &                           & \nc{Z3}               & \cc{cvc4} CVC4 \nc{Z3}  \\
    QF\_LIA     & \cc{smti} SMTInterpol  & \cc{smti} SMTInterpol     & \cc{smti} SMTInterpol & \cc{smti} SMTInterpol   \\
    QF\_NIA     & \cc{cvc3} CVC3         &                           & \cc{yices} Yices2     &                         \\
    QF\_NRA     &                        &                           &                       & \cc{cvc4} CVC4          \\
    QF\_UFLIA   &                        &                           & \nc{Z3}               &                         \\
    QF\_UFNRA   &                        & \cc{cvc4} CVC4 \nc{Z3}    & \cc{cvc4} CVC4        & \cc{cvc4} CVC4 \nc{Z3}  \\
    UF          &                        &                           & \cc{cvc4} CVC4        & -/CVC4                  \\
    UFIDL       &                        & \cc{vamp} Vampire \nc{Z3} & \nc{Z3}               &                         \\
    UFLRA       &                        &                           & \cc{verit} veriT      & \cc{verit} veriT        \\
    \bottomrule
  \end{tabular}

\end{table}

\begin{table}
  \caption{Alternative best solvers in the \maintrack if only \emph{unsatisfiable} benchmarks had been considered. These results are in reference to those given in Table~\ref{tab:winners-main}. NEW}
  \label{tab:results:unsat}
  \centering
  \scalebox{.82}{%
    \renewcommand{\arraystretch}{1.01}%
\begin{tabular}{r>{\columncolor{white}[.25em][.5em]}c|>{\columncolor{white}[.5em][.5em]}c>{\columncolor{white}[.5em][.5em]}c|>{\columncolor{white}[.5em][.5em]}c>{\columncolor{white}[.5em][.5em]}c|>{\columncolor{white}[.5em][.5em]}c>{\columncolor{white}[.5em][.5em]}c|>{\columncolor{white}[.5em][0.5em]}c}
\toprule
Division &                   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2015} &                    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2016} &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2017} &                    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2018} \\
\midrule
AUFLIRA   &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
BV        &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &              \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{q3b} Q3B} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
LIA       &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
NIA       &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire [Z3]} \\
QF\_AUFBV  &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [MathSAT]} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_AUFNIA &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_BV     &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                  \hspace*{3em}                            & \cc{bool} Boolector &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_FP     &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{coli} COLIBRI} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_IDL    &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
QF\_LIA    &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol} &   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol} &   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{smti} SMTInterpol} \\
QF\_NIA    &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc3} CVC3 [Z3]} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_NRA    &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_UFLIA  &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_UFNRA  &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &                                                           & \cc{cvc4} CVC4 [Z3] &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
UF        &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &                          \hspace*{3em}                       & \cc{cvc4} CVC4 \\
UFBV      &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
UFIDL     &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire [Z3]} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
UFLRA     &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{verit} veriT} &        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{verit} veriT} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{verit} veriT} &        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{verit} veriT} \\
\bottomrule
\end{tabular}

  }
\end{table}