### 2019-08-30 2018/report/table-winners-main-families.tex
Report: Fix results and layout for families table.

\begin{table}
  \caption{Alternative best solvers in the \maintrack if families were not considered in computing scores (i.e., the 2015 scoring system were used)}
  \label{tab:results:total:solved}
  \centering
  \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][0.5em]}c}
    \toprule
    Division & 2015 &  2016           &  2017                &  2018                  \\
    \hline\hline
    AUFNIRA  &      & \cc{cvc4} CVC4  &                      &                        \\
    UFIDL    &      & \nc{Z3}         &                      &                        \\
    BV       &      &                 & \cc{bool} Boolector  &                        \\
    QF\_LRA  &      &                 & \cc{yices} Yices2    & \cc{spass} SPASS-SATT  \\
    QF\_NIA  &      &                 & \cc{yices} Yices2    & \cc{yices} Yices2      \\
    QF\_BV   &      &                 &                      & -/Boolector            \\
    QF\_FP   &      &                 &                      & \cc{cvc4} CVC4         \\
  \bottomrule
  \end{tabular}
\end{table}

\begin{table}
  \caption{Alternative best solvers in the \maintrack if families were not considered in computing scores (i.e., the 2015 scoring system were used). NEW}
  \label{tab:results:total:solved}
  \centering
  {%
    \renewcommand{\arraystretch}{1.01}%

\begin{tabular}{r>{\columncolor{white}[.25em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][0.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][0.5em]}l}
\toprule
Division & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2015}              &                 \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2016} &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2017} &                   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2018} \\
\midrule
AUFNIRA  &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\hspace*{5em}} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
BV       &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{}              &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector [Z3]} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_BV    &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{}             &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                     \cc{mink} Minkeyrink &  \hspace*{3em}       &                           \hspace*{3em}                & \cc{bool} Boolector \\
QF\_FP    &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{}             &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_IDL   &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{}             &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_LRA   &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{}             &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{spass} SPASS-SATT} \\
QF\_NIA   &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{}             &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_NRA   &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{}             &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
UFIDL    &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{}              &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &                          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
\bottomrule
\end{tabular}
  }
\end{table}