### 2019-08-30 2018/report/table-winners-main-unknown.tex
Report: Fix results and layout for no-unknowns table.

\begin{table}
  \caption{Alternative best solvers in the \maintrack if benchmarks with unknown status had not bee included in 2017 onwards.}
  \label{tab:results:unknowns}
  \centering
    \begin{tabular}{r@{\hskip 1em}>{\columncolor{white}[.25em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][.5em]}c@{\hskip 1em}>{\columncolor{white}[.5em][0.5em]}c}
    \toprule
    Division &  2017          &  2018                  \\
    \hline\hline
    AUFNIRA  & \cc{cvc4} CVC4 &                        \\
    QF\_NIRA & \nc{Z3}        & \cc{cvc4} CVC4 \nc{Z3} \\
    QF\_NRA  & \nc{Z3}        &                        \\
    UFDTLIA  & \cc{cvc4} CVC4 &                        \\
    BV       &                & \cc{bool} Boolector    \\
    QF\_FP   &                & \cc{cvc4} CVC4 \nc{Z3} \\
    UFNIA    &                & \cc{cvc4} CVC4         \\
    NRA      &                & -/\nc{Z3}              \\
    UF       &                & -/CVC4                 \\
  \bottomrule
  \end{tabular}
\end{table}

\begin{table}
  \caption{Alternative best solvers in the \maintrack if benchmarks with unknown status had not bee included in 2017 onwards. NEW}
  \label{tab:results:unknowns}
  \centering
  {%
    \renewcommand{\arraystretch}{1.01}%
\begin{tabular}{r>{\columncolor{white}[.25em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][0.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][0.5em]}l}
\toprule
Division & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2015} & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2016} &                   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2017} &                 \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2018} \\
\midrule
AUFLIRA  &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                 \cc{cvc4} CVC4 [Z3] &  \hspace*{3em}         &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
AUFNIRA  &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
BV       &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &           \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{q3b} Q3B} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} \\
LRA      &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
NRA      &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                                   & \cc{vamp} Vampire [Z3] \\
QF\_FP    &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_NRA   &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
UF       &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                                            \cc{cvc4} CVC4 &  &                                                           & \cc{cvc4} CVC4 \\
UFNIA    &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                                     \cc{cvc4} CVC4 [Z3] &  \\
\bottomrule
\end{tabular}
  }
\end{table}