### 2019-08-30 2018/report/table-winners-main-24s.tex
Minor

\begin{table}
  \caption{Alternative best solvers in the \maintrack if the time limit had been 24 seconds.}
  \label{tab:results:24s}
  \centering
  \scalebox{.755}{%
    \renewcommand{\arraystretch}{1.01}%
\begin{tabular}{r>{\columncolor{white}[.25em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l}
\toprule
Division &                 \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2015} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2016} &                        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2017} &                   \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2018} \\
\midrule
AUFLIA   &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire [Z3]} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{verit} veriT} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
AUFLIRA  &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc3} CVC3 [Z3]} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                                     & \cc{alt} Alt-Ergo [Z3] \\
AUFNIRA  &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc3} CVC3 [Z3]} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
BV       &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                               \cc{cvc4} CVC4 [Z3] & \cc{bool} Boolector [Z3] \\
NRA      &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
QF\_ABV   &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_ANIA  &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \nonc \cc{cvc4} CVC4 [Z3]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_BV    &              \hspace*{3em}                                & \cc{cvc4} CVC4 &           \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{stp} STP [Z3]} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                                        & \cc{bool} Boolector \\
QF\_IDL   &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_LIA   &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [MathSAT]} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [MathSAT]} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_LIRA  &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_LRA   &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_NIA   &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} \\
QF\_NIRA  &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} \\
QF\_NRA   &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_UF    &     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{verit} veriT} &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{verit} veriT} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_UFLIA &                                                         & \cc{yices} Yices &            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
QF\_UFNRA &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} \\
UF       &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &           \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                                         \cc{vamp} Vampire &  \\
UFBV     &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector [Z3]} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
UFIDL    &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &         \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} &                       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
UFLIA    &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &    \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4 [Z3]} \\
UFLRA    &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &           \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{vamp} Vampire} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{verit} veriT [Z3]} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{verit} veriT} \\
UFNIA    &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\nonc \cc{cvc4} CVC4 [Z3]} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                            \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                 \cc{cvc4} CVC4 [Z3] & \cc{vamp} Vampire [Z3] \\
\bottomrule
\end{tabular}
  }
\end{table}