### 2019-08-31 2018/report/SMT-COMP.tex
Source file cleanup

\documentclass[dvipsnames,table,twoside,11pt]{article}

\usepackage{jsat}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{siunitx}

\usepackage{pifont}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{subcaption}

\usepackage{listings}
\lstdefinelanguage{smtlib2}
  {alsoletter={-, =},
   morekeywords={set-logic,declare-fun,assert,check-sat,set-info,exit},
   sensitive=false,
   morecomment=[l]{;}}
\lstset{language=smtlib2,
basicstyle=\sffamily,
numbers=none,
keywordstyle=\bfseries\sffamily,
keywordstyle={[2]\bfseries\sffamily\underbar}
}

%\input epsf
\special{papersize=8.5in,11in}

\jsatheading{?}{2019}{??-??}
\ShortHeadings{The SMT Competition 2015--2018}
{Weber et al.}
\firstpageno{1}

% macros for tracks for more consistency
\newcommand{\maintrack}{Main Track\xspace}
\newcommand{\apptrack}{Application Track\xspace}
\newcommand{\ucoretrack}{Unsat-Core Track\xspace}

% float layout
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\colorlet{bool}{blue!10}
\colorlet{cvc4}{cyan!10}
\colorlet{smti}{yellow!10}
\colorlet{yices}{green!10}
\colorlet{vamp}{orange!10}
\colorlet{q3b}{red!10}
\colorlet{prob}{gray!10}
\colorlet{z}{Fuchsia!10}
\colorlet{coli}{Brown!10}
\colorlet{spass}{Lavender!10}
\colorlet{apr}{PineGreen!10}
\colorlet{rat}{Goldenrod!10}
\colorlet{cvc3}{Melon!10}
\colorlet{redl}{Yellow!10}
\colorlet{alt}{LimeGreen!10}
\colorlet{verit}{MidnightBlue!10}
\colorlet{stp}{Fuchsia!10}
\colorlet{mink}{Magenta!10}
\colorlet{nonc}{black!55}

\newcommand{\cc}[1]{\cellcolor{#1}}
\newcommand{\wc}{\cc{white}}
\newcommand{\rc}[1]{\rowcolor{#1}}
\newcommand{\nc}[1]{{[}#1{]}}

\newcommand{\nonc}{\color{nonc}}

\newcommand{\gray}[1]{\textcolor{gray}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{The SMT Competition 2015--2018}

\author{%
  \name{Tjark Weber {\normalfont (competition chair, 2015--2018)}}
  \email{tjark.weber@it.uu.se} \\
  \addr Uppsala University \\
  Sweden
  \AND
  \name{Sylvain Conchon {\normalfont (co-organizer, 2015--2016)}}
  \email{Sylvain.Conchon@lri.fr} \\
  \addr Paris-Sud University \\
  France
  \AND
  \name{David D\'{e}harbe {\normalfont (co-organizer, 2015--2016)}}
  \email{david.deharbe@clearsy.com} \\
  \addr ClearSy Systems Engineering \\
  France
  \AND
  \name{Matthias Heizmann {\normalfont (co-organizer, 2016--2018)}}
  \email{heizmann@informatik.uni-freiburg.de} \\
  \addr University of Freiburg \\
  Germany
  \AND
  \name{Aina Niemetz {\normalfont (co-organizer, 2018)}}
  \email{niemetz@cs.stanford.edu} \\
  \addr Stanford University\\
  USA
  \AND
  \name{Giles Reger {\normalfont (co-organizer, 2017--2018)}}
  \email{giles.reger@manchester.ac.uk} \\
  \addr University of Manchester \\
  UK}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
  The International Satisfiability Modulo Theories Competition is an
  annual competition between Satisfiability Modulo Theories~(SMT)
  solvers.  The 2018 edition of the competition was part of the FLoC
  Olympic Games, which comprised 14 competitions in various areas of
  computational logic.  We report on the design and selected results
  of the SMT Competition during the last FLoC Olympiad, from 2015 to
  2018.  These competitions set several new records regarding the
  number of participants, number of benchmarks used, and amount of
  computation performed.
\end{abstract}

\keywords{SMT solver, SMT-COMP, SMT-LIB, Satisfiability Modulo Theories, competitions}

\published{November 2018}{March 2019}{TBA}

%% Expected content according to the JSAT Call for Papers:

%% Full articles by competition organizers reporting on SAT 2018
%% competitions and evaluation. The articles should describe the
%% competition, its criteria, why it is interesting to the SAT
%% research community, execution environment used, analysis of the
%% results (including how they compare to previous instantiations, if
%% appropriate), and give a summary of the main technical
%% contributions to the field, as well as discussions on lessons
%% learned and suggestions for improvements for future competitions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:introduction}

Satisfiability Modulo Theories (SMT) is a generalization of Boolean
satisfiability (SAT), the satisfiability decision problem for
propositional logic.  In place of Boolean variables, SMT formulas may
contain terms that are built from function and predicate symbols drawn
from a number of background theories.  Background theories, which are
motivated by application domains, include the theory of arrays,
integer and real arithmetic, bit-vectors, and floating-point numbers,
among others~\cite{BarFT-RR-17}.  For instance, the following is an
SMT formula over the combination of integer arithmetic and
uninterpreted functions:
%
$$x \leq y \wedge y \leq x \wedge P (f(x) - f(y)) \wedge \neg P(0).$$
%
This formula asserts that~$x$ is less than or equal to~$y$, that~$y$
is less than or equal to~$x$, that~$P$ holds for the difference
of~$f(x)$ and~$f(y)$, and that~$P$ does not hold for~$0$.  (Here,~$x$
and~$y$ are integers,~$f$ is a function from integers to integers,
and~$P$ is a predicate over integers.)
%
Software tools to determine the satisfiability of such formulas are
called SMT solvers.  With its rich input language, SMT has
applications in software engineering, optimization, and many other
areas~\cite{DeMoura:2011:SMT}.

Internally, many SMT solvers employ SAT solving techniques to deal
with the propositional structure of the formula, and combine these
with (semi-)decision procedures for the background theories that the
solver
supports~\cite{DBLP:conf/cav/BarrettCDHJKRT11,DBLP:conf/cade/BoutonODF09,DBLP:conf/tacas/MouraB08,Nieuwenhuis:2006:SSS:1217856.1217859}.
Historically, SMT solvers have focused on quantifier-free formulas and
decidable combinations of background theories.  Increasingly, however,
SMT solvers also support quantified formulas.  Hence there is overlap
between SMT solving and automated theorem proving for quantified
Boolean formulas~(QBF), first-order logic, and even higher-order
logic~\cite{DBLP:journals/corr/abs-1712-01486}.

The International Satisfiability Modulo Theories Competition
(SMT-COMP)~\cite{BdMS05,BdMS07,BDdMOS13,BDOS08,BDOS10,CDW14,CGBD12} is
an annual competition between SMT solvers.  It was instituted in~2005,
and is affiliated with the International Workshop on Satisfiability
Modulo Theories.  Solvers are submitted to the competition by their
developers, and pitted against each other in a number of tracks and
divisions.

The SMT Competition is one of a growing number of tool competitions in
the broad area of formal methods, which are summarised in a recent
review of the area~\cite{toolympics}. The two closest competitions in
terms of goals and design are the SAT competition
(\url{http://www.satcompetition.org/}), which evaluates SAT solvers,
and CASC (\url{http://www.tptp.org/CASC/}), which evaluates automated
theorem provers (ATPs). In recent years, we have seen a blurring of
lines between SMT-COMP and CASC with SMT solvers competing in CASC and
ATPs competing in SMT-COMP. The main (informal) distinction remains
that SMT-COMP focuses on problems that are heavy in theories, whereas
CASC focuses on problems heavy in quantifiers.  Another closely
related competition is SV-COMP (\url{https://sv-comp.sosy-lab.org})
where many tools rely on advances in SMT solving technology.

The SMT Competition was part of the Federated Logic Conference (FLoC)
Olympic Games, which comprised 14 competitions in various areas of
computational logic, in Vienna in~2014, and again in Oxford~(UK)
in~2018.  It was last described in a 2014 competition
report~\cite{CDW14}.  In this paper, we report on the design and
selected results of the competition during the last FLoC Olympiad,
from 2015 to~2018.

Complete data about past competitions since 2015 (and partly for
earlier competitions) is available from the competition website,
\url{http://www.smtcomp.org}.  This includes detailed results for each
solver/benchmark combination, totaling nearly 6 million data records
(about \SI{1.1}{GB} in CSV format).  Competition solvers and
benchmarks may be downloaded from StarExec,
\url{https://www.starexec.org}.

The rest of this paper is structured as follows.  In
Section~\ref{sec:goals}, we present the goals and organization of the
competition.  Section~\ref{sec:smtlib} introduces the SMT-LIB language
and library, and Section~\ref{sec:benchmarks} discusses competition
divisions and benchmarks.  The competition is being run on StarExec,
which is described in Section~\ref{sec:starexec}.
Section~\ref{sec:participants} gives an overview of participating
solvers, and Section~\ref{sec:scoring} discusses the scoring rules.
Selected competition results are presented in
Section~\ref{sec:results}, and analyzed in more detail in
Section~\ref{sec:analysis}.  We conclude with suggestions for future
SMT Competitions in Section~\ref{sec:conclusions}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The Competition Goals and Organization}
\label{sec:goals}

The original goals of the SMT Competition were to spur adoption of the
community-designed SMT-LIB format (see Section~\ref{sec:smtlib}), and
to spark further advances in SMT, especially for
verification~\cite{BdMS05}.  These, together with providing a useful
yardstick of performance for users and developers of SMT solvers, are
still its main goals to date.  The competition has been successful in
establishing SMT-LIB as the de~facto standard language for SMT
solvers.  In recent years, the focus in this regard has shifted
towards promoting some of the newer extensions of the SMT-LIB
language~\cite{BarFT-RR-17}, such as floating-point numbers and
algebraic datatypes.

The competition is organized under the direction of the SMT Steering
Committee, which appoints the competition chair~\cite{smt-bylaws}.
The competition chair then assembles a team of organizers, and
oversees the work for the next edition of the competition.
%
The SMT Workshop includes a block of time to present the competitors
and results of the competition.  The workshop is affiliated with a
major conference in the field of automated reasoning each year: with
CAV in 2015, with IJCAR in 2016, again with CAV in 2017, and with
IJCAR (as part of FLoC) in~2018.  Consequently, its date varies.
During 2015--2018, the workshop always took place in July.

Important competition deadlines are determined by calculating
backwards from the date of the next SMT Workshop.  The competition's
computational workload should be completed approximately two weeks
before the workshop date.  This gives participants and other parties a
chance to scrutinize the data for errors, allows partial re-runs of
competition jobs if necessary, and gives the organizers time to
prepare the results presentation.  While making all data publicly
available well in advance of the official results presentation may
impact the suspense, it has on several occasions helped to uncover
serious problems with the competition tools or the StarExec framework
(see Section~\ref{sec:starexec}) that could have led to invalid
results otherwise.

It typically takes up to three weeks to run all competition jobs on
the StarExec cluster, so the final submission deadline for solvers
needs to be about five weeks before the workshop date.  The
competition imposes a separate deadline for first versions of solvers
about two weeks before the final submission deadline.  This earlier
deadline has proved useful for the organizers to obtain an accurate
estimate on the number of competing solvers, and to run preliminary
tests with some of the submitted solvers to identify potential issues.
Participants can still make changes to their solver until the final
deadline but are encouraged to use this period for bug-fixing only.

Around the initial solver deadline, the organizers also aim to publish
the latest version of the competition tools, and---in collaboration
with the SMT-LIB maintainers---to release a new version of the SMT
Library (see Section~\ref{sec:smtlib}).  SMT-LIB releases were made on
June~1 in 2015, on May~23 in 2016, on June~5 in 2017, and on May~20 in
2018.  Anyone may submit new benchmarks to the SMT Library, and
competitors are allowed to tune their solvers accordingly.  Therefore
all benchmarks eligible for the competition must be released at least
some time before the final solver submission deadline.  To give the
SMT-LIB maintainers sufficient time for curation, the deadline for new
benchmark contributions has usually been about four to six weeks
before the initial solver deadline, typically in April or early May.

The competition rules are revised each year, and a (near final) draft
version of the rules is made public on the competition website around
mid-April.  The competition chair has ultimate responsibility for the
rules, but changes are often preceded by discussions among the
organizers or within the SMT community.  Such discussions are
initiated by a Call for Comments that is sent out to the SMT-COMP
mailing list~\cite{smtcomp-mailinglist} in February or early March.
Publicly, this call is the first harbinger of a new edition of the SMT
Competition.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The SMT-LIB Language and Library}
\label{sec:smtlib}

All problems used in the SMT Competition come from the SMT Library
(SMT-LIB).  This is a large collection of benchmark problems from
various sources.  The SMT Library is available from a public
repository~\cite{smtlib-repository}, and each new release is mirrored
on StarExec.  Users of SMT solvers are encouraged to submit new and
interesting benchmarks to the library, which has grown from \num{1352}
benchmarks in~2005~\cite{BdMS05} to \num{347011} benchmarks in 2018.

Benchmarks in the SMT Library are written in the SMT-LIB
language~\cite{BarFT-RR-17}, a text-based format that defines the
syntax and semantics of solver input and output.  This includes a
language for terms and formulas, and a command language for
interacting with SMT solvers.  For instance,
Figure~\ref{fig:smtlib-example} shows the example formula from
Section~\ref{sec:introduction} as a benchmark in SMT-LIB syntax, with
commands to the SMT solver typeset in bold.

In~2016, benchmarks were updated from version~2.0 to version~2.5 of
the SMT-LIB language.  Since~2017, benchmarks are written in
version~2.6.  Fortunately, versions~2.5 and~2.6 are largely
backwards-compatible to 2.0, so that older solvers are still able to
compete with at most minor modifications.

\begin{figure}
\begin{lstlisting}
(set-info :smt-lib-version 2.6)
(set-logic QF_UFLIA)
(set-info :status unsat)
(declare-fun x () Int)
(declare-fun y () Int)
(declare-fun f (Int) Int)
(declare-fun P (Int) Bool)
(assert (and (<= x y) (<= y x) (P (- (f x) (f y))) (not (P 0))))
(check-sat)
(exit)
\end{lstlisting}
\caption{A benchmark problem in SMT-LIB syntax.}
\label{fig:smtlib-example}
\end{figure}

The SMT Library classifies benchmarks according to their \emph{logic},
that is, according to the specific combination of background theories
that the benchmark uses.  Logic names are composed of letter groups
that refer to these background theories: {A} or {AX} for the theory of
arrays, {BV} for the theory of bit-vectors, {DT} for algebraic
datatypes, {FP} for the floating-point theory, {IA} for integer
arithmetic, {RA} for real arithmetic, {IRA} for mixed integer and real
arithmetic, {S} for the theory of strings, {UF} for uninterpreted
functions (i.e., free symbols).  Additionally, logics may impose
further syntactic restrictions, such as the absence of quantifiers
({QF\_}), the restriction to difference logic over the
integers~({IDL}) or reals ({RDL}), or to the (non-)linear fragment of
arithmetic ({N} or {L}, respectively, before {IA}, {RA} or {IRA}).
For instance, the logic QF\_UFLIA contains quantifier-free formulas
with uninterpreted functions over linear integer arithmetic.  These
restrictions are typically motivated by the existence of efficient
decision procedures for certain syntactic fragments.

Additionally, the SMT Library distinguishes between \emph{incremental}
and \emph{non-incremental} benchmarks.  Non-incremental benchmarks
contain a single {\tt check-sat} command, which instructs the SMT
solver to determine the satisfiability of the benchmark.  (Valid
solver responses are {\tt sat}, {\tt unsat}, and {\tt unknown}.)
Incremental benchmarks exercise additional solver features, such as
the ability to retract assertions, and typically contain multiple {\tt
  check-sat} commands.  These benchmarks originate from applications
such as bounded model
checking~\cite{Gunther:2014:IBS:2632362.2632374}, where SMT solvers
interact with other tools in a feedback loop.

The 2018 release of the SMT Library contained \num{336844}
non-incremental benchmarks in 51 logics, as well as~\num{10167}
incremental benchmarks (with a total of nearly 32 million {\tt
  check-sat} commands) in 26 logics.  Table~\ref{table:smtlib}
presents further data on the size of the SMT Library since~2015.
While the library usually grows from one year to the other because of
the inclusion of new benchmarks and logics, continuous curation
efforts may also cause benchmarks to be removed or reclassified on
occasion, thereby causing individual logics or even the entire library
to shrink.  From~2015 to~2018, the total number of benchmarks has
increased by~68\,\%.  The largest benchmark in the 2018 release had a
size of \SI{1041}{MB}; the total size of the release was \SI{91}{GB}.

\begin{table}
  \caption{Number of benchmarks and logics in the SMT Library}
  \label{table:smtlib}
  \centering
  \begin{tabular}{r@{\ \ }lrrrr}
    \toprule
                               & & \multicolumn{1}{c}{2015} & \multicolumn{1}{c}{2016} & \multicolumn{1}{c}{2017} & \multicolumn{1}{c}{2018} \\
    \midrule
    \multirow{2}{*}{Non-incremental $\{$} & benchmarks & 196375 & 196114 & 258079 & 336844 \\
                                          & logics     &     40 &     40 &     49 &     51 \\
    \multirow{2}{*}{Incremental $\{$}     & benchmarks &  10019 &  10024 &   6262 &  10167 \\
                                          & logics     &     15 &     15 &     22 &     26 \\
    \bottomrule
  \end{tabular}
\end{table}

Benchmarks in the SMT Library are annotated with additional
information, such as the source of the benchmark and its status (i.e.,
whether the benchmark is satisfiable or unsatisfiable) if this is
known.  (For incremental benchmarks, each {\tt check-sat} command has
a separate status annotation.)  To prevent solvers from simply looking
up the benchmark's status in the SMT Library, benchmarks in the
competition are lightly scrambled before they are presented to
solvers~\cite{DBLP:conf/cade/Weber16}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Competition Divisions and Benchmarks}
\label{sec:benchmarks}

The SMT Competition consists of several tracks.  Each track is further
split into (logic) \emph{divisions}, where each division corresponds
to a logic from the SMT Library.  As indicated in
Tables~\ref{table:smtlib}--\ref{table:divisions}, there are slightly
fewer divisions (Table~\ref{table:divisions}) than logics in the SMT
Library (Table~\ref{table:smtlib}) because not every logic contains
benchmarks eligible for the competition.

\paragraph{Tracks.}
The \emph{\maintrack}, which has been a staple of the competition
since its inception in 2005, features non-incremental benchmarks.  The
\emph{\apptrack}, which has been part of the competition since
2011~\cite{BDdMOS13}, features incremental benchmarks.  These
benchmarks are passed to each solver one command at a time by a trace
executor, which waits for the solver's response before passing the
next benchmark command (to prevent look-ahead behavior).  The
\emph{\ucoretrack} uses \maintrack benchmarks that are known to be
unsatisfiable, and solvers are evaluated on their ability to find
small unsatisfiable cores (i.e., a small but still unsatisfiable
subset of the benchmark's assertions).  The \ucoretrack had first been
part of the competition in 2012~\cite{CGBD12}, but was then
discontinued (partly because of the competition's move to StarExec and
inadequate tool support).  It was re-introduced as an experimental
track in 2016, and has been a regular competition track again
since~2017.

\paragraph{Divisions.}
Table~\ref{table:divisions} shows the number of divisions in each
track since~2015.  Participants can choose which tracks and divisions
they want to enter, and competition results are reported separately
for each track and division.  Thus, competing solvers do not need to
support all logics, and may or may not support features such as
incremental solving or unsat-core generation.

\paragraph{Eligible benchmarks.}
Some benchmarks typically have to be excluded for syntactic reasons:
in 2015 and 2016, \num{11966} benchmarks were excluded because they
contained partial or underspecified operations (e.g., bit-vector
division), whose precise semantics were being debated at the time.
Additionally in~2016, one benchmark was excluded because it contained
incorrect status information.  In~2017, \num{1104} benchmarks were
excluded because their logic had been classified incorrectly, and
8~benchmarks were excluded because of incorrect or missing status
information.  Moreover, divisions containing datatypes ({DT})---which
were new in~2017---were excluded from the Unsat-Core track because of
inadequate tool support.  In~2018, \num{3603}~benchmarks were excluded
because they contained named terms (a feature of the SMT-LIB language
that is not permitted in competition benchmarks), and 5~benchmarks
were excluded because of missing status information.

\begin{table}
  \caption{Number of divisions in each competition track}
  \label{table:divisions}
  \centering
  \begin{tabular}{lcccc}
    \toprule
    & 2015 & 2016 & 2017 & 2018 \\
    \midrule
    \maintrack   &  40 & 40 & 49 & 50 \\
    \apptrack    &  14 & 14 & 16 & 21 \\
    \ucoretrack  &  -- & 40 & 41 & 44 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Benchmark selection.}
SMT Competitions until~2014 employed a pseudo-random selection process
to further cull the number of benchmarks used in the competition.
However, a 2013 evaluation of SMT-COMP and SMT-LIB found that this
``significantly lessens the ability of a competition to determine
`best' solvers''~\cite{CSW15}. Together with the increased
computational resources available to the competition since its move to
StarExec, this finding prompted a paradigm shift.  Since 2015, the
competition is evaluating all solvers on all eligible benchmarks.

\paragraph{Benchmarks with unknown status.}
Another paradigm shift concerns the treatment of benchmarks with
unknown status.  Because of the difficulty to judge solver responses
for such benchmarks, SMT Competitions until 2015 only used benchmarks
whose status was known (i.e., that had been solved previously by
existing SMT solvers).  In 2015, \num{30717} non-incremental
benchmarks from the SMT Library were not used for the competition
because their status was unknown.  However, this approach was
criticized for rewarding imitation of existing solvers rather than
true innovation.  In~2016, the SMT Library contained \num{29724}
non-incremental benchmarks with unknown status; solver performance on
these was evaluated but reported separately.  Since 2017, all
non-incremental benchmarks are eligible for the competition's
\maintrack regardless of status, and they directly affect the
competition results.

Solver responses for such benchmarks are assumed to be correct, unless
there is a disagreement between two or more solvers that are otherwise
(i.e., on benchmarks with known status) sound.  Benchmarks with
disagreements are removed from the competition results.  This
conservative approach was chosen over a voting system as the latter is
unreliable (especially in divisions with few solvers), could lead to a
solver being punished wrongly, and could be gamed by entering multiple
copies of a solver into the competition.  Moreover, the size and time
frame of the competition demand an automated resolution.
Disagreements are reported to the solver developers involved, who
typically perform a manual investigation at their leisure, and in all
cases have led to bugs in solvers being identified and fixed.  In
2016, there were disagreements on 79 benchmarks; in 2017, only one
benchmark was removed; while in 2018, there were no disagreements
between otherwise sound solvers.

For incremental benchmarks, the situation is different: an incremental
benchmark is ineligible for the \apptrack if its first {\tt check-sat}
command has unknown status.  Otherwise, the prefix of the benchmark
that only contains {\tt check-sat} commands with known status is
eligible.  The main reason for this difference is that the competition
tools for the \apptrack (in particular, the trace executor) had not
been adapted yet to support {\tt check-sat} commands with unknown
status. Even though the trace executor has now been modified, we
cannot check what impact this may have had on previous results without
rerunning the \apptrack as the trace executor drives the solver.

Tables~\ref{table:benchmarks-main-track}--\ref{table:benchmarks-unsat-core-track}
show the number of benchmarks that were used in each competition track
and division since~2015.  A dash indicates that no eligible benchmarks
were available, and hence the corresponding SMT-LIB logic was not
actually a competition division in that track and year.  The size of
divisions varies widely, from just one benchmark in ABVFP
to~\num{72705} benchmarks in QF\_SLIA.

\begin{table}
  \caption{Number of benchmarks used in the \maintrack (by division). Numbers in parentheses indicate the (rounded) percentage of benchmarks with different statuses, which are (sat/unsat) in 2015 and 2016, and (sat/unsat/unknown) in 2017 and 2018.}
  \label{table:benchmarks-main-track}
  \centering
  \scalebox{.9}[.775]{\begin{tabular}{lr@{\hskip3pt}cr@{\hskip3pt}cr@{\hskip3pt}cr@{\hskip3pt}c}
    \toprule
    Division & \multicolumn{2}{c}{2015} & \multicolumn{2}{c}{2016} & \multicolumn{2}{c}{2017} & \multicolumn{2}{c}{2018} \\
    \midrule
    ABVFP & -- & & -- & & -- & & 1 & \gray{(0/0/100)}\\
    ALIA & 42 & \gray{(2/98)} & 42 & \gray{(2/98)} & 42 & \gray{(2/98/0)} & 42 & \gray{(2/98/0)}\\
    AUFBVDTLIA & -- & & -- & & 1709 & \gray{(0/1/98)} & 1709 & \gray{(0/1/98)}\\
    AUFDTLIA & -- & & -- & & 728 & \gray{(0/0/100)} & 728 & \gray{(0/0/100)}\\
    AUFLIA & 4 & \gray{(25/75)} & 4 & \gray{(25/75)} & 4 & \gray{(25/75/0)} & 4 & \gray{(25/75/0)}\\
    AUFLIRA & 19849 & \gray{(1/99)} & 19849 & \gray{(1/99)} & 20011 & \gray{(0/99/1)} & 20011 & \gray{(0/99/1)}\\
    AUFNIRA & 1050 & \gray{(0/100)} & 1050 & \gray{(0/100)} & 1480 & \gray{(0/71/29)} & 1480 & \gray{(0/71/29)}\\
    BV & 85 & \gray{(34/66)} & 85 & \gray{(34/66)} & 5151 & \gray{(2/2/97)} & 5751 & \gray{(10/86/4)}\\
    BVFP & -- & & -- & & -- & & 24 & \gray{(0/0/100)}\\
    FP & -- & & -- & & -- & & 61 & \gray{(0/0/100)}\\
    LIA & 201 & \gray{(5/95)} & 201 & \gray{(5/95)} & 388 & \gray{(38/60/2)} & 388 & \gray{(38/60/2)}\\
    LRA & 339 & \gray{(6/94)} & 339 & \gray{(6/94)} & 2419 & \gray{(17/46/38)} & 2419 & \gray{(29/64/8)}\\
    NIA & 9 & \gray{(67/33)} & 9 & \gray{(67/33)} & 14 & \gray{(43/29/29)} & 14 & \gray{(43/29/29)}\\
    NRA & 3788 & \gray{(0/100)} & 3788 & \gray{(0/100)} & 3811 & \gray{(0/100/0)} & 3813 & \gray{(0/100/0)}\\
    QF\_ABV & 14720 & \gray{(68/32)} & 14720 & \gray{(68/32)} & 15061 & \gray{(69/31/0)} & 15066 & \gray{(69/31/0)}\\
    QF\_ABVFP & -- & & -- & & -- & & 18129 & \gray{(78/22/1)}\\
    QF\_ALIA & 134 & \gray{(40/60)} & 139 & \gray{(42/58)} & 139 & \gray{(42/58/0)} & 139 & \gray{(42/58/0)}\\
    QF\_ANIA & 6 & \gray{(0/100)} & 8 & \gray{(0/100)} & 8 & \gray{(0/100/0)} & 8 & \gray{(0/100/0)}\\
    QF\_AUFBV & 37 & \gray{(16/84)} & 37 & \gray{(16/84)} & 31 & \gray{(19/81/0)} & 31 & \gray{(19/81/0)}\\
    QF\_AUFLIA & 1009 & \gray{(49/51)} & 1009 & \gray{(49/51)} & 1009 & \gray{(49/51/0)} & 1009 & \gray{(49/51/0)}\\
    QF\_AUFNIA & 21 & \gray{(29/71)} & 21 & \gray{(29/71)} & 17 & \gray{(29/71/0)} & 17 & \gray{(29/71/0)}\\
    QF\_AX & 551 & \gray{(49/51)} & 551 & \gray{(49/51)} & 551 & \gray{(49/51/0)} & 551 & \gray{(49/51/0)}\\
    QF\_BV & 26414 & \gray{(35/65)} & 26414 & \gray{(35/65)} & 40043 & \gray{(32/59/8)} & 40102 & \gray{(36/64/0)}\\
    QF\_BVFP & 7 & \gray{(14/86)} & 7 & \gray{(14/86)} & 17215 & \gray{(81/18/0)} & 17215 & \gray{(81/18/0)}\\
    QF\_DT & -- & & -- & & 8000 & \gray{(0/0/100)} & 8000 & \gray{(45/55/0)}\\
    QF\_FP & 34413 & \gray{(50/50)} & 34413 & \gray{(50/50)} & 40302 & \gray{(50/50/1)} & 40300 & \gray{(50/50/1)}\\
    QF\_IDL & 2094 & \gray{(61/39)} & 2094 & \gray{(61/39)} & 2193 & \gray{(58/37/5)} & 2193 & \gray{(58/37/4)}\\
    QF\_LIA & 5839 & \gray{(51/49)} & 5839 & \gray{(51/49)} & 6141 & \gray{(52/46/2)} & 6947 & \gray{(55/43/1)}\\
    QF\_LIRA & 6 & \gray{(17/83)} & 6 & \gray{(17/83)} & 7 & \gray{(14/71/14)} & 7 & \gray{(14/71/14)}\\
    QF\_LRA & 1626 & \gray{(61/39)} & 1626 & \gray{(61/39)} & 1649 & \gray{(57/41/2)} & 1649 & \gray{(57/41/1)}\\
    QF\_NIA & 8475 & \gray{(98/2)} & 8593 & \gray{(96/4)} & 23876 & \gray{(59/13/28)} & 23876 & \gray{(59/20/20)}\\
    QF\_NIRA & 2 & \gray{(0/100)} & 2 & \gray{(0/100)} & 3 & \gray{(0/67/33)} & 3 & \gray{(0/67/33)}\\
    QF\_NRA & 10184 & \gray{(52/48)} & 10245 & \gray{(52/48)} & 11354 & \gray{(44/47/10)} & 11489 & \gray{(44/47/9)}\\
    QF\_RDL & 220 & \gray{(49/51)} & 220 & \gray{(49/51)} & 255 & \gray{(43/44/13)} & 255 & \gray{(43/44/13)}\\
    QF\_SLIA & -- & & -- & & -- & & 72705 & \gray{(0/0/100)}\\
    QF\_UF & 6649 & \gray{(38/62)} & 6649 & \gray{(38/62)} & 6650 & \gray{(38/62/0)} & 7423 & \gray{(42/58/0)}\\
    QF\_UFBV & 31 & \gray{(0/100)} & 31 & \gray{(0/100)} & 31 & \gray{(0/100/0)} & 1224 & \gray{(53/47/0)}\\
    QF\_UFIDL & 441 & \gray{(24/76)} & 441 & \gray{(24/76)} & 428 & \gray{(25/75/0)} & 428 & \gray{(25/75/0)}\\
    QF\_UFLIA & 598 & \gray{(67/33)} & 598 & \gray{(67/33)} & 583 & \gray{(69/31/0)} & 583 & \gray{(69/31/0)}\\
    QF\_UFLRA & 1627 & \gray{(48/52)} & 1627 & \gray{(48/52)} & 1284 & \gray{(60/40/0)} & 1284 & \gray{(60/40/0)}\\
    QF\_UFNIA & 7 & \gray{(0/100)} & 7 & \gray{(0/100)} & 7 & \gray{(0/100/0)} & 7 & \gray{(0/100/0)}\\
    QF\_UFNRA & 34 & \gray{(47/53)} & 34 & \gray{(47/53)} & 36 & \gray{(58/31/11)} & 36 & \gray{(58/31/11)}\\
    UF & 2839 & \gray{(28/72)} & 2839 & \gray{(28/72)} & 7562 & \gray{(10/44/46)} & 7562 & \gray{(10/46/44)}\\
    UFBV & 71 & \gray{(25/75)} & 71 & \gray{(25/75)} & 200 & \gray{(34/48/17)} & 200 & \gray{(34/48/17)}\\
    UFDT & -- & & -- & & 4535 & \gray{(0/9/90)} & 4527 & \gray{(1/41/58)}\\
    UFDTLIA & -- & & -- & & 303 & \gray{(0/0/100)} & 303 & \gray{(0/0/100)}\\
    UFIDL & 68 & \gray{(9/91)} & 68 & \gray{(9/91)} & 68 & \gray{(4/84/12)} & 68 & \gray{(4/84/12)}\\
    UFLIA & 8404 & \gray{(0/100)} & 8404 & \gray{(0/100)} & 10137 & \gray{(0/76/24)} & 10137 & \gray{(0/76/23)}\\
    UFLRA & 25 & \gray{(20/80)} & 25 & \gray{(20/80)} & 15 & \gray{(33/67/0)} & 15 & \gray{(33/67/0)}\\
    UFNIA & 2319 & \gray{(0/100)} & 2319 & \gray{(0/100)} & 3308 & \gray{(0/74/26)} & 3308 & \gray{(0/74/26)}\\
    \midrule
    Total      & 154238 & \gray{(40/60)} & 154424 & \gray{(40/60)} & 256973 & \gray{(37/46/17)} & 333241 & \gray{(33/39/28)}\\
    \bottomrule
  \end{tabular}}
\end{table}

\begin{table}
  \caption{Number of benchmarks used in the \apptrack (by division)}
  \label{table:benchmarks-application-track}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    Division & \multicolumn{1}{c}{2015} & \multicolumn{1}{c}{2016} & \multicolumn{1}{c}{2017} & \multicolumn{1}{c}{2018} \\
    \midrule
    ALIA       &   24 &   24 &   24 &   24 \\
    ANIA       &    3 &    3 &    3 &    3 \\
    AUFNIRA    &   -- &   -- &   -- &  117 \\
    BV         &   -- &   -- &   -- &   17 \\
    LIA        &    6 &    6 &    6 &    6 \\
    QF\_ABV    &   -- &   -- &   -- &   15 \\
    QF\_ALIA   &   44 &   44 &   44 &   44 \\
    QF\_ANIA   &    5 &    5 &    5 &    5 \\
    QF\_AUFBV  &   -- &   -- &   -- &   10 \\
    QF\_AUFLIA &   72 &   72 &   72 &   72 \\
    QF\_BV     &   18 &   18 &   18 &  815 \\
    QF\_BVFP   &   -- &   -- &    2 &    2 \\
    QF\_FP     &   -- &   -- &    2 &    2 \\
    QF\_LIA    &   65 &   69 &   68 &   69 \\
    QF\_LRA    &   10 &   10 &   10 &   10 \\
    QF\_NIA    &   10 &   10 &   10 &   10 \\
    QF\_UFBV   &   -- &   -- &   -- & 2327 \\
    QF\_UFLIA  &  905 &  905 &  780 &  780 \\
    QF\_UFLRA  & 3331 & 3331 & 3056 & 3058 \\
    QF\_UFNIA  &    1 &    1 &    1 &    1 \\
    UFLRA      & 5358 & 5358 & 1870 & 1870 \\
    \midrule
    Total      & 9852 & 9856 & 5971 & 9257 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}
  \caption{Number of benchmarks used in the \ucoretrack (by
    division)}
  \label{table:benchmarks-unsat-core-track}
  \centering
  \scalebox{1}[.91]{\begin{tabular}{lrrr}
    \toprule
    Division & \multicolumn{1}{c}{2016} & \multicolumn{1}{c}{2017} & \multicolumn{1}{c}{2018} \\
    \midrule
    ALIA       &     41 &     41 &     41 \\
    AUFBVDTLIA &     -- &     -- &     25 \\
    AUFLIA     &      3 &      3 &      3 \\
    AUFLIRA    &  19749 &  19771 &  19771 \\
    AUFNIRA    &   1046 &   1050 &   1053 \\
    BV         &     56 &     94 &   4937 \\
    LIA        &    191 &    233 &    233 \\
    LRA        &    319 &   1106 &   1539 \\
    NIA        &      3 &      4 &      4 \\
    NRA        &   3788 &   3801 &   3801 \\
    QF\_ABV    &   4644 &   4673 &   4677 \\
    QF\_ABVFP  &     -- &   3934 &   3934 \\
    QF\_ALIA   &     80 &     80 &     80 \\
    QF\_ANIA   &      8 &      8 &      8 \\
    QF\_AUFBV  &     31 &     25 &     25 \\
    QF\_AUFLIA &    516 &    516 &    516 \\
    QF\_AUFNIA &     15 &     12 &     12 \\
    QF\_AX     &    279 &    279 &    279 \\
    QF\_BV     &  17172 &  23732 &  25700 \\
    QF\_BVFP   &      6 &   3174 &   3174 \\
    QF\_DT     &     -- &     -- &   4422 \\
    QF\_FP     &  17213 &  20028 &  20026 \\
    QF\_IDL    &    816 &    816 &    816 \\
    QF\_LIA    &   2840 &   2844 &   3019 \\
    QF\_LIRA   &      5 &      5 &      5 \\
    QF\_LRA    &    633 &    671 &    683 \\
    QF\_NIA    &    316 &   3130 &   4842 \\
    QF\_NIRA   &      2 &      2 &      2 \\
    QF\_NRA    &   4948 &   5296 &   5357 \\
    QF\_RDL    &    113 &    113 &    113 \\
    QF\_UF     &   4100 &   4101 &   4330 \\
    QF\_UFBV   &     31 &     31 &    575 \\
    QF\_UFIDL  &    335 &    322 &    322 \\
    QF\_UFLIA  &    195 &    183 &    183 \\
    QF\_UFLRA  &    853 &    511 &    511 \\
    QF\_UFNIA  &      7 &      7 &      7 \\
    QF\_UFNRA  &     18 &     11 &     11 \\
    UF         &   2039 &   3316 &   3442 \\
    UFBV       &     53 &     97 &     97 \\
    UFDT       &     -- &     -- &   1863 \\
    UFIDL      &     62 &     57 &     57 \\
    UFLIA      &   8377 &   7714 &   7743 \\
    UFLRA      &     20 &     10 &     10 \\
    UFNIA      &   2318 &   2432 &   2457 \\
    \midrule
    Total      &  93241 & 114233 & 130705 \\
    \bottomrule
  \end{tabular}}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Infrastructure}
\label{sec:starexec}

Since 2014 (and previously for the evaluation in 2013~\cite{CSW15}),
the computational workload of the competition is being run on
StarExec, a ``cross-community infrastructure for logic
solving''~\cite{DBLP:conf/cade/StumpST14} developed at the University
of Iowa and supported by the National Science Foundation.  Each
machine in the StarExec cluster is equipped with two E5-2609 Intel
Xeon processors (10~MiB Cache, 2.4~GHz).  This hardware configuration
has been unchanged since~2013.  Up to two solvers are run
simultaneously on each machine: one solver on each processor, with
four cores available.  Memory usage is limited to~60~GiB per solver.

On the software side, the cluster is running Red Hat Enterprise Linux
Workstation.  The~2015 competition was running on release~6.3
(Santiago, kernel 2.6.32-431).  This was updated to~7.2 (Maipo, kernel
3.10.0-327) before the~2016 competition, and the kernel was further
updated to~3.10.0-514 before the~2017 competition.  The platform
emphasizes stability over bleeding edge technology.  This has on
occasion caused problems with solver binaries that had been compiled
on other Linux distributions and linked against library versions not
(yet) available on the cluster.  Usually, such problems are easily
solved by building the solver on StarExec instead.  The StarExec
developers have made a virtual machine image available to facilitate
this and similar tasks.

Interaction with StarExec is mostly via its web interface,
\url{https://www.starexec.org}.  The interface allows registered users
to upload and share benchmarks and solvers, to create and manage jobs
that apply selected solvers to a selection of benchmarks, and to view
and download job results.  Other features, such as pre-processing of
benchmarks (e.g., for scrambling) and post-processing of solver
output, are also supported.

Jobs consist of individual \emph{job pairs}.  Each job pair
corresponds to a specific solver/benchmark combination.
Figure~\ref{fig:job-pairs} shows the total number of job pairs for the
SMT Competition from~2015 to~2018.  Over this period, the number has
grown by~72\,\%, from~\num{1028615} job pairs in~2015 to~\num{1776062}
job pairs in~2018.  (This is the number of unique solver/benchmark
combinations that were part of the competition.  The number of job
pairs that were actually run is even higher: test jobs and job pairs
that had to be re-run, e.g., because of issues with the StarExec
infrastructure, are not included.)  For comparison, the~2014
competition only ran~\num{339714} job pairs~\cite{CDW14}.

Also shown in Figure~\ref{fig:job-pairs} is the total wall-clock time
(summed over all job pairs) for the competition.  This has more than
doubled, from 3.5~years in~2015 to 7.6~years in~2018.  These are
solver run-times only; time that the cluster spent on pre- and
post-processing of benchmarks and solver output or on other auxiliary
tasks is not included.  The increase in~2016 is largely due to the
inclusion of (non-incremental) benchmarks with unknown status; these
were evaluated with a wall-clock time limit of 10~minutes.  In~2017,
when these benchmarks had become eligible for the \maintrack,
\maintrack job pairs were first run with a time limit of 10~minutes.
When it became apparent to the organizers that there were sufficient
resources to increase this limit, job pairs that had timed out were
re-run with a time limit of 20~minutes.  (In this way, about 2.9 years
of wall-clock time in 2017 were spent on job pairs that were then
re-run.  This time is not included in Figure~\ref{fig:job-pairs}.)
Table~\ref{table:time-limits} shows the wall-clock time limits that
were imposed on job pairs for each track and year.

\begin{table}
  \caption{Wall-clock time limits (in minutes)}
  \label{table:time-limits}
  \centering
  \begin{tabular}{lcccc}
    \toprule
    & 2015 & 2016 & 2017 & 2018 \\
    \midrule
    \maintrack   & 40 & 40 & 20\protect\footnotemark & 20 \\
    \apptrack    & 40 & 40 & 40\phantom{$^{1.}$}      & 40 \\
    \ucoretrack  & -- & 40 & 40\phantom{$^{1.}$}      & 40 \\
    \bottomrule
  \end{tabular}
\end{table}
\footnotetext{These job pairs were first run with a time limit of
  10~minutes.  See the text for details.}

StarExec typically reserves between~80 to~150 machines for the SMT
Competition, allowing its workload to be completed within two to three
weeks.

\begin{figure}
  \centering
  \scalebox{.98}{\begin{tikzpicture}
    \draw [gray] (0,0) -- (12,0);
    \draw [gray] (0,1.5) -- (12,1.5);
    \draw [gray] (0,3) -- (12,3);
    \draw [gray] (0,4.5) -- (12,4.5);
    \draw [gray] (0,6) -- (12,6);

    \node [left] at (0,0) {\footnotesize 0};
    \node [left] at (0,1.5) {\footnotesize 500,000};
    \node [left] at (0,3) {\footnotesize 1,000,000};
    \node [left] at (0,4.5) {\footnotesize 1,500,000};
    \node [left] at (0,6) {\footnotesize 2,000,000};

    \node [rotate=90] at (-2,3) {\small Job pairs};

    \node [right] at (12,0) {\footnotesize 0};
    \node [right] at (12,1.5) {\footnotesize 2};
    \node [right] at (12,3) {\footnotesize 4};
    \node [right] at (12,4.5) {\footnotesize 6};
    \node [right] at (12,6) {\footnotesize 8};

    \node [rotate=-90] at (12.9,3) {\small Wall-clock time (years)};

    \draw [fill=black] (1,0) rectangle (2,3.085845); % 1,028,615
    \draw [fill=black] (3.75,0) rectangle (4.75,4.687632); % 1,562,544
    \draw [fill=black] (6.5,0) rectangle (7.5,4.366839); % 1,455,613
    \draw [fill=black] (9.25,0) rectangle (10.25,5.328186); % 1,776,062

    \draw [fill=white] (2,0) rectangle (3,2.6088709094); % 3,4784945459
    \draw [fill=white] (4.75,0) rectangle (5.75,4.3030490392); % 5,7373987189
    \draw [fill=white] (7.5,0) rectangle (8.5,5.6286127759); % 7,5048170345
    \draw [fill=white] (10.25,0) rectangle (11.25,5.7280186691); % 7,6373582255

    \node [below] at (2,0) {\footnotesize 2015};
    \node [below] at (4.75,0) {\footnotesize 2016};
    \node [below] at (7.5,0) {\footnotesize 2017};
    \node [below] at (10.25,0) {\footnotesize 2018};
  \end{tikzpicture}}
  \caption{Number of job pairs ($\blacksquare$) and total wall-clock time ($\square$).}
  \label{fig:job-pairs}
\end{figure}

The use of StarExec ensures the replicability of the competition. The
publicly accessible space
\url{https://www.starexec.org/starexec/secure/explore/spaces.jsp?id=2641}
stores all of the information required to rerun the competition, e.g.,
all benchmarks and copies of all tools. In addition, a GitHub
organization (\url{https://github.com/SMT-COMP/smt-comp}) has been set
up to archive previous results and store the tools used by organizers
to run the competition.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Participants}
\label{sec:participants}

Participants must upload their solvers to StarExec to enter them into
the competition, and additionally submit information about each
competing solver---such as the tracks and divisions into which the
solver is being entered---to the competition organizers.  Until 2016,
this information was collected by email.  Since 2017, a web form is
being used for this purpose.  This has greatly reduced the number of
incomplete or ambiguous submissions.

Participants are required to be authors of the tool they submit.  In
2015--2018, the competition rules did not further define authorship,
but it is understood that an author contributed code to the tool.  As
a special case of this rule, participants are required to identify if
their tool is a wrapper tool, i.e., any solver that calls (or is in
some other way based on) one or more SMT solvers of which the author
of the tool is not an author.

Tables~\ref{table:participants1}--\ref{table:participants3} list the
solvers that participated between~2015 and~2018, together with
contributing team members (in alphabetical order) and their
institutional affiliations (if any; these may have changed over time)
as reported by the development teams.

Further, to promote a wide comparison between tools, there was no
restriction on the number of solver versions participants may submit,
e.g., for different competition tracks or with differently tuned
internal components.  Organizers reserved the right to not accept
multiple versions of a solver in case the number of solver submissions
is too large given the computational resources, but in the past four
years, exercising this right was not necessary.
Table~\ref{table:participants-history} shows in more detail how many
versions of each solver were submitted in each year.

Participants were not required to provide source code with their
submission.  There were no licensing requirements/restrictions for the
submitted binaries of the participating solvers.
%
Participants were encouraged to submit a short system description of
their solver, including a list of its authors and an explanation of
the SMT solving approach used.  System descriptions are published on
the competition website.  They were submitted for about 60\,\% of all
solver versions since 2015.

In addition to the solvers that were submitted to the competition by
their respective developers, the organizers included the most recent
stable versions of MathSAT and Z3 for comparison purposes.  Both
solvers are strong tools, but---except for two experimental versions
of Z3 in 2015---their development teams chose not to prepare
competition versions.  These solvers were therefore participating hors
concours.  Moreover, in 2015 a bug was discovered in the \apptrack
version of Boolector and a fixed version was submitted after the
deadline; in 2016, the CVC4 team did not enter their solver into the
\apptrack; and in 2018, an experimental version of CVC4 as well as the
OpenSMT2 solver were submitted after the deadline.  These solvers were
also participating hors concours.  In result tables, they are listed
with their name in square brackets (e.g., [MathSAT]).

\include{table-participants}

\begin{table}
  \caption{Number of versions submitted for each solver (by year).
    Numbers in square brackets indicate solver versions that were
    participating hors concours.}
  \label{table:participants-history}
  \centering
  \begin{tabular}{lr@{\,\,}rr@{\,\,}rr@{\,\,}rr@{\,\,}r}
    \toprule
    Solver & \multicolumn{2}{c}{2015} & \multicolumn{2}{c}{2016} & \multicolumn{2}{c}{2017} & \multicolumn{2}{c}{2018} \\
    \midrule
    ABC               &    &      &  2 &      &    &      &    &      \\
    Alt-Ergo          &    &      &    &      &    &      &  1 &      \\
    AProVE            &  1 &      &  1 &      &  1 &      &  1 &      \\
    Boolector         &  3 & [+1] &  2 &      &  2 &      &  2 &      \\
    COLIBRI           &    &      &    &      &  1 &      &  1 &      \\
    Ctrl-Ergo         &    &      &    &      &    &      &  1 &      \\
    CVC3              &  2 &      &    &      &    &      &    &      \\
    CVC4              &  4 &      &  1 & [+1] &  3 &      &  3 & [+1] \\
    MapleSTP          &    &      &  4 &      &    &      &    &      \\
    {[}MathSAT{]}     &    & [+2] &    & [+3] &    & [+3] &    & [+3] \\
    MinkeyRink        &    &      &  1 &      &  1 &      &  2 &      \\
    OpenSMT2          &  2 &      &  1 &      &  1 &      &    & [+1] \\
    ProB              &    &      &  1 &      &    &      &    &      \\
    Q3B               &    &      &  1 &      &  1 &      &  1 &      \\
    raSAT             &  1 &      &  2 &      &    &      &    &      \\
    Redlog            &    &      &    &      &  1 &      &    &      \\
    SMT-RAT           &  2 &      &  1 &      &  1 &      &  2 &      \\
    SMTInterpol       &  1 &      &  1 &      &  1 &      &  1 &      \\
    SPASS-SATT        &    &      &    &      &    &      &  1 &      \\
    STP               &  4 &      &  8 &      &  2 &      &  3 &      \\
    toysmt            &    &      &  1 &      &    &      &    &      \\
    Vampire           &    &      &  2 &      &  1 &      &  1 &      \\
    veriT             &  1 &      &  1 &      &  3 &      &  2 &      \\
    XSat              &    &      &    &      &  1 &      &    &      \\
    Yices            &  3 &      &  2 &      &  2 &      &  3 &      \\
    Z3                &  2 & [+1] &    & [+1] &    & [+1] &    & [+1] \\
    \midrule
    Total             & 26 & [+4] & 32 & [+5] & 22 & [+4] & 25 & [+6] \\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{table:participation-by-track} summarizes these numbers and
shows how many solver versions were submitted to each track of the
competition.  Note that solver versions may be entered into multiple
tracks.  Therefore, the total number of solver versions for each year
is typically less than the sum over all tracks.

It can be easy for developers to create multiple versions of their
solver, which may differ only in configuration settings or other minor
details.  We caution against over-interpreting these numbers, which to
some extent depend on how adventurous development teams were in any
given year.

By number of solver versions submitted, the competitions in 2015--2018
were the four largest in the history of SMT-COMP.  On average, 26
solver versions were submitted each year since 2015.  In contrast, the
competitions from 2005--2014 only received an average of~12
submissions~\cite{CDW14}.

\begin{table}
  \caption{Participation by track and year.  Numbers in square
    brackets indicate solver versions that were participating hors
    concours.}
  \label{table:participation-by-track}
  \centering
  \begin{tabular}{lr@{\,\,}rr@{\,\,}rr@{\,\,}rr@{\,\,}r}
    \toprule
                      & \multicolumn{2}{c}{2015} & \multicolumn{2}{c}{2016} & \multicolumn{2}{c}{2017} & \multicolumn{2}{c}{2018} \\
    \midrule
    \maintrack   & 21 &               [+2] & 25 & [+2] & 19 & [+2] & 20 & [+4] \\
    \apptrack    & 10 &               [+3] &  8 & [+3] &  4 & [+2] &  4 & [+2] \\
    \ucoretrack  & \multicolumn{2}{c}{---} &  1 & [+4] &  2 & [+2] &  3 & [+2] \\
    \midrule
    Total             & 26 &               [+4] & 32 & [+5] & 22 & [+4] & 25 & [+6] \\
    \bottomrule
  \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Scoring}
\label{sec:scoring}

A detailed description of the scoring can be found in the competition
rules~\cite{rules18}.  We summarize the most important aspects in this
section.

\subsection{Division Rankings}
\label{sec:scoring-division-rankings}

Solvers in each division (of each track) are ranked according to a
metric that is based, first, on the number of erroneous responses
(which is usually~0, but solvers that did give erroneous responses are
ranked below solvers that did not); second, on the number of correct
responses; third and fourth, on the wall-clock and CPU time,
respectively, that the solver process consumed.
%
The \emph{raw benchmark score} of a solver is thus a quadruple
$\langle e, n, w, c\rangle$, with
\begin{itemize}[noitemsep]
  \item \makebox[5em][l]{$e\geq 0$} number of erroneous responses
    (usually~$e = 0$),
  \item \makebox[5em][l]{$0 \leq n \leq N$} number of correct
    responses (resp.~\emph{reduction} for the \ucoretrack),
  \item \makebox[5em][l]{$w \in [0,T]$} wall-clock time in seconds
    (real-valued),
  \item \makebox[5em][l]{$c \in [0, 4T]$} CPU time in seconds
    (real-valued).
\end{itemize}
%
For the \maintrack and \apptrack, $e$ is the number of returned
statuses that disagree with the given expected status; $e \in \{0,1\}$
for the \maintrack.  For the \ucoretrack, $e$ includes, in addition,
the number of returned unsat cores that are ill-formed or are not, in
fact, unsatisfiable (as validated by a selection of solvers selected
by the organizers).

For the \maintrack and \apptrack, $N$ is defined as the number of {\tt
  check-sat} commands ($N=1$ for the \maintrack), and $n$ is defined
as the number of correct responses.  For the \ucoretrack (which uses
unsatisfiable benchmarks only), $N$ is defined as the number of named
top-level assertions, and~$n$ is defined as the \emph{reduction},
i.e., the difference between $N$ and the size of the unsat core
returned by the solver.

As discussed in Section~\ref{sec:benchmarks}, \maintrack benchmarks
with unknown status are removed from the competition results if two
(or more) solvers that are sound on benchmarks with known status
disagree on their result.  Otherwise, solver responses for benchmarks
with unknown status are assumed to be correct.

A division is \emph{competitive} if at least two substantially
different solvers (i.e., solvers from two different teams) were
submitted.  (Experimental divisions are never competitive.)  Official
winners were declared only for competitive divisions.

\paragraph{Parallel vs.\ sequential performance.}

All solvers are run with four cores available, and the time limit~$T$
that is imposed on each job pair (see Table~\ref{table:time-limits})
is a wall-clock limit.  Thus, solvers that take advantage of
parallelism might use up to~$4T$ CPU time.  For the \maintrack and the
\ucoretrack, the competition reports parallel and sequential
performance separately.  The raw benchmark score corresponds to the
\emph{raw parallel score}---all values are determined for solver
performance within the wall-clock time limit.  We derive a \emph{raw
  sequential score} by imposing a (virtual) CPU time limit equal to
the wall-clock time limit~$T$.  A solver response is taken into
consideration for the sequential score only if the solver process
terminates within this CPU time limit.  For the \apptrack, only
parallel performance is reported (because response times to individual
{\tt check-sat} commands are not recorded by StarExec).

\paragraph{Benchmark weights.}

Until 2015, the competition used these raw scores for all tracks to
determine the division score of a solver as the sum of the scores of
all benchmarks in that division.  However, there are vast differences
in size between benchmark \emph{families} (i.e., groups of benchmarks
with similar characteristics) in the SMT Library.  The pseudo-random
selection process of benchmarks that was used until~2014 (see
Section~\ref{sec:benchmarks}) ensured that large families were not
overemphasized in the competition.  Since~2016, the competition uses a
scoring method that assigns a weight to each benchmark in the
\maintrack and \ucoretrack to achieve the same
goal~\cite{rules16,rules17,rules18}.

The \emph{weight} of a benchmark~$b$ in the \maintrack is computed as
$\alpha_b = (1 + \ln F_b)/F_b$, where $F_b \geq 1$ is the size of the
benchmark's family.  We then define its \emph{normalized weight} as
$\alpha'_b = \alpha_b / (\sum_{b'} \alpha_{b'})$, where the sum is
over all benchmarks in the division.  The logarithmic scaling seems a
reasonable compromise between linearly combining numbers of
benchmarks, which would overemphasize large families, and simply
averaging the fraction of benchmarks solved, which would overemphasize
small families.

The \emph{(weighted) score} of a solver in a (\maintrack or
\ucoretrack) division~$i$ is then defined as the weighted sum of all
raw scores
%
$$\sum_{b\in i} \alpha'_b \cdot \langle e_b \cdot N_i, n_b \cdot N_i, w_b, c_b\rangle$$
%
where the sum is over all benchmarks in the division, and~$N_i$ is the
number of benchmarks in~$i$.  This is computed separately for parallel
and sequential scores.

For \apptrack divisions, division scores were still based on raw
scores only, since \apptrack benchmarks may be partially solved.

\subsection{Competition-Wide Ranking (\maintrack)}
\label{sec:scoring-floc}

In 2014, the SMT Competition was part of the FLoC Olympic Games, which
sponsored three Kurt G\"odel medals.  These were awarded to the winner
of the \maintrack's QF\_BV division (which had historically been the
logic with the largest number of solver submissions and job pairs), as
well as to the two best \maintrack solvers according to a
competition-wide ranking that had been instituted especially for this
purpose~\cite{CDW14}.  This ranking combines each solver's performance
across all competitive \maintrack divisions into a single score,
giving more weight to larger divisions, and to solvers that solve a
large fraction of the benchmarks~\cite{rules18}.  It emphasizes
breadth of solver participation---a solver participating in many
logics does not need to be the best in any one of them to rank in high
position.

The competition-wide ranking for the \maintrack was retained and
refined after~2014.  Sequential and parallel performance are reported
separately.  Let $N_i$ be the total number of benchmarks in
division~$i$ that were used in the competition (and not removed
because of disagreements), and let $\langle e_i, n_i, w_i, c_i\rangle$
be a solver's raw score for this division.
%
In 2015 and~2016, the \emph{competition-wide raw score} for a solver
was defined as
%
$$\sum_i (e_i == 0 \,?\, (n_i/N_i)^2 : -e_i) \ln N_i$$
%
where the sum is over all competitive divisions into which the solver
was entered.
%
In 2017, the (per-division) penalty for erroneous results was changed
from the number of errors to a fixed constant, since the division
weight~$\ln N_i$ already takes the number of benchmarks in the
division into account, and multiplying it by the number of errors
overemphasizes large divisions.  In 2017 and 2018, the
\emph{competition-wide raw score} for a solver was thus defined as
%
$$\sum_i (e_i == 0 \,?\, (n_i/N_i)^2 : -4) \ln N_i$$
%
The value~$-4$ was chosen to balance the strong interest in correct
solvers against the risk of stifling innovation.  In the division
rankings (Section~\ref{sec:scoring-division-rankings}), an incorrect
response causes a solver to be ranked below all correct solvers for
that division.  But the competition's \maintrack consists of many
independent divisions; a solver bug that affects one division need not
manifest itself in other divisions.  With this value, entering a
(possibly buggy) solver that can solve all benchmarks into a division
has a positive expected score if the probability that a soundness bug
in the solver will be triggered is below~$20\,\%$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}
\label{sec:results}

\subsection{Division Rankings}
\label{sec:division-rankings}

Tables~\ref{tab:winners-main}--\ref{tab:winners-unsat-core} show the
highest-ranked solvers for all (competitive and non-competitive)
divisions from 2015--2018.  We include solvers that were running hors
concours (and hence non-competing and ineligible to be recognized as
official competition winners).  Non-com-petitive divisions use gray
font.  Entries $A [B]$ indicate that solver $A$ was the highest
competing solver, but non-competing solver $B$ actually ranked best if
non-competing solvers were taken into consideration.  Entries $A|B$
indicate that solver~$A$ ranked first considering sequential
performance, while~$B$ ranked first considering parallel performance.
When the same solver ranked first in both, which is most often the
case, its name appears just once.  Table cell colors (other than
white) indicate the best competing solver performance; no cell color
indicates that the best solver performance was by a non-competing
solver and no competing solvers participated in that division.  Full
division rankings are available from the competition website.

It is interesting to note that the non-competing solvers MathSAT and
Z3 often appear as the best performing solver.  We view this as
confirmation of the industrial significance of the competition
results.  Both solvers are industrial-strength tools, actively
maintained and tuned for applications.  The reasons for non-entry into
the competition are orthogonal to their suitability for the
competition.  Indeed, Z3 has been entered into the competition
competitively in the past, and one can assume that it was trained on
the current benchmarks.  The fact that industrial-strength solvers
perform well in the competition supports the notion that the
benchmarks in the competition are industrially relevant.

\include{table-winners-main}

\include{table-winners-application}

\include{table-winners-unsat-core}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Competition-Wide Ranking}
\label{sec:floc}

Table~\ref{table:competition-wide-ranking} shows the best three
\maintrack solvers according to the competition-wide ranking for
2015--2018.  Also shown is Z3, which ranked in high position each year
but was running hors concours.  Sequential and parallel performance
are reported separately.
%
Note that scores are not directly comparable between different years:
the set of benchmarks changes from one year to the other, and also the
formula to compute the score changed (see
Section~\ref{sec:scoring-floc}).  Moreover, readers should keep in
mind that this ranking---like any method that maps the performance of
solvers for a broad range of logics to a one-dimensional value---is
inherently limited.  We refer to the individual division rankings
(Section~\ref{sec:division-rankings}) for more detailed information
about each solver.

\begin{table}
  \caption{Competition-wide ranking for the \maintrack: best solvers (by year)}
  \label{table:competition-wide-ranking}
  \centering
  \scalebox{.89}{\begin{tabular}{cp{2.2cm}rrp{0cm}cp{2.2cm}rr}
    %\cmidrule[\heavyrulewidth]{1-4} \cmidrule[\heavyrulewidth]{6-9}
    \multicolumn{4}{c}{2015} & & \multicolumn{4}{c}{2016} \\ \cmidrule[\heavyrulewidth]{1-4} \cmidrule[\heavyrulewidth]{6-9}
    Rank & Solver & Score  & Score  & & Rank & Solver & Score & Score \\
         &        & (sequential) & (parallel) & & & & (sequential) & (parallel) \\ \cmidrule{1-4} \cmidrule{6-9}
         & [Z3]              & 159.36 & 159.36 & &    & [Z3]   & 185.09 & 185.09 \\
      1. & CVC4              & 144.67 & 144.74 & & 1. & CVC4   & 180.95 & 181.19 \\
      2. & CVC4(exp)\footnotemark & 140.47 & 140.51 & & 2. & Yices & 119.29 & 119.29 \\
      3. & Yices             & 101.91 & 101.91 & & 3. & veriT  &  75.11 &  75.11 \\ \cmidrule[\heavyrulewidth]{1-4} \cmidrule[\heavyrulewidth]{6-9}
  \end{tabular}}

  \medskip

  \scalebox{.89}{\begin{tabular}{cp{2.2cm}rrp{0cm}cp{2.2cm}rr}
    %\cmidrule[\heavyrulewidth]{1-4} \cmidrule[\heavyrulewidth]{6-9}
    \multicolumn{4}{c}{2017} & & \multicolumn{4}{c}{2018} \\ \cmidrule[\heavyrulewidth]{1-4} \cmidrule[\heavyrulewidth]{6-9}
    Rank & Solver & Score  & Score  & & Rank & Solver & Score & Score \\
         &        & (sequential) & (parallel) & & & & (sequential) & (parallel) \\ \cmidrule{1-4} \cmidrule{6-9}
         & [Z3]        & 171.99 & 171.99 & & 1. & CVC4        & 211.99 & 211.99 \\
      1. & CVC4        & 161.38 & 161.76 & &    & [Z3]        & 186.19 & 186.19 \\
      2. & Yices      & 110.63 & 110.63 & & 2. & Yices      & 115.26 & 115.26 \\
      3. & SMTInterpol &  65.96 &  66.00 & & 3. & SMTInterpol &  65.32 &  65.38 \\ \cmidrule[\heavyrulewidth]{1-4} \cmidrule[\heavyrulewidth]{6-9}
  \end{tabular}}
\end{table}
\footnotetext{This was a distinct version of CVC4 that was designated as experimental.}

The 2018 edition of the competition was part of the FLoC Olympic
Games, which sponsored seven medals.  The competition organizers
awarded the following solvers:
\begin{itemize}
\item The best three \maintrack solvers according to the
  competition-wide ranking (CVC4, Yices, and SMTInterpol; see
  Table~\ref{table:competition-wide-ranking}),
\item the solvers that won the QF\_BV division in the \maintrack
  (Boolector and MinkeyRink; see Table~\ref{tab:winners-main}),
\item the solver that won the most competitive divisions in the
  \apptrack (Yices; see Table~\ref{tab:winners-application}),
\item the solver that won the most competitive divisions in the
  \ucoretrack (Yices; see Table~\ref{tab:winners-unsat-core}).
\end{itemize}

In comparison to the competition-wide ranking for the \maintrack, the
metric used for the \apptrack and \ucoretrack---simply counting the
number of competitive divisions that a solver has won---was rather
unsophisticated.  Future editions of the competition should perhaps
refine this to use a ranking scheme that is more similar to the
\maintrack ranking, by weighing divisions according to their relative
importance, and also by rewarding good performance from solvers other
than division winners.  A difficulty in adapting the \maintrack
ranking for the other tracks is that each \maintrack benchmark is
worth one point (before benchmark weights are applied) if solved
correctly, while benchmarks in the \apptrack and \ucoretrack are worth
widely varying numbers of points.  Moreover, in the \ucoretrack, the
size of the smallest unsatisfiable core---and hence the number of
points attainable on any given benchmark---may not be known precisely.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
  \centering
  \scalebox{.45}{\includegraphics{%
        plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2015-2018.pdf}}
  \caption{VBS on common benchmarks 2015--2018 of all divisions in the
    \maintrack with a time limit of \SI{1200}{s}.  Number of common
    benchmarks is indicated with a gray vertical line.}
  \label{fig:progress-all}
\end{figure}

\begin{figure}
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
        plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2014-2015.pdf}}
    \caption{2014 vs.~2015}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
      plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2015-2016.pdf}}
    \caption{2015 vs.~2016}
  \end{subfigure}
  \\[2ex]
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
      plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2016-2017.pdf}}
    \caption{2016 vs.~2017}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \scalebox{.35}{\includegraphics{%
    plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2017-2018.pdf}}
    \caption{2017 vs.~2018}
  \end{subfigure}
  \caption{VBS on common benchmarks for pairs of years of all \maintrack
  divisions with a time limit of \SI{1200}{s}.  Number of common benchmarks is
  indicated with a gray vertical line.}
  \label{fig:progress}
\end{figure}

\begin{figure}
  \centering
  \begin{minipage}{0.47\textwidth}
    \centering
    \scalebox{.31}{\includegraphics{%
        plot/pdf/bar_plot-Virtual_Best_Solver_(VBS)_Application_Track_2015-2018.pdf}}
    \vspace{-5ex}
    \caption{VBS in terms of correct \texttt{check-sat} answers (out
      of \num{30370517} total) within a time limit of \SI{2400}{s} on
      common benchmarks 2015--2018 of all divisions in the \apptrack.}
    \label{fig:progress-all-app}
  \end{minipage}\hfill
  \begin{minipage}{0.47\textwidth}
    \centering
    \scalebox{.31}{\includegraphics{%
        plot/pdf/bar_plot-Virtual_Best_Solver_(VBS)_Unsat_Core_Track_2016-2018.pdf}}
    \vspace{-5ex}
    \caption{VBS in terms of number of unsat core reductions within a
      time limit of \SI{2400}{s} on common benchmarks 2016--2018 of
      all divisions in the \ucoretrack.\\${}$}
    \label{fig:progress-all-uc}
  \end{minipage}
\end{figure}

\section{Further Analysis}
\label{sec:analysis}

Beyond the competition rankings, the competition data provides ample
opportunity for further analysis.  Here, we report on several
additional aspects: progress in solver performance, number of uniquely
solved benchmarks, use of parallelism in SMT solvers, performance on
satisfiable versus unsatisfiable benchmarks, performance for a low
time limit, and the impact of benchmark families and of benchmarks
with unknown status.

\subsection{Progress in Solver Performance}

The SMT Competition consists of several tracks with a multitude of
divisions per track.  Not all participating solvers enter all tracks,
and the majority of solvers participates only in a handful of
divisions.  To give a measure of overall progress in solver
performance we use, for each of the three tracks, the notion of a
\emph{virtual best solver} (VBS) over all divisions.

For the \maintrack, we determine the VBS as the best wall-clock
performance per (correctly solved) instance with the time limit of
\SI{1200}{s} used in the competitions 2017 and 2018.
Figure~\ref{fig:progress-all} visualizes VBS performance on common
benchmarks from 2015--2018 as cactus plot.  Figure~\ref{fig:progress}
shows VBS performance on common benchmarks over pairs of consecutive
years from 2014 to 2018.  Both plots show the wall-clock solving time
per instance over all instances, sorted by solving time.
%
Between 2014 (\num{67426} benchmarks) and~2015 (\num{154238}
benchmarks), the number of common benchmarks is \num{67070}; between
2015 and 2016 (\num{154424} benchmarks) it is \num{154238}; between
2016 and 2017 (\num{238758} benchmarks) it is \num{145236}; and
between 2017 and 2018 (\num{333241} benchmarks) it is \num{226429}.
The number of solved instances on these sets of common benchmarks
increased from 2014 to 2015 by 0.1\%, from 2016 to 2017 by 3.9\%, and
from 2017 to 2018 by 0.3\%.  Interestingly, it decreased from 2015 to
2016 by 3.5\%.  This is mainly due to the fact that, in 2015, more
than \num{5600} benchmarks of division QF\_FP were solved by one of
the two participating configurations of Z3 but by no other solver, and
in~2016 Z3 did not participate in this division.

For the \apptrack, we determine the VBS as the best performance in
terms of number of correct \texttt{check-sat} answers within the time
limit of \SI{2400}{s} (the time limit of the \apptrack in the
competitions 2015--2018).  The total number of eligible
\texttt{check-sat} queries (i.e., all queries with status sat or unsat
up to the first query with status unknown) on common benchmarks from
2015--2018 is \num{30370517}.  Figure~\ref{fig:progress-all-app}
visualizes VBS performance on these benchmarks as a bar plot.  In
2015, already 99.95\% of the eligible queries were answered correctly,
which naturally does not leave a lot of room for improvement.  In
2016, performance decreased by \num{185625} correct answers (0.61\%),
which can be entirely attributed to a performance regression of Z3 on
two benchmarks in the QF\_NIA division.

For the \ucoretrack, we determine the VBS as the best performance in
terms of reduction, i.e., the difference in the number of assertions
of the input formula and the unsat core, within the time limit of
\SI{2400}{s} (the time limit of the \ucoretrack in the competitions
2016--2018).  Figure~\ref{fig:progress-all-uc} visualizes VBS
performance on the common benchmarks (\num{91455} total) from
2016--2018 as a bar plot.  For \num{32377} of these benchmarks,
extracting an unsat core is trivial since they contain only one single
\texttt{assert} command.  They are thus not considered in the bar
plot, since their maximum possible reduction is trivially 0.

\subsection{Unique Solutions}

For each SMT solver, the number of uniquely solved benchmarks---that
is, the number of benchmarks that were solved by this solver only---is
a measure of how much the solver contributes to the state-of-the-art,
and the extent to which it is complementary to other solvers.  In
particular, if a solver that uniquely solves~$n$ benchmarks had not
participated in the competition, the virtual best solver would have
been able to solve~$n$ fewer benchmarks.
Table~\ref{table:unique-solutions} shows the number of uniquely solved
benchmarks for each \maintrack solver from 2015--2018.  (Multiple
versions of a solver are consolidated in this table, and answers that
were given by unsound solver versions are excluded.)

Solvers that rank in high position in the competition-wide ranking
(Section~\ref{sec:floc}) often also solve many benchmarks uniquely,
but there are exceptions.  For instance, SMTInterpol has few unique
solutions, despite achieving third place in the competition-wide
ranking in~2017 and~2018.  However, almost every (sound) solver in the
competition solves at least some benchmarks uniquely, and hence
contributes to the state-of-the-art.

The largest numbers in Table~\ref{table:unique-solutions} are due to
divisions in which there was only one sound solver: in~2016, MathSAT
was the only solver to support QF\_FP; in~2017, Z3 was the only sound
solver in QF\_FP; and in~2018, CVC4 was the only solver to support
QF\_SLIA, as well as the only sound solver in QF\_ABVFP.  Note that
divisions with fewer than two competing solvers are non-competitive
and do not affect the competition-wide ranking.

\begin{table}
  \caption{Number of \maintrack benchmarks uniquely solved by each
    solver (by year)}
  \label{table:unique-solutions}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    Solver & 2015 & 2016 & 2017 & 2018 \\
    \midrule
    ABC               &       & 7     &       &        \\
    Alt-Ergo          &       &       &       & 1      \\
    AProVE            & 5     & 0     & 413   & 553    \\
    Boolector         & 37    & 25    & 97    & 62     \\
    COLIBRI           &       &       & 3     & 22     \\
    Ctrl-Ergo         &       &       &       & 12     \\
    CVC3              & 70    &       &       &        \\
    CVC4              & 4863  & 113   & 10258 & 100679 \\
    MapleSTP          &       & 0     &       &        \\
    {[}MathSAT{]}     & 21    & 28745 & 11    & 5      \\
    MinkeyRink        &       & 13    & 27    & 16     \\
    OpenSMT2          & 0     & 0     & 0     & 0      \\
    ProB              &       & 2     &       &        \\
    Q3B               &       & 0     & 89    & 0      \\
    raSAT             & 0     & 14    &       &        \\
    Redlog            &       &       & 4     &        \\
    SMT-RAT           & 12    & 0     & 85    & 2      \\
    SMTInterpol       & 1     & 1     & 5     & 4      \\
    SPASS-SATT        &       &       &       & 34     \\
    STP               & 4     & 1     & 4     & 5      \\
    toysmt            &       & 0     &       &        \\
    Vampire           &       & 130   & 753   & 652    \\
    veriT             & 7     & 19    & 55    & 57     \\
    XSat              &       &       & 0     &        \\
    Yices            & 86    & 128   & 2379  & 1535   \\
    Z3                & 6118  & 244   & 40992 & 888    \\
    \midrule
    Total             & 11224 & 29442 & 55175 & 104527 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Use of Parallelism}

It is evident both from the division rankings
(Section~\ref{sec:division-rankings}) and the competition-wide ranking
(Section~\ref{sec:floc}) that there is very little difference between
sequential and parallel performance.  Even though parallel solvers may
use up to four times as much CPU time, sequential and parallel
division winners are most often identical, as are sequential and
parallel scores for the competition-wide ranking.

In Table~\ref{table:parallelism}, we report the quotient of CPU time
over wall-clock time (summed up over all benchmarks) for each
\maintrack solver, as well as for all solvers combined.  When there
were multiple versions of a solver (such as a sequential and a
parallel version), only the highest value is shown.  We note that this
quotient is close to~1 for most solvers: only about 4--8 \maintrack
solvers each year were participating with a version that made use of
parallelism.

\begin{table}
  \caption{Quotient of CPU time over wall-clock time for each
    \maintrack solver (by year).  Values below~1.1 are shown in gray.}
  \label{table:parallelism}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    Solver & 2015 & 2016 & 2017 & 2018 \\
    \midrule
    ABC               &     & \textcolor{gray}{1.0} &     &     \\
    Alt-Ergo          &     &     &     & 3.0 \\
    AProVE            & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{0.8} & \textcolor{gray}{0.8} \\
    Boolector         & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & 1.1 & 1.3 \\
    COLIBRI           &     &     & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} \\
    Ctrl-Ergo         &     &     &     & 4.0 \\
    CVC3              & \textcolor{gray}{1.0} &     &     &     \\
    CVC4              & 1.2 & 1.1 & 1.1 & \textcolor{gray}{1.0} \\
    MapleSTP          &     & 3.8 &     &     \\
    {[}MathSAT{]}     & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} \\
    MinkeyRink        &     & 3.6 & 2.0 & 3.9 \\
    OpenSMT2          & 3.7 & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} \\
    ProB              &     & \textcolor{gray}{1.0} &     &     \\
    Q3B               &     & 2.9 & 2.9 & 3.0 \\
    raSAT             & \textcolor{gray}{1.0} & 2.0 &     &     \\
    Redlog            &     &     & \textcolor{gray}{1.0} &     \\
    SMT-RAT           & 1.3 & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} \\
    SMTInterpol       & 1.4 & 1.1 & 1.1 & 1.1 \\
    SPASS-SATT        &     &     &     & \textcolor{gray}{1.0} \\
    STP               & \textcolor{gray}{1.0} & 3.6 & 3.8 & 3.8 \\
    toysmt            &     & \textcolor{gray}{1.0} &     &     \\
    Vampire           &     & 3.4 & 3.9 & 4.0 \\
    veriT             & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} \\
    XSat              &     &     & 1.8 &     \\
    Yices            & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} \\
    Z3                & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} & \textcolor{gray}{1.0} \\
    \midrule
    All solvers       & 1.0 & 1.3 & 1.3 & 1.3 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Focusing on Satisfiability or Unsatisfiability}

In many cases, an SMT application produces either mainly satisfiable
or mainly unsatisfiable queries to the SMT solver.  A typical example
is bounded model checking, where a query is only satisfiable if a
property of the model does not hold.  As a consequence, some solvers
may implement specialized techniques and optimizations for the more
frequent (the expected) case.
%
Therefore, it is interesting to ask whether the winners of divisions
of a track might have changed if benchmarks in that division had been
restricted to one of these results.
Tables~\ref{tab:results:sat}--\ref{tab:results:unsat} show the changes
in division winners if only satisfiable or unsatisfiable results are
considered.  These tables only lists the result for a division if it
has changed with respect to Table~\ref{tab:winners-main}.  A
non-competing solver is listed by itself if there was no competing
solver with a positive division score under the given constraints.  It
is interesting to note that we see some new solvers winning divisions
that did not win any division in the main results.  For example, veriT
solves the most satisfiable benchmarks in QF\_UFIDL in 2018 and the
most unsatisfiable benchmarks in all four years in UFLRA, and
SMTInterpol performs very well on unsatisfiable benchmarks in QF\_LIA.

We only consider the \maintrack here, which contains a mixture of
benchmarks that are satisfiable and unsatisfiable.  A percentage of
satisfiable, unsatisfiable and unknown status benchmarks per division
and year is given in Table~\ref{table:benchmarks-main-track}.  For the
\apptrack, this analysis is less meaningful since a benchmark usually
contains a multitude of both satisfiable and unsatisfiable incremental
queries (which may depend on each other).  For the \ucoretrack, only
known unsatisfiable benchmarks from the \maintrack are used, which
renders this analysis useless for this track.

\input{table-winners-main-sat-unsat}

%%%%%%%%%%%%%%%%%%%%

\subsection{Very Quick Solutions}

Different application domains of SMT typically impose a wide range of
time limits (from hours to seconds). In domains with very low time
limits, these limits are, in the incremental case, usually even
defined per incremental query.  With 20 and 40 minutes, the time
limits used in the competition for all tracks in all four years were
relatively long and the same for all benchmarks, thus agnostic to the
application domain of a benchmark.  Hence, it is interesting to
investigate how the results change if we consider a much shorter time
limit.  Table~\ref{tab:results:24s} shows the results of virtually
imposing a time limit of 24 seconds\footnote{This time limit is used
  for the new 24 seconds score introduced in SMT-COMP 2019 and was
  inspired by a use case at Amazon.} in the \maintrack.  As in the
previous analysis, this highlights the different winners under this
new scheme only. Here we see Alt-Ergo (AUFLIRA in 2018) and STP
(QF\_BV in 2016) winning a division for the first time in our
analysis, and Yices performing well.

We only consider the \maintrack here, since for the other two tracks,
the results are less interesting.  In the \apptrack, for such low time
limits we are rather interested in a time limit per query, which we
are unable to impose on the existing data since it is not possible to
extract wallclock run time information per incremental query from the
data of the competitions (as mentioned above, StarExec does not record
this information).  In the \ucoretrack, the main focus is on producing
small unsat-cores and solving times are of secondary interest.
Additionally, benchmarks with known unsatisfiable status from the
\maintrack are used, and thus these results are already available from
the \maintrack analysis in Table~\ref{tab:results:24s} (disregarding
the fact that enabling unsat core production may slow down a solver,
depending on the technique).

\input{table-winners-main-24s}

%%%%%%%%%%%%%%%%%

\subsection{Do Families Matter?}

In 2016 the \maintrack competition scoring was changed to weight the
correctness score for benchmarks by the size of the family the
benchmark belongs to. Here we ask what the winners would have been if
this change had not been made. The answer to this question is given in
Table~\ref{tab:results:total:solved}. Relatively few changes would
have occurred. Unsurprisingly, the differences are in the larger and
more structured divisions.

Whilst producing these results we noticed that the description of what
constitutes a family had been incorrectly interpreted by the scoring
scripts in 2016--2018. The scoring scripts interpreted the bottom most
directory containing benchmarks in a logic as family, while the rules
stated that the top most subdirectory in a logic is to be interpreted
as family. After fixing this misinterpretation we note that only five
division winners change---in 2018, division AUFLIRA should have been
won by Vampire instead of CVC4; in 2016, AUFNIRA should have been won
by CVC4 instead of Vampire; in 2016 and 2017, QF\_LIA should have been
won by Yices instead of CVC4; and in 2018, QF\_NRA should have been
won by CVC4 instead of Yices.  Table~\ref{tab:results:total:top} shows
the results in the Main Track when the top most subdirectory in a
logic is interpreted as family in reference to the competition results
in Table~\ref{tab:winners-main}, where the bottom most directory was
interpreted as family.

The combination of the small changes in winners and the lack of
difference a misinterpretation of the notion of family made suggests
that families do not have a significant impact on the competition.

\input{table-winners-main-families}

%%%%%%%%%%%%%%%%%

\subsection{What Impact Did Adding Unknowns in the \maintrack Have?}

Until 2016, only benchmarks with known status were used in the
\maintrack. In 2016, benchmarks with unknown status were evaluated
separately, but not considered in the overall results of the track.
Since 2017, non-incremental benchmarks with unknown status were
eligible for the \maintrack.  Table~\ref{tab:results:unknowns} shows
alternative winners for 2017 and 2018 if benchmarks with unknown
status had not been included.

\input{table-winners-main-unknown}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:conclusions}

The SMT Competition was instituted in~2005.  Thirteen years later, it
is a well-established event that garners wide interest from developers
and users of SMT solvers.  The competition continues to contribute to
its original goals, including adoption of the SMT-LIB format and
sparking further advances in SMT, and provides a yardstick of
performance for SMT solvers.

The last FLoC Olympiad was a period of significant growth for the SMT
Competition.  The average number of solver submissions increased
from~12 (for 2005--2014) to~26 (for 2015--2018).  The number of
SMT-LIB benchmarks used in the competition increased from~\num{77352}
in~2014~\cite{CDW14} to \num{342498} in 2018.  This is partly because
new benchmarks were added to the SMT Library, but also because
benchmarks with unknown status are now eligible for the competition,
and because the competition now uses all eligible benchmarks (rather
than a randomly selected subset).

Since the re-introduction of the \ucoretrack in~2016, the competition
again has three separate tracks.  This, together with new logics in
the SMT Library, has brought the total number of competition divisions
to 115 in~2018, up from just 40 divisions in~2014~\cite{CDW14}.  The
number of job pairs increased from \num{339714} in~2014
to~\num{1776062} in 2018, and the total wall-clock time in~2018
exceeded 7.6~years.  By these measures, the SMT Competition is one of
the largest competitions in automated reasoning.

The StarExec infrastructure has been vitally important in allowing the
competition to grow to its current size.  The competition organizers
have given the StarExec developers comprehensive feedback, and filed
dozens of feature requests and bug reports since 2013.  The
infrastructure has become increasingly stable; serious issues are now
rare.  When they did surface during ongoing competitions, they
received a swift response by the StarExec team (under the lead of
Aaron Stump).  This has been critical to the success of the
competition on more than one occasion.

Despite the substantial amount of computational resources available to
the competition through StarExec, the competition has reached a size
where future organizers will have to carefully consider the number of
participants and benchmarks each year, to determine time limits that
will allow the computational workload to complete before the SMT
Workshop.

\paragraph{The Next Competition}

Preparations for the 2019 iteration of the competition are underway and a number of changes have already been decided. These include

\begin{itemize}
  \item the introduction of new \emph{additional scores} for
    satisfiable/unsatisfiable benchmarks and benchmarks solved within 24
    seconds in the \maintrack,
	\item the inclusion of {\tt check-sat} commands with unknown status in
  the \apptrack,
  \item the removal of \emph{non-competitive divisions} and the filtering of
    benchmarks to exclude those seen as \emph{trivial} (taken as those solved
    by all competitors in the previous competition in less than 1 second),
  \item the introduction of a new experimental \emph{model validation track}
    with a single QF\_BV division where the models produced by solvers are
    checked,
  \item the introduction of a new \emph{industrial challenge track} consisting
    of unsolved benchmarks submitted specifically for the track,
  \item the replacement of the notion of \emph{competition-wide scoring} with
    awards that do not attempt to compare divisions.
\end{itemize}

\paragraph{The Future}

Looking to the future of the competition, it is clear that further development should be undertaken with care,
keeping in mind that major changes require not just the competition
organizers, but also participants to adapt.

For the inclusion of new tracks or divisions, a two-phase approach has
proven successful.  A new division is first declared
\emph{experimental}, meaning that results will be reported but no
official winner will be announced.  This allows competition organizers
and participants to gain valuable experience, while reducing the
impact of mistakes.  If successful, the division may then become
regular in the following year.

Organizing the competition is a non-trivial amount of work.  For
instance, the competition tools (such as the benchmark scrambler and
trace executor) often require updating when new divisions are added,
or when changes are made to StarExec.  One could imagine adding
further tracks to the competition (e.g., for proof-producing solvers,
or for variants such as max-SMT), but the effort required to support
such tracks will need to be balanced against their added value.

To ensure that the competition results are practically relevant, it is
important that the SMT Library contains benchmarks that cover a wide
range of applications.  To this end, we again encourage users of SMT
solvers to submit new and interesting benchmarks to the library,
especially for underrepresented logics.  This is not just a valuable
contribution to the community, but also---since these benchmarks will
then be used in the competition---a straightforward way to incentivize
solver developers to tune their solvers accordingly.

\paragraph{Acknowledgments}

We would like to thank Aaron Stump for allowing the SMT Competition to
run on StarExec; the SMT-LIB maintainers, namely Clark Barrett, Pascal
Fontaine, Aina Niemetz and Mathias Preiner, for preparing the annual
SMT-LIB release; all contributors of new benchmarks to the SMT
Library; and, last but not least, all solver developers who entered
their solver into the competition.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{plainbv}
\bibliography{SMT-COMP}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Local Variables:
% ispell-local-dictionary: "american"
% mode: LaTeX
% mode: flyspell
% LocalWords: concours hors logics satisfiability SMT Tjark
% End: