### 2019-08-31 2018/report/table-winners-main-unknown.tex
Remove 2015 and 2016 columns

\begin{table}
  \caption{Alternative best solvers in the \maintrack if benchmarks with unknown status had not been included in 2017 onwards}
  \label{tab:results:unknowns}
  \centering
  \scalebox{.755}{%
    \renewcommand{\arraystretch}{1.01}%
\begin{tabular}{r>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l>{\columncolor{white}[.5em][.8em]}r@{\hskip .8em}|@{\hskip .5em}>{\columncolor{white}[.5em][.5em]}l}
\toprule
Division &                  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2017} &                 \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{2018} \\
\midrule
AUFLIRA  &                                \cc{cvc4} CVC4 [Z3] &  \hspace*{3em}         &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
AUFNIRA  &        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
BV       &          \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{q3b} Q3B} &  \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{bool} Boolector} \\
LRA      &        \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
NRA      &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                                   & \cc{vamp} Vampire [Z3] \\
QF\_FP   &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &       \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{cvc4} CVC4} \\
QF\_NRA  & \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{\cc{yices} Yices [Z3]} &                     \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} \\
UF       &                                                           \cc{cvc4} CVC4 &  &                                                           & \cc{cvc4} CVC4 \\
UFNIA    &                      \multicolumn{2}{>{\columncolor{white}[.5em][.5em]}c}{} &                                                     \cc{cvc4} CVC4 [Z3] &  \\
\bottomrule
\end{tabular}
  }
\end{table}