### 2023-07-05 2023/slides/presresults/input_for_certificates.tex
Start of presentation result 2023

\MakeOnePage{COLIBRI}{}{}{\withtrack{QF\_ABVFPLRA}{\seq,\paral,\unsat,\fast}, \withtrack{QF\_FP}{\unsat,\fast}, \withtrack{QF\_FPLRA}{\unsat,\fast}}{}{s}
\newpage
\MakeOnePage{Z3++}{\withtrack{Biggest Lead}{\mv}, \withtrack{Largest Contribution}{\mv}}{\withtrack{QF\_LinearIntArith}{\seq,\paral,\sat,\mv}, \withtrack{QF\_NonLinearIntArith}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{QF\_NonLinearRealArith}{\seq,\paral,\sat}}{\withtrack{QF\_IDL}{\unsat}}{s}{}
\newpage
\MakeOnePage{OSTRICH}{}{}{\withtrack{QF\_S}{\seq,\paral,\unsat}}{}{}
\newpage
\MakeOnePage{SMTInterpol}{\withtrack{Biggest Lead}{\fast}}{\withtrack{QF\_Datatypes}{\fast,\uc\textsuperscript\paral}, \withtrack{QF\_Equality+LinearArith}{\seq,\paral,\sat,\inc,\mv}, \withtrack{QF\_NonLinearIntArith}{\inc}}{\withtrack{ALIA}{\sat,\uc}, \withtrack{AUFDTLIA}{\uc}, \withtrack{QF\_ALIA}{\uc}, \withtrack{QF\_ANIA}{\seq,\paral,\sat,\unsat,\fast,\inc,\uc}, \withtrack{QF\_AUFNIA}{\seq,\paral,\sat,\unsat,\fast,\uc}, \withtrack{QF\_LIA}{\unsat}, \withtrack{QF\_UF}{\uc\textsuperscript\paral}, \withtrack{QF\_UFDT}{\uc\textsuperscript\seq}, \withtrack{QF\_UFDTLIA}{\unsat,\fast,\uc}, \withtrack{UFIDL}{\sat}, \withtrack{UFLIA}{\sat}}{s}{s}
\newpage
\MakeOnePage{Bitwuzla}{\withtrack{Biggest Lead}{\inc}}{\withtrack{Bitvec}{\sat,\fast}, \withtrack{Equality+MachineArith}{\inc}, \withtrack{FPArith}{\seq,\paral,\sat,\unsat,\fast,\inc,\uc}, \withtrack{QF\_Bitvec}{\inc}, \withtrack{QF\_Equality+Bitvec}{\seq,\paral,\sat,\unsat,\fast,\inc,\mv}, \withtrack{QF\_FPArith}{\seq,\paral,\sat,\unsat,\fast,\inc}}{\withtrack{ABVFP}{\sat,\fast}, \withtrack{ABVFPLRA}{\fast}, \withtrack{AUFBV}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{AUFBVFP}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{QF\_ABVFP}{\uc}, \withtrack{QF\_ABVFPLRA}{\uc}, \withtrack{QF\_BVFP}{\uc}, \withtrack{QF\_BVFPLRA}{\uc}, \withtrack{UFBV}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{UFBVFP}{\seq,\paral,\unsat}}{s}{s}
\newpage
\MakeOnePage{Vampire}{}{\withtrack{Equality}{\fast}}{\withtrack{ALIA}{\seq,\paral,\unsat,\fast}, \withtrack{AUFLIA}{\unsat,\fast}, \withtrack{AUFLIRA}{\fast}, \withtrack{AUFNIRA}{\paral,\unsat,\fast}, \withtrack{UF}{\seq,\paral,\sat}, \withtrack{UFDTLIA}{\fast,\uc}, \withtrack{UFDTNIA}{\seq,\paral,\unsat,\fast,\uc}}{}{s}
\newpage
\MakeOnePage{Yices2}{}{\withtrack{QF\_Bitvec}{\uc}, \withtrack{QF\_Equality}{\seq,\paral,\sat,\unsat,\fast,\uc,\mv}, \withtrack{QF\_Equality+Bitvec}{\uc}, \withtrack{QF\_Equality+LinearArith}{\fast,\uc}, \withtrack{QF\_Equality+NonLinearArith}{\sat}, \withtrack{QF\_LinearIntArith}{\fast,\inc,\uc}, \withtrack{QF\_LinearRealArith}{\seq,\paral,\fast,\uc}}{\withtrack{QF\_ALIA}{\unsat}, \withtrack{QF\_AUFBV}{\fast}, \withtrack{QF\_AUFBVLIA}{\inc}, \withtrack{QF\_AUFLIA}{\seq,\paral,\sat,\unsat}, \withtrack{QF\_LIRA}{\seq,\paral,\sat,\unsat,\mv}, \withtrack{QF\_RDL}{\sat,\unsat,\mv}, \withtrack{QF\_UFBVLIA}{\inc}, \withtrack{QF\_UFLIA}{\unsat,\inc}, \withtrack{QF\_UFLRA}{\seq,\paral,\sat,\unsat,\inc,\mv}, \withtrack{QF\_UFNRA}{\seq,\paral,\unsat,\fast,\inc}}{s}{s}
\newpage
\MakeOnePage{YicesQS}{}{\withtrack{Arith}{\sat,\fast}}{\withtrack{LRA}{\seq,\paral,\unsat}, \withtrack{NRA}{\seq,\paral,\unsat}}{}{s}
\newpage
\MakeOnePage{STP}{}{\withtrack{QF\_Bitvec}{\seq,\paral,\sat,\unsat,\fast,\mv}}{}{}{}
\newpage
\MakeOnePage{cvc5}{\withtrack{Biggest Lead}{\seq,\paral,\sat,\unsat,\uc}, \withtrack{Largest Contribution}{\seq,\paral,\sat,\unsat,\fast,\inc,\uc}}{\withtrack{Arith}{\seq,\paral,\unsat,\inc,\uc}, \withtrack{Bitvec}{\seq,\paral,\unsat,\inc,\uc}, \withtrack{Equality}{\seq,\paral,\sat,\unsat,\inc,\uc}, \withtrack{Equality+LinearArith}{\seq,\paral,\sat,\unsat,\fast,\inc,\uc}, \withtrack{Equality+MachineArith}{\seq,\paral,\sat,\unsat,\fast,\uc}, \withtrack{Equality+NonLinearArith}{\seq,\paral,\sat,\unsat,\fast,\inc,\uc}, \withtrack{QF\_Datatypes}{\seq,\paral,\sat,\unsat,\uc\textsuperscript\seq}, \withtrack{QF\_Equality}{\inc}, \withtrack{QF\_Equality+Bitvec+Arith}{\inc}, \withtrack{QF\_Equality+NonLinearArith}{\seq,\paral,\unsat,\fast,\inc,\uc}, \withtrack{QF\_FPArith}{\uc}, \withtrack{QF\_LinearIntArith}{\unsat}, \withtrack{QF\_LinearRealArith}{\unsat}, \withtrack{QF\_NonLinearRealArith}{\unsat,\fast}, \withtrack{QF\_Strings}{\seq,\paral,\sat,\unsat}}{\withtrack{LIA}{\sat,\fast}, \withtrack{NIA}{\sat,\fast}, \withtrack{QF\_ALIA}{\inc}, \withtrack{QF\_AUFLIA}{\uc}, \withtrack{QF\_BVFPLRA}{\inc}, \withtrack{QF\_FP}{\sat}, \withtrack{QF\_IDL}{\uc}, \withtrack{QF\_NIRA}{\seq,\paral,\unsat}, \withtrack{QF\_UFBVDT}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{QF\_UFDTLIRA}{\seq,\paral,\sat,\unsat,\fast,\uc}, \withtrack{QF\_UFLRA}{\uc}, \withtrack{QF\_UFNIA}{\sat}, \withtrack{UFDT}{\fast}}{s}{s}
\newpage
\MakeOnePage{OpenSMT}{}{\withtrack{QF\_Equality+LinearArith}{\unsat}, \withtrack{QF\_LinearRealArith}{\sat,\inc,\mv}}{\withtrack{QF\_LIA}{\seq,\paral}, \withtrack{QF\_LRA}{\seq,\paral,\unsat,\fast}, \withtrack{QF\_UFIDL}{\seq,\paral,\sat,\mv}}{s}{s}
\newpage
\MakeOnePage{z3-alpha}{}{\withtrack{QF\_Strings}{\fast}}{\withtrack{QF\_S}{\sat}, \withtrack{QF\_SNIA}{\seq,\paral,\sat}}{}{s}
\newpage
\MakeOnePage{iProver}{}{}{\withtrack{ANIA}{\unsat}}{}{}
\newpage