### 2024-07-18 2024/slides/presresults/input_for_certificates.tex
SMT Workshop 2024 slides draft.

\MakeOnePage{Algaroba}{}{\withtrack{QF\_Datatypes}{\seq,\paral,\fast}}{}{}{}
\newpage
\MakeOnePage{Amaya}{}{}{\withtrack{NIA}{\fast}}{}{}
\newpage
\MakeOnePage{Bitwuzla}{}{\withtrack{Bitvec}{\sat}, \withtrack{FPArith}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{QF\_Bitvec}{\seq,\paral,\unsat,\fast}, \withtrack{QF\_Equality\_Bitvec}{\seq,\paral,\sat,\unsat}, \withtrack{QF\_FPArith}{\seq,\paral,\sat,\unsat,\fast}}{\withtrack{ABVFP}{\sat}, \withtrack{ABVFPLRA}{\sat}, \withtrack{AUFBV}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{AUFBVFP}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{QF\_UFBV}{\fast}, \withtrack{UFBV}{\fast}, \withtrack{UFBVFP}{\seq,\paral,\unsat,\fast}}{s}{s}
\newpage
\MakeOnePage{COLIBRI}{}{}{\withtrack{QF\_FP}{\fast}, \withtrack{QF\_FPLRA}{\fast}}{}{s}
\newpage
\MakeOnePage{cvc5}{\withtrack{Biggest Lead}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{Largest Contribution}{\seq,\paral,\sat,\unsat,\fast}}{\withtrack{Arith}{\seq,\paral,\unsat}, \withtrack{Bitvec}{\seq,\paral,\unsat}, \withtrack{Equality}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{Equality\_LinearArith}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{Equality\_MachineArith}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{Equality\_NonLinearArith}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{QF\_Datatypes}{\sat}, \withtrack{QF\_Equality}{\sat}, \withtrack{QF\_Equality\_NonLinearArith}{\unsat}, \withtrack{QF\_LinearIntArith}{\unsat}, \withtrack{QF\_LinearRealArith}{\unsat}, \withtrack{QF\_NonLinearRealArith}{\unsat}}{\withtrack{FP}{\fast}, \withtrack{LIA}{\sat,\fast}, \withtrack{NIA}{\sat}, \withtrack{NRA}{\sat}, \withtrack{QF\_ABVFPLRA}{\seq,\paral,\sat,\unsat}, \withtrack{QF\_AUFNIA}{\seq,\paral,\fast}, \withtrack{QF\_AX}{\fast}, \withtrack{QF\_FP}{\sat}, \withtrack{QF\_LIRA}{\fast}, \withtrack{QF\_UFBVDT}{\seq,\paral,\fast}, \withtrack{QF\_UFDT}{\seq,\paral}, \withtrack{QF\_UFDTNIA}{\seq,\paral,\sat,\fast}, \withtrack{QF\_UFFPDTNIRA}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{QF\_UFLRA}{\unsat}}{s}{s}
\newpage
\MakeOnePage{iProver v3.9}{}{}{\withtrack{ALIA}{\paral,\unsat}, \withtrack{UF}{\fast}, \withtrack{UFDTNIA}{\paral,\unsat}, \withtrack{UFLRA}{\seq,\paral,\unsat,\fast}, \withtrack{UFNIRA}{\seq,\unsat}}{}{s}
\newpage
\MakeOnePage{OpenSMT}{}{\withtrack{QF\_Equality\_LinearArith}{\unsat}, \withtrack{QF\_LinearIntArith}{\seq,\paral}, \withtrack{QF\_LinearRealArith}{\seq,\paral,\sat}}{\withtrack{QF\_AX}{\seq,\paral,\sat}, \withtrack{QF\_LIA}{\sat}, \withtrack{QF\_LRA}{\unsat,\fast}, \withtrack{QF\_UFIDL}{\seq,\paral,\sat}}{s}{s}
\newpage
\MakeOnePage{OSTRICH}{}{}{\withtrack{QF\_SNIA}{\seq,\paral,\sat,\fast}}{}{}
\newpage
\MakeOnePage{SMT-RAT}{}{}{\withtrack{NRA}{\unsat,\fast}}{}{}
\newpage
\MakeOnePage{SMTInterpol}{}{\withtrack{QF\_Equality\_LinearArith}{\seq,\paral,\sat}}{\withtrack{ALIA}{\sat}, \withtrack{ANIA}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{QF\_ANIA}{\unsat,\fast}, \withtrack{QF\_AUFLIA}{\fast}, \withtrack{QF\_AX}{\unsat}, \withtrack{QF\_LIA}{\unsat}, \withtrack{QF\_LIRA}{\sat}, \withtrack{QF\_NIRA}{\seq,\paral,\unsat}, \withtrack{QF\_UF}{\sat}, \withtrack{QF\_UFBVDT}{\sat,\unsat}, \withtrack{QF\_UFDTLIA}{\unsat,\fast}, \withtrack{QF\_UFDTLIRA}{\unsat,\fast}, \withtrack{QF\_UFLIA}{\unsat}, \withtrack{UFLIA}{\sat}}{}{s}
\newpage
\MakeOnePage{STP}{}{\withtrack{QF\_Bitvec}{\sat}}{}{}{}
\newpage
\MakeOnePage{Yices2}{}{\withtrack{QF\_Equality}{\fast}, \withtrack{QF\_Equality\_Bitvec}{\fast}, \withtrack{QF\_Equality\_LinearArith}{\fast}, \withtrack{QF\_Equality\_NonLinearArith}{\seq,\paral,\sat,\fast}, \withtrack{QF\_LinearIntArith}{\fast}, \withtrack{QF\_LinearRealArith}{\fast}, \withtrack{QF\_NonLinearIntArith}{\fast}, \withtrack{QF\_NonLinearRealArith}{\fast}}{\withtrack{QF\_LIRA}{\seq,\paral,\unsat}, \withtrack{QF\_RDL}{\seq,\paral,\sat}, \withtrack{QF\_UFLRA}{\seq,\paral}, \withtrack{QF\_UFNRA}{\unsat}}{s}{s}
\newpage
\MakeOnePage{YicesQS}{}{\withtrack{Arith}{\sat,\fast}, \withtrack{Bitvec}{\fast}}{\withtrack{LRA}{\seq,\paral,\unsat}, \withtrack{NRA}{\seq,\paral}}{s}{s}
\newpage
\MakeOnePage{Z3-alpha}{}{\withtrack{QF\_Datatypes}{\unsat}, \withtrack{QF\_Equality}{\seq,\paral,\unsat}, \withtrack{QF\_LinearIntArith}{\sat}, \withtrack{QF\_NonLinearIntArith}{\seq,\paral,\sat,\unsat}, \withtrack{QF\_NonLinearRealArith}{\seq,\paral,\sat}}{\withtrack{QF\_AUFBV}{\unsat}, \withtrack{QF\_IDL}{\seq,\paral,\unsat}, \withtrack{QF\_NIRA}{\fast}, \withtrack{QF\_RDL}{\unsat}}{s}{s}
\newpage
\MakeOnePage{Z3-Noodler}{}{\withtrack{QF\_Strings}{\seq,\paral,\sat,\unsat,\fast}}{}{}{}
\newpage