### 2024-07-19 2024/slides/presresults/input_for_certificates.tex
Updated slides with QF_BVFP incr. results.

\MakeOnePage{Algaroba}{}{\withtrack{QF\_Datatypes}{\seq,\paral,\fast}}{\withtrack{QF\_DT}{\sat}}{}{}
\newpage
\MakeOnePage{Amaya}{}{}{\withtrack{NIA}{\fast}}{}{}
\newpage
\MakeOnePage{Bitwuzla}{\withtrack{Biggest Lead}{\inc}}{\withtrack{Bitvec}{\sat}, \withtrack{Equality\_MachineArith}{\inc,\uc}, \withtrack{FPArith}{\seq,\paral,\sat,\unsat,\fast,\inc}, \withtrack{QF\_ADT\_BitVec}{\mv}, \withtrack{QF\_Bitvec}{\seq,\paral,\unsat,\fast,\inc}, \withtrack{QF\_Equality\_Bitvec}{\seq,\paral,\sat,\unsat,\inc}, \withtrack{QF\_FPArith}{\seq,\paral,\sat,\unsat,\fast,\inc,\uc}}{\withtrack{ABVFP}{\sat}, \withtrack{AUFBV}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{AUFBVFP}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{BVFPLRA}{\uc}, \withtrack{QF\_ABV}{\uc\textsuperscript\seq}, \withtrack{QF\_ABVFP}{\mv}, \withtrack{QF\_BVFPLRA}{\mv}, \withtrack{QF\_FPLRA}{\mv}, \withtrack{QF\_UFBV}{\fast}, \withtrack{UFBV}{\fast}}{s}{s}
\newpage
\MakeOnePage{COLIBRI}{}{}{\withtrack{QF\_ABVFPLRA}{\unsat}, \withtrack{QF\_FP}{\fast}, \withtrack{QF\_FPLRA}{\unsat,\fast}}{}{s}
\newpage
\MakeOnePage{cvc5}{\withtrack{Biggest Lead}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{Largest Contribution}{\seq,\paral,\sat,\unsat,\fast}}{\withtrack{Arith}{\seq,\paral,\unsat,\inc}, \withtrack{Bitvec}{\seq,\paral,\unsat,\inc,\uc}, \withtrack{Equality}{\seq,\paral,\sat,\unsat,\fast,\inc}, \withtrack{Equality\_LinearArith}{\seq,\paral,\sat,\unsat,\fast,\inc}, \withtrack{Equality\_MachineArith}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{Equality\_NonLinearArith}{\seq,\paral,\sat,\unsat,\fast,\inc}, \withtrack{FPArith}{\uc}, \withtrack{QF\_Datatypes}{\sat,\uc}, \withtrack{QF\_Equality\_LinearArith}{\uc\textsuperscript\paral}, \withtrack{QF\_Equality\_NonLinearArith}{\unsat,\inc,\uc\textsuperscript\seq,\mv}, \withtrack{QF\_FPArith}{\mv}, \withtrack{QF\_LinearIntArith}{\unsat}, \withtrack{QF\_LinearRealArith}{\unsat}, \withtrack{QF\_NonLinearRealArith}{\unsat}, \withtrack{QF\_Strings}{\uc}}{\withtrack{ANIA}{\uc}, \withtrack{AUFBV}{\uc\textsuperscript\paral}, \withtrack{AUFBVDTLIA}{\uc}, \withtrack{AUFBVDTNIRA}{\uc}, \withtrack{AUFDTLIRA}{\uc\textsuperscript\seq}, \withtrack{AUFFPDTNIRA}{\uc}, \withtrack{AUFLIRA}{\uc\textsuperscript\seq}, \withtrack{AUFNIA}{\uc\textsuperscript\paral}, \withtrack{BVFPLRA}{\unsat}, \withtrack{FP}{\fast}, \withtrack{LIA}{\sat,\fast}, \withtrack{NIA}{\sat}, \withtrack{QF\_AUFLIA}{\inc,\uc\textsuperscript\seq}, \withtrack{QF\_AX}{\mv}, \withtrack{QF\_BVFP}{\inc}, \withtrack{QF\_FP}{\sat,\uc}, \withtrack{QF\_IDL}{\uc}, \withtrack{QF\_LIA}{\mv}, \withtrack{QF\_SNIA}{\seq,\paral,\sat,\fast}, \withtrack{QF\_UFBVDT}{\seq,\paral,\unsat,\fast,\uc,\mv}, \withtrack{QF\_UFDT}{\seq,\paral,\mv}, \withtrack{QF\_UFDTLIRA}{\seq,\paral,\sat,\unsat,\fast,\uc\textsuperscript\seq,\mv}, \withtrack{QF\_UFDTNIA}{\seq,\paral,\sat,\fast}, \withtrack{QF\_UFFPDTNIRA}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{QF\_UFLRA}{\inc,\uc\textsuperscript\seq}, \withtrack{UFDT}{\uc\textsuperscript\seq}, \withtrack{UFDTLIA}{\uc\textsuperscript\seq}, \withtrack{UFDTNIA}{\uc\textsuperscript\seq}, \withtrack{UFFPDTNIRA}{\uc}, \withtrack{UFIDL}{\uc\textsuperscript\seq}, \withtrack{UFLRA}{\uc}, \withtrack{UFNIRA}{\uc\textsuperscript\seq}}{s}{s}
\newpage
\MakeOnePage{iProver v3.9}{}{}{\withtrack{ALIA}{\paral,\unsat}, \withtrack{UF}{\fast}, \withtrack{UFNIRA}{\seq,\unsat}}{}{s}
\newpage
\MakeOnePage{OpenSMT}{}{\withtrack{QF\_Equality}{\mv}, \withtrack{QF\_Equality\_LinearArith}{\unsat,\mv}, \withtrack{QF\_LinearIntArith}{\seq,\paral,\mv}, \withtrack{QF\_LinearRealArith}{\seq,\paral,\sat,\inc,\mv}}{\withtrack{QF\_LIA}{\sat}, \withtrack{QF\_LRA}{\unsat,\fast}, \withtrack{QF\_UFIDL}{\seq,\paral,\sat}}{s}{s}
\newpage
\MakeOnePage{SMT-RAT}{}{\withtrack{QF\_NonLinearRealArith}{\mv}}{\withtrack{NRA}{\seq,\paral,\unsat,\fast}}{}{}
\newpage
\MakeOnePage{SMTInterpol}{\withtrack{Biggest Lead}{\uc,\mv}}{\withtrack{Arith}{\uc}, \withtrack{Equality}{\uc\textsuperscript\paral}, \withtrack{Equality\_LinearArith}{\uc}, \withtrack{Equality\_NonLinearArith}{\uc}, \withtrack{QF\_ADT\_LinArith}{\mv}, \withtrack{QF\_Datatypes}{\mv}, \withtrack{QF\_Equality\_LinearArith}{\seq,\paral,\sat,\inc,\uc\textsuperscript\seq}, \withtrack{QF\_Equality\_NonLinearArith}{\uc\textsuperscript\paral}, \withtrack{QF\_NonLinearIntArith}{\inc,\uc}, \withtrack{QF\_NonLinearRealArith}{\uc}}{\withtrack{ALIA}{\sat}, \withtrack{ANIA}{\seq,\paral,\sat,\unsat,\fast}, \withtrack{QF\_ALIA}{\uc\textsuperscript\paral}, \withtrack{QF\_ANIA}{\unsat,\fast,\inc,\uc\textsuperscript\seq,\mv}, \withtrack{QF\_AUFNIA}{\sat,\uc\textsuperscript\seq,\mv}, \withtrack{QF\_LIA}{\unsat}, \withtrack{QF\_LIRA}{\uc\textsuperscript\paral}, \withtrack{QF\_UF}{\uc\textsuperscript\paral}, \withtrack{QF\_UFBVDT}{\sat}, \withtrack{QF\_UFDTLIA}{\unsat,\fast,\uc\textsuperscript\paral}, \withtrack{QF\_UFDTNIA}{\uc\textsuperscript\seq,\mv}, \withtrack{QF\_UFLIA}{\unsat,\mv}, \withtrack{QF\_UFNRA}{\uc\textsuperscript\seq}, \withtrack{UFIDL}{\sat}, \withtrack{UFLIA}{\sat}}{s}{s}
\newpage
\MakeOnePage{STP}{}{\withtrack{QF\_Bitvec}{\sat,\mv}}{}{}{}
\newpage
\MakeOnePage{Yices2}{\withtrack{Largest Contribution}{\uc,\mv}}{\withtrack{Equality}{\uc\textsuperscript\seq}, \withtrack{QF\_Bitvec}{\uc}, \withtrack{QF\_Equality}{\seq,\paral,\sat,\unsat,\fast,\inc,\uc}, \withtrack{QF\_Equality\_Bitvec}{\fast,\uc,\mv}, \withtrack{QF\_Equality\_Bitvec\_Arith}{\inc}, \withtrack{QF\_Equality\_LinearArith}{\fast}, \withtrack{QF\_Equality\_NonLinearArith}{\seq,\paral,\sat,\fast}, \withtrack{QF\_LinearIntArith}{\fast,\inc,\uc}, \withtrack{QF\_LinearRealArith}{\fast,\uc}, \withtrack{QF\_NonLinearIntArith}{\fast,\mv}, \withtrack{QF\_NonLinearRealArith}{\fast}}{\withtrack{QF\_ALIA}{\unsat}, \withtrack{QF\_AUFLIA}{\seq,\paral,\sat,\unsat}, \withtrack{QF\_AUFNIA}{\unsat}, \withtrack{QF\_IDL}{\mv}, \withtrack{QF\_LIRA}{\sat,\mv}, \withtrack{QF\_RDL}{\seq,\paral,\sat,\unsat,\mv}, \withtrack{QF\_UFLIA}{\inc,\uc\textsuperscript\seq}, \withtrack{QF\_UFLRA}{\seq,\paral,\sat,\unsat,\mv}, \withtrack{QF\_UFNRA}{\unsat,\mv}, \withtrack{UF}{\uc\textsuperscript\paral}}{s}{s}
\newpage
\MakeOnePage{YicesQS}{}{\withtrack{Arith}{\sat,\fast}, \withtrack{Bitvec}{\fast}}{\withtrack{LRA}{\seq,\paral,\unsat}}{s}{}
\newpage
\MakeOnePage{Z3-alpha}{}{\withtrack{QF\_Datatypes}{\unsat}, \withtrack{QF\_LinearIntArith}{\sat}, \withtrack{QF\_NonLinearIntArith}{\seq,\paral,\sat,\unsat}, \withtrack{QF\_NonLinearRealArith}{\seq,\paral,\sat}}{\withtrack{QF\_AUFBV}{\unsat}, \withtrack{QF\_IDL}{\seq,\paral,\unsat}, \withtrack{QF\_LIRA}{\seq,\paral,\unsat}, \withtrack{QF\_NIRA}{\fast}}{s}{s}
\newpage
\MakeOnePage{Z3-Noodler}{}{\withtrack{QF\_Strings}{\seq,\paral,\sat,\unsat,\fast}}{}{}{}
\newpage